### 2017-02-03 miscellaneous/AF literature review.tex
rearranged files

%Instruct LaTeX that you would like to format an article and are using a4paper, default is US Letter size!
\documentclass{article}

%Set up a4paper size and sensible margins
\usepackage[a4paper]{geometry}

\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{amsthm, amssymb}
\usepackage[ansinew]{inputenc}
%Include support for graphics! 
\usepackage{graphicx}
%Include support for url linking
\usepackage{hyperref}

%for the image in this so it can be positoned where desired
\usepackage{float}

\title{A Literature Review of Atrial Fibrillation}
\author{ Mattia Gaggi CID:00863268\\
 \vspace{0.2cm}
Supervisor: Kim Christensen}

\date{\today}

\begin{document}
\maketitle
\vspace{-0.65cm}
\centerline{\textbf{Word Count:} 2498}
\begin{abstract}
In this paper, a summary of the physiology of Atrial Fibrillation (AF) and the research about it was carried out. After a short introduction about AF, the dynamics of electrical impulse propagation through the myocardiac tissue and the dynamics of arrhythmia were described. Then, a historical review of the main biophysical models and the Cellular Automata (CA) models was done. Finally, some ideas for future research developments were presented.
\end{abstract}

\vspace{0.2cm}

\vspace{0.2cm}

\section{Introduction}

Atrial Fibrillation is a type of supraventricular tachycardia and one of the major causes of stroke and heart failure. Population based studies estimate that over 30 millions of people worldwide are suffering from AF with a higher incidence on people over 65 years old~\cite{stats_af}. 
Only in the UK, expenses related to this condition account for more than $1\%$ of the total budget of the National Health Service, which is more than 1 billion US dollars~\cite{stats_health}. Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research.

The incredible growth in computational power over the last 30 years has enabled the implementation of increasingly more accurate models of the heart.  These computational models are a powerful tool to study the mechanisms underlying cardiac arrhythmias and are usually classified either as biophysically accurate (physiological) models or as CA models.

CA models are the simplest kind, where the cardiac tissue is modelled by a grid in which cells exist in a number of discrete states. Some of the advantages of these models over more biophysically accurate ones are the reduced computational time to run simulations and the possibility to easily implement tissue anisotropies~\cite{kim2015}.

CA models have been successful in exhibiting fibrillation behaviour and might be the key to a better understanding of AF.

\section{ Cardiac Electrophysiology}

\subsection{ Heart structure and Electrical Impulses}

Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres and finally from the ventricular endocardium to the ventricular epicardium~\cite{cardiacphys_book1987}. The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell membrane.

When a muscle cell (myocyte) is at rest, there is a built up potential difference between the outside and the inside of the cell 
(polarisation). If excited by its neighbour cells, the cell's membrane  becomes permeable to the passage of $ Na^{+}$, $Ca_{2}^{+}$ (in the cell) and $K^{+}$ ions (out of the cell) through voltage and time dependent channels.

First, the $Na^{+}$  gates are open which results in a potential spike due to the active transport of ions propagating within the myocyte (phase $0$). The intracellular potential in this phase goes from $ -75  mV$ to $+55 mV $~\cite{cardiacphys_ap2009} .

Then, the $Na^{+}$ gate closes and, after a short repolarisation period (phase $1$), the $Ca_{2}^{+}$ ions gates open causing a smaller spike in potential (phase $2$). Finally the $K^{+}$ ions leave the cell repolarising the myocyte to negative potential (phase 3).

This process causes the propagation of a short electrical impulse ($\approx 0.2s$) called the action potential (AP) which propagates through the myocytes. Once the cell has been excited, the enzyme Na/K-ATPase restores the original concentration gradients (phase $4$). For a more accurate analysis also other currents should be taken into account, even though their impact on the AP is negligible and difficult to model. After the excitation in phase $0$, the myocyte can't be excited again until phase 4 during a period of time called refractory period. The impulse propagates through junctions in the cilindrically shaped myocytes, which are arranged in bundles (fibres). It is worth mentioning that the impulse propagation changes according to the kind of muscle cells it propagates through (figure \ref{APimage}). In the case of fibrosis, myocytes are replaced by tissue that is unable to conduct~\cite{arrhythmia_fibrosis2007}. This condition is particularly relevant to disorders of impulse conduction arrhythmia which are described in the next paragraph.

 \begin{figure}[H]
\centering
\includegraphics[width=8cm]{AP_impulse}
\caption{This is an AP electric signals, where all of the phases from $0$ to $4$ are shown.~\cite{cardiacphys_ap2009} }
\label{AP_impulse}
\end{figure}

\subsection{Mechanics of Arrhythmias}

Cardiac arrhythmias are typically divided into two groups: disorders of impulse formation and disorders impulse conduction arrhythmias (reentry).

Disorders of impulse formations include irregularities in the formation of the AP, the shape of the AP and irregularities in the chemical concentrations~\cite{arrhythmia_review2012}. The SA node fires electrical impulses with the firing rate depending on the maximum diastolic potential, the threshold potential at which the AP starts and the rate of the repolarisation phase~\cite{cardiacphys_ap2009}. In addition to the SA node, also the Purkinjie fibers and the AV node can fire impulses, but at a lower rate than the SA node (that is why they are called subsidiary pacemakers). When one of the three factors regulating the firing rate of the SA note is altered, the balance between the firing rate of the SA node and the subsidiary pacemakers is broken generating a kind of impulse formation arrhythmia (called altered normal automaticity).

The normal cardiac functioning can also be disrupted by an alteration in the intracellular or extracellular chemical levels. High extracellular $K^{+}$, low $pH$ and catecholamine excess can cause abnormal impulse shape (abnormal automaticity)~\cite{arrhythmia_review2012}. Abnormal level of $Ca_{2}^{+}$ instead can cause oscillation in the cellular voltage either delaying repolarisation or causing early repolarisation ~\cite{arrhythmia_careview2016}. These voltage anomalies can also lead to disorders of impulse conduction.

Disorders of impulse conductions are mainly due to irregularities in the structure of the cardiac tissue. We will focus on these kind of disorders because they can be effectively modelled by CA models. The most common disorder of conduction is reentry which occurs when the AP impulse propagates in a self sustaining circuit~\cite{arrhythmia_review2001}.  
This is caused either by an actual physical obstacle in the cardiac tissue (anatomical reentry) or by the heterogeneities of the myocytes (functional reentry). Reentry occurs only when the impulse is sufficiently delayed in a pathway to allow for expiration of the refractory period of the fibers it propagates through. For this to happen, the length of the circuit (path-length) must exceed or equal that of the wavelength (the product of conduction velocity and refractory period).

Functional reentries can be classified according to the way they are generated. The most studied form of reentry is the circus movement reentry in which the impulse propagates in a loop around a non-conducting area. Another kind of reentry is the figure of eight reentry in which two AP impulses propagate in opposite directions around two blocks and merge into a common pathway (figure \ref{APimage}). The third kind of reentry is reflection reentry in which an impulse propagates through the same pathway back and forth. Reflection reentries happen only in presence of very impaired conduction.

\begin{figure}[H]
\vspace{1cm}
\hspace{-4cm}
\centering
\begin{minipage}[t]{.5\textwidth}
  \centering
  \includegraphics[height=0.78\linewidth]{AP_kinds}
  \label{AP}
\end{minipage}%
\begin{minipage}[t]{.5\textwidth}
  \centering
  \includegraphics[height=.78\linewidth]{reentry}
  \label{reentry}
\end{minipage}
\caption{On the left, the AP propagation in case of arrhythmia and in normal conditions~\cite{arrhythmia_review2012}.On the right,in a) a schematic drawing of a figure of eight reentry, in b) a schematic drawing of a circular reentry~\cite{arrhythmia_figurecircle}.}
\label{APimage}
\end{figure}

The last kind of reentries are rotors (called scroll waves if in 3D) in which a wavefront of AP impulses propagates around a point in a circular fashion. This happens because parts of the wavefront propagate with different speed due to anisotropies in the cardiac tissue. Unlike the circus movement reentry, there is no fixed wavelenght in rotors. This is because electrotonic gradients established between cells at the core and cells in the near vicinity shorten the action potentials of cells near the core~\cite{arrhythmia_rotorsreview2013} . Another key difference with the circus movement reentry is that the tissue in the center of the rotor is conducting, which allows rotors to shift to different areas. Recent developments in medical imaging techniques shed light on the dynamics of rotors and their role in AF~\cite{arrhythmia_rotorsreview2013}. In a 2012 study, physiologically guided computational mapping showed the presence of rotors in 96\% of the patients with AF studied~\cite{arrhythmia_rotors2012}. Another study shows that, during AF, many small rotors are generated with most of them located near the pulmonary vein~\cite{arrhythmia_pulmonaryvein1998}, even though other studies also suggests that rotors can be generated elsewhere~\cite{arrhythmia_venacava2000}.

\subsection{ Current Treatments }

A cure for atrial fibrillation is a major goal in cardiac electrophysiology. Current treatments can be divided into three categories: rhythm control, rate control and ablation procedures~\cite{article_aftreatments}.

Rhythm control and rate control mainly involve the use of drugs to keep the heart rate regular by affecting the chemical concentrations of ions and the blood coagulation. Both these strategies have a plethora of side effects and their success rate is heavily dependent on the condition of the patient.

Ablation procedures instead are surgical procedures acting on the area of the heart affected by AF. Among these, radio frequency cathered ablation (RFCA) has proved to be effective with a success rate that can be as high as 80\% in patients affected by AF arising in tissue near the pulmonary vein ~\cite{cardiacphys_raterythmcontrol}. In RFCA, high-frequency, low-voltage radio waves are sent by two electrodes burning the area affected, avoiding therefore the risks of open-heart surgery. Other techniques are evolving to address the challenge of a catheter-based cure for all forms of AF (balloon-based technologies using cryoablation,ultrasound, and laser)~\cite{article_afablation}, but, in order for them to be successful, a better understanding of the mechanisms underlying AF is needed. That is why computational models could be the first step towards a definitive cure for AF.

\section{ Current research}

\subsection{ Biophysically Accurate Models }

Many different models of the impulse propagation in myocytes have been developed in the last 60 years~\cite{physmodel_review}. 
One of the very first model was the Fitz Hugh Nagumo model (1969) ~\cite{physmodel_fhnrabbit2014} where muscle cells were modelled with an excitability variable and a recovery variable and whose dynamics were governed by simple differential equations.  Then, many other models were created trying to describe more accurately the $Na+$,$Ca2+$ and $K+$ currents. The models can be divided into monodomain models, which consider an isotropic medium, and bidomain models, which instead take into account both the intracellular and extracellular environment~\cite{kim_computationalmodels2001}.

Other noteworthy models are the Noble-Noble model (1984) describing the behaviour of the SA node and the Ramirez-Nattel-Courtemanche model (2000) describing the behaviour of canine  atrial cells~\cite{physmodel_review}. These models are quite convoluted, involving the use of many variables.

Another step forward was done first with the Karma model (1993), and then with the Fenton-Karma model (1998) where the ion currents were modelled as fast currents and slow currents ~\cite{physmodel_review}.

The Fenton-Karma model has only 3 variables and it is today one of the most used models, showing a certain degree of success in obtaining arrhythmias~\cite{physmodel_fkkktcomparison}. The main shortcoming of all these model is not taking into account anisotropies in the myocardiac tissue. As a result, functional arrhythmia can only be generated in these model by using specific initial conditions.

\subsection{ Cellular Automata Models }

In CA models the muscle fiber are modelled by a 2D (or 3D) grid made up by cells linked to each other~\cite{cellauto_weimar1992}. The cells can be in an excited state, a refractory state or an excitable state according to a discrete parameter (the currents of ions are not modelled explicitly).

The very first attempt of building a CA model of myocytes was made by Moe in 1964, modelling the myocytes as exagons in 2D~\cite{cellauto_review}. The model was then improved by Markus in 1990 who used squares of unit 1 to describe the cells (where the exact position of each cell was a random point in each cell). In the model, a variable $S$ with $0<S<N$ is used to describe the excitation states (being $0$, $N$ the resting end excited state respectively). The neighbours of each cell are the cells within radius one unit from the centre of the cell. If the sum over the neighbours $S$ is greater than a specific value, the cell is excited.  In another paper a similar version of this model was used to show that waves propagating with a certain curvature can generate arrhythmias and rotors ~\cite{cellauto_curvature1990}.

Another noteworthy model is the Weimar model (1992)\cite{cellauto_review} which modifies the neighbours rule giving more excitation weight to the closer cells. In spite of the ability to show some form of arrhythmia, these models are not very computationally efficient. 
A further step was made in 2005 in the paper "A Probabilistic Model of Cardiac Electrical Activity Based on a Cellular Automata System"~\cite{cellauto_model2005}. This model shares many features with the Markus model except that, for the first time, the depolarisation of the cell membrane was modelled as a probabilistic phenomenon.

Then, in the publication "Simple Model for Identifying Critical Regions in Atrial Fibrillation"~\cite{kim2015} an even simpler model was produced. The cells were arranged in a squared grid with horizontal connections, with a percentage of transversal connections (to reproduce the effect of heterogeneities) and a percentage of unexcitable cells. Another key difference with the previous models is that each cell becomes excited the moment one of its four neighbours is excited. The results from this model show that there is a threshold value of vertical connections beyond which AF is produced spontaneously (without changing refractory period or conduction speed). This matches real data according to which AF is correlated with structural heterogeneities~\cite{arrhythmia_fibrosis2007}. 
The model also shows that burning the area affected by AF effectively stops the arrhythmia, which has been recently discovered to be the case ~\cite{article_kimpaper}.

All the models described up to now are 2D models and 3D models haven't really been developed. An attempt was done in 1998~\cite{cellauto_3dmodel}, but the computational resources available were too limited to obtain meaningful results. Another publication from 2015 "Modeling Sources of Atrial Fibrillation on a Triangulated Sphere" shows a cellular automaton modelled on a 2D sphere where the cells were modelled as a simple graph~\cite{cellauto_sphere2015} and where the excitability rules were similar to the ones developed by Weimar. The results show rotors for specific values of the refractory period and the conduction speed.

\begin{figure}[H]
\centering
\begin{minipage}[t]{.5\textwidth}
  \centering
  \includegraphics[height=0.73\linewidth]{rotor}
  \label{rot}
\end{minipage}%
\begin{minipage}[t]{.5\textwidth}
  \centering
  \includegraphics[height=.73\linewidth]{rotorkim}
  \label{rotkim}
\end{minipage}
\caption{On the left, the dynamics of a rotor generated in a biophysical model~\cite{arrhythmia_rotorsreview2013}. On the right, the dynamics of a rotor spontaneously arising in a CA model~\cite{kim2015}.}
\label{rotor}
\end{figure}

\section{Possible Developments}

Even though models that simulate AF have been reproduced, the mechanisms underlying AF are still obscure. The role of $Ca_{2}^{+}$ in the AP is not completely understood and fibrillation doesn't spontaneously arise in the current biophysically accurate models available~\cite{arrhythmia_ca2016}\cite{arrhythmia_careview2016}. As for CA models, improvements and further testing are necessary.

A few ideas can be used to test the efficiency of CA models. Optical imaging techniques have shown a correlation between rotor frequency and action potential shortening~\cite{physmodel_rotorspeed2015}. By comparing the data from the model with real data, the accuracy of a CA model can be tested.

In addition, in a very recent publication~\cite{physmodel_heterogeneities2015} the effect of patches with different refractory period was observed and analysed in a biophysical model. The same can be done in a CA model by modelling patches of tissue heterogeineties and a comparison can be made. Third, in another publication~\cite{arrhythmia_rotorsreview2013} a scaling relation between ventricular fibrillation frequency and body mass index was found. Finding a similar scaling relation in a CA model will give further credit to the validity of the model. Another further research development could be mimicking the effects of different anti-arrhythmic drugs in CA model and  compare the results with real data. Finally, the construction and implementation of a functioning 3D CA model could be carried out.

 \newpage

\section{Bibliography}

\bibliographystyle{ieeetr}
\bibliography{Bibliography}
\end{document}

### 2017-02-03 General Network/Network.py
Add files via upload

@@ -1,0 +1,308 @@
+import numpy as np
+from scipy.sparse import csr_matrix
+import operator
+from itertools import chain
+import time
+
+
+def decision(p):
+    if np.random.random()<p:
+        return True
+    else:
+        return False
+def trial():
+    nodes=range(40000)
+    a=np.arange(40000)
+    b=np.arange(40000)
+    c=np.arange(40000)
+    np.random.shuffle(b)
+    np.random.shuffle(c)
+    l=[]
+    for e in np.arange(40000):
+        l.append((a[e],b[e]))
+        l.append((c[e],b[e]))
+        l.append((a[e],c[e]))
+    array_transv=[]
+    p=[0,1]
+    return nodes,l,array_transv,1,p,0.5,0.5
+        
+    
+
+class create_network:
+    def __init__(self,array_nodesindices,array_vertical,array_transv,p_transv,impulse_start,p_dysf,p_unexcitable):
+        """
+        this inputs array_nodesindices as a list of faces indices,
+        
+        array_vertical and array_transv as  list of tuples
+        """
+        t0=time.time()
+            
+        # watch it because here we assume that faces go from 0 to n
+        
+        self.p_dysf=p_dysf
+        self.p_transv=p_transv
+        self.p_unexcitable=p_unexcitable
+        self.excitation=50
+        self.heartbeatssteps=220
+        self.totalruns = 0
+    
+        self.array_vertical=array_vertical
+        self.impulse_start=impulse_start
+        self.size=len(array_nodesindices)
+        self.nodes=np.zeros(self.size)
+       
+        self.array_transv=[]
+        self.excited=[]
+        self.dysf=np.random.rand(self.size)  #create array of dysf cells
+        self.dysf[self.dysf < self.p_dysf] = 1
+        self.dysf[self.dysf != 1] = 0
+        self.unexcited=np.random.rand(self.size)
+        self.array_alltransv=array_transv
+               
+        for elements in self.array_alltransv: #append transversal connections to a new list according to the probability of transv connect
+          if decision(self.p_transv):
+              self.array_transv.append(elements)
+        
+        self.connections = {}   #create dictionary with connections ex d={1:[0],2:[3,4],.....}
+        for key,value in (self.array_vertical+self.array_transv):
+            try:
+                self.connections[key].append(value)
+            except KeyError:
+                self.connections[key]=[value]
+        
[…]

### 2017-02-03 General Network/Network2.py

@@ -1,0 +1,312 @@
+import numpy as np
+from scipy.sparse import csr_matrix
+import operator
+from itertools import chain
+import time
+
+
+def decision(p):
+    if np.random.random()<p:
+        return True
+    else:
+        return False
+def trial():
+    nodes=range(40000)
+    a=np.arange(40000)
+    b=np.arange(40000)
+    c=np.arange(40000)
+    np.random.shuffle(b)
+    np.random.shuffle(c)
+    l=[]
+    for e in np.arange(40000):
+        l.append((a[e],b[e]))
+        l.append((c[e],b[e]))
+        l.append((a[e],c[e]))
+    array_transv=[]
+    p=[0,1]
+    return nodes,l,array_transv,1,p,0.5,0.5
+        
+    
+
+class create_network:
+    def __init__(self,array_nodesindices,array_vertical,array_transv,p_transv,impulse_start,p_dysf,p_unexcitable):
+        """
+        this inputs array_nodesindices as a list of faces indices,
+        
+        array_vertical and array_transv as  list of tuples
+        """
+        t0=time.time()
+            
+        # watch it because here we assume that faces go from 0 to n
+        
+        self.p_dysf=p_dysf
+        self.p_transv=p_transv
+        self.p_unexcitable=p_unexcitable
+        self.excitation=50
+        self.heartbeatssteps=220
+        self.totalruns = 0
+    
+        self.array_vertical=array_vertical
+        self.impulse_start=impulse_start
+        self.size=len(array_nodesindices)
+        self.nodes=np.zeros(self.size)
+       
+        self.array_transv=[]
+        self.excited=[]
+        self.dysf=np.random.rand(self.size)  #create array of dysf cells
+        self.dysf[self.dysf < self.p_dysf] = 1
+        self.dysf[self.dysf != 1] = 0
+        self.unexcited=np.random.rand(self.size)
+        self.array_alltransv=array_transv
+               
+        for elements in self.array_alltransv: #append transversal connections to a new list according to the probability of transv connect
+          if decision(self.p_transv):
+              self.array_transv.append(elements)
+        
+        self.connections = {}   #create dictionary with connections ex d={1:[0],2:[3,4],.....}
+        for key,value in (self.array_vertical+self.array_transv):
+            try:
+                self.connections[key].append(value)
+            except KeyError:
+                self.connections[key]=[value]
[…]

### 2017-02-03 General Network/Network_anim.py

@@ -1,0 +1,405 @@
+import numpy as np
+from scipy.sparse import csr_matrix
+import operator
+from itertools import chain
+import time
+from matplotlib import animation
+from matplotlib import pyplot as plt
+from matplotlib import cm
+from mpl_toolkits.mplot3d import Axes3D
+
+def decision(p):
+    if np.random.random()<p:
+        return True
+    else:
+        return False
+def trial():
+    nodes=range(40000)
+    a=np.arange(40000)
+    b=np.arange(40000)
+    c=np.arange(40000)
+    np.random.shuffle(b)
+    np.random.shuffle(c)
+    l=[]
+    for e in np.arange(40000):
+        l.append((a[e],b[e]))
+        l.append((c[e],b[e]))
+        l.append((a[e],c[e]))
+    array_transv=[]
+    p=[0,1]
+    return nodes,l,array_transv,1,p,0.5,0.5
+        
+    
+
+class create_network:
+    def __init__(self,array_nodesindices,array_vertical,array_transv,p_transv,impulse_start,p_dysf,p_unexcitable):
+        """
+        this inputs array_nodesindices as a list of faces indices,
+        
+        array_vertical and array_transv as  list of tuples
+        """
+        t0=time.time()
+            
+        # watch it because here we assume that faces go from 0 to n
+        
+        self.p_dysf=p_dysf
+        self.p_transv=p_transv
+        self.p_unexcitable=p_unexcitable
+        self.excitation=50
+        self.heartbeatssteps=220
+        self.totalruns = 0
+    
+        self.array_vertical=array_vertical
+        self.impulse_start=impulse_start
+        self.size=len(array_nodesindices)
+        self.nodes=np.zeros(self.size)
+       
+        self.array_transv=[]
+        self.excited=[]
+        self.dysf=np.random.rand(self.size)  #create array of dysf cells
+        self.dysf[self.dysf < self.p_dysf] = 1
+        self.dysf[self.dysf != 1] = 0
+        self.unexcited=np.random.rand(self.size)
+        self.array_alltransv=array_transv
+               
+        for elements in self.array_alltransv: #append transversal connections to a new list according to the probability of transv connect
+          if decision(self.p_transv):
+              self.array_transv.append(elements)
+        
+        self.connections = {}   #create dictionary with connections ex d={1:[0],2:[3,4],.....}
+        for key,value in (self.array_vertical+self.array_transv):
+            try:
+                self.connections[key].append(value)
+            except KeyError:
+                self.connections[key]=[value]
+        
[…]

### 2017-02-03 General Network/Network_anim2.py

@@ -1,0 +1,470 @@
+import numpy as np
+from scipy.sparse import csr_matrix
+import operator
+from itertools import chain
+import time
+from matplotlib import animation
+from matplotlib import pyplot as plt
+from matplotlib import cm
+from mpl_toolkits.mplot3d import Axes3D
+
+def decision(p):
+    if np.random.random()<p:
+        return True
+    else:
+        return False
+def trial():
+    nodes=range(40000)
+    a=np.arange(40000)
+    b=np.arange(40000)
+    c=np.arange(40000)
+    np.random.shuffle(b)
+    np.random.shuffle(c)
+    l=[]
+    for e in np.arange(40000):
+        l.append((a[e],b[e]))
+        l.append((c[e],b[e]))
+        l.append((a[e],c[e]))
+    array_transv=[]
+    p=[0,1]
+    return nodes,l,array_transv,1,p,0.5,0.5
+        
+    
+
+class create_network:
+    def __init__(self,array_nodesindices,array_vertical,array_transv,p_transv,impulse_start,p_dysf,p_unexcitable):
+        """
+        this inputs array_nodesindices as a list of faces indices,
+        
+        array_vertical and array_transv as  list of tuples
+        """
+        t0=time.time()
+            
+        # watch it because here we assume that faces go from 0 to n
+        
+        self.p_dysf=p_dysf
+        self.p_transv=p_transv
+        self.p_unexcitable=p_unexcitable
+        self.excitation=50
+        self.heartbeatssteps=220
+        self.totalruns = 0
+    
+        self.array_vertical=array_vertical
+        self.impulse_start=impulse_start
+        self.size=len(array_nodesindices)
+        self.nodes=np.zeros(self.size)
+       
+        self.array_transv=[]
+        self.excited=[]
+        self.dysf=np.random.rand(self.size)  #create array of dysf cells
+        self.dysf[self.dysf < self.p_dysf] = 1
+        self.dysf[self.dysf != 1] = 0
+        self.unexcited=np.random.rand(self.size)
+        self.array_alltransv=array_transv
+               
+        for elements in self.array_alltransv: #append transversal connections to a new list according to the probability of transv connect
+          if decision(self.p_transv):
+              self.array_transv.append(elements)
+        
+        self.connections = {}   #create dictionary with connections ex d={1:[0],2:[3,4],.....}
+        for key,value in (self.array_vertical+self.array_transv):
+            try:
+                self.connections[key].append(value)
+            except KeyError:
+                self.connections[key]=[value]
+        
[…]

### 2017-02-03 General Network/ico_propagation.py

@@ -1,0 +1,421 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Jan  5 17:03:51 2017
+
+@author: tigan_5ytncvu
+"""
+from mpl_toolkits.mplot3d import Axes3D
+from scipy.spatial import Delaunay
+from scipy.spatial import ConvexHull
+import matplotlib.tri as mtri
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+t = (1.0 + (5.0)**(0.5) )/ 2.0
+
+icosahedron_vertices = [[-1,  t,  0],
+                            [ 1,  t,  0],
+                            [-1, -t,  0],
+                            [1, -t,  0],
+                            [ 0, -1,  t],
+                            [0,  1,  t],
+                            [ 0, -1, -t],
+                            [ 0,  1, -t],
+                            [ t,  0, -1],
+                            [t,  0,  1],
+                            [-t,  0, -1],
+                            [-t,  0,  1]]
+
+faces = np.array(
+            [[0, 11, 5],
+            [0, 5, 1],
+            [0, 1, 7],
+            [0, 7, 10],
+            [0, 10, 11],
+            [1, 5, 9],
+            [5, 11, 4],
+            [11, 10, 2],
+            [10, 7, 6],
+            [7, 1, 8],
+            [3, 9, 4],
+            [3, 4, 2],
+            [3, 2, 6],
+            [3, 6, 8],
+            [3, 8, 9],
+            [4, 9, 5],
+            [2, 4, 11],
+            [6, 2, 10],
+            [8, 6, 7],
+            [9, 8, 1]])
+
+
+
+class Sphere:
+    
+    def __init__(self, vertices = icosahedron_vertices, faces = faces, recursion_level = 4 ):
+        
+        self.icosahedron_vertices = vertices
+        self.faces = faces
+        self.recursion_level = recursion_level
+        self.ch = []
+
+
+    
+    def normalize_v3(self, arr):
+        ''' Normalize a numpy array of 3 component vectors shape=(n,3) '''
+    
+        lens = arr[:,0]**2 + arr[:,1]**2 + arr[:,2]**2 
+        lens = np.sqrt(lens)
+        arr[:,0] /= lens
+        arr[:,1] /= lens
+        arr[:,2] /= lens        
+        return arr
+        
+    def normalize(self, arr):
+        ''' Normalize components '''
+     
[…]

### 2017-02-03 Sphere Model/General Network/Network_anim2.py
small

@@ -1,5 +1,4 @@
 import numpy as np
-from scipy.sparse import csr_matrix
 import operator
 from itertools import chain
 import time
@@ -7,6 +6,8 @@ from matplotlib import animation
 from matplotlib import pyplot as plt
 from matplotlib import cm
 from mpl_toolkits.mplot3d import Axes3D
+import ico_propagation as sp
+
 
 def decision(p):
     if np.random.random()<p:
@@ -185,7 +186,7 @@ class run:
         
         
     def propagate_n(self):
-        for times in range(runs):
+        for times in range(self.runs):
             if self.store==True:
                 
                 if self.network.totalruns%self.network.heartbeatssteps == 0: #self.time%self.network.heartbeatssteps==0:
@@ -198,7 +199,7 @@ class run:
             self.network.onestep()
             
     def propagate_a(self):
-        for times in range(runs):
+        for times in range(self.runs):
             if self.store==True:
                 
                 if self.network.totalruns%self.network.heartbeatssteps == 0: #self.time%self.network.heartbeatssteps==0:
@@ -287,177 +288,178 @@ class run:
         return self.surf,
         
                 
-import ico_propagation as sp
 
 
-t = (1.0 + (5.0)**(0.5) )/ 2.0
-
-icosahedron_vertices = [[-1,  t,  0],
-                            [ 1,  t,  0],
-                            [-1, -t,  0],
-                            [1, -t,  0],
-                            [ 0, -1,  t],
-                            [0,  1,  t],
-                            [ 0, -1, -t],
-                            [ 0,  1, -t],
-                            [ t,  0, -1],
-                            [t,  0,  1],
-                            [-t,  0, -1],
-                            [-t,  0,  1]]
-
-faces = np.array(
-            [[0, 11, 5],
-            [0, 5, 1],
-            [0, 1, 7],
-            [0, 7, 10],
-            [0, 10, 11],
-            [1, 5, 9],
-            [5, 11, 4],
-            [11, 10, 2],
[…]

### 2017-02-03 Sphere Model/General Network/Network_anim2.py
small changes

@@ -395,9 +395,9 @@ class Create_Sphere_and_Connections:
                     [8, 6, 7],
                     [9, 8, 1]])
         
-        self.s = sp.Sphere(vertices = self.icosahedron_vertices, faces = self.faces, recursion_level = 5 )
+        self.s = sp.Sphere(vertices = self.icosahedron_vertices, faces = self.faces, recursion_level = 6 )
         self.s.ch = self.s.construct_icosphere()
-        x, y, z   = self.s.ch.points[self.s.ch.vertices][:,0],self.s.ch.points[s.ch.vertices][:,1], self.s.ch.points[self.s.ch.vertices][:,2]
+        x, y, z   = self.s.ch.points[self.s.ch.vertices][:,0],self.s.ch.points[self.s.ch.vertices][:,1], self.s.ch.points[self.s.ch.vertices][:,2]
         vertex1, vertex2, vertex3 = self.s.ch.points[self.s.ch.simplices[:,0]],self.s.ch.points[self.s.ch.simplices[:,1]], self.s.ch.points[self.s.ch.simplices[:,2]]# hull_pts[ch.simplices[:,1]], hull_pts[ch.simplices[:,2]]
         
         

### 2017-02-03 Sphere Model/General Network/ico_propagation.py

@@ -53,7 +53,7 @@ faces = np.array(
 
 class Sphere:
     
-    def __init__(self, vertices = icosahedron_vertices, faces = faces, recursion_level = 4 ):
+    def __init__(self, vertices = icosahedron_vertices, faces = faces, recursion_level = 6 ):
         
         self.icosahedron_vertices = vertices
         self.faces = faces

### 2017-02-03 Sphere Model/General Network/ico_propagation.py
a

@@ -53,7 +53,7 @@ faces = np.array(
 
 class Sphere:
     
-    def __init__(self, vertices = icosahedron_vertices, faces = faces, recursion_level = 6 ):
+    def __init__(self, vertices = icosahedron_vertices, faces = faces, recursion_level = 4 ):
         
         self.icosahedron_vertices = vertices
         self.faces = faces

### 2017-02-03 Network_Clean.py
Add files via upload

@@ -1,0 +1,360 @@
+import numpy as np
+from scipy.sparse import csr_matrix
+import operator
+from itertools import chain
+import time
+from matplotlib import animation
+from matplotlib import pyplot as plt
+from matplotlib import cm
+from mpl_toolkits.mplot3d import Axes3D
+
+def decision(p):
+    if np.random.random()<p:
+        return True
+    else:
+        return False
+def trial():
+    nodes=range(40000)
+    a=np.arange(40000)
+    b=np.arange(40000)
+    c=np.arange(40000)
+    np.random.shuffle(b)
+    np.random.shuffle(c)
+    l=[]
+    for e in np.arange(40000):
+        l.append((a[e],b[e]))
+        l.append((c[e],b[e]))
+        l.append((a[e],c[e]))
+    array_transv=[]
+    p=[0,1]
+    return nodes,l,array_transv,1,p,0.5,0.5
+        
+    
+
+class create_network:
+    def __init__(self,array_nodesindices,array_vertical,array_transv,p_transv,impulse_start,p_dysf,p_unexcitable):
+        """
+        this inputs array_nodesindices as a list of faces indices,
+        
+        array_vertical and array_transv as  list of tuples
+        """
+        t0=time.time()
+            
+        # watch it because here we assume that faces go from 0 to n
+        
+        self.p_dysf=p_dysf
+        self.p_transv=p_transv
+        self.p_unexcitable=p_unexcitable
+        self.excitation=50
+        self.heartbeatssteps=220
+        self.totalruns = 0
+    
+        self.array_vertical=array_vertical
+        self.impulse_start=impulse_start
+        self.size=len(array_nodesindices)
+        self.nodes=np.zeros(self.size)
+       
+        self.array_transv=[]
+        self.excited=[]
+        self.dysf=np.random.rand(self.size)  #create array of dysf cells
+        self.dysf[self.dysf < self.p_dysf] = 1
+        self.dysf[self.dysf != 1] = 0
+        self.unexcited=np.random.rand(self.size)
+        self.array_alltransv=array_transv
+               
+        for elements in self.array_alltransv: #append transversal connections to a new list according to the probability of transv connect
+          if decision(self.p_transv):
+              self.array_transv.append(elements)
+        
+        self.connections = {}   #create dictionary with connections ex d={1:[0],2:[3,4],.....}
+        for key,value in (self.array_vertical+self.array_transv):
+            try:
+                self.connections[key].append(value)
+            except KeyError:
+                self.connections[key]=[value]
+        
[…]

### 2017-02-06 Sphere Model/Network_Clean.py
folders

@@ -7,7 +7,8 @@ from matplotlib import animation
 from matplotlib import pyplot as plt
 from matplotlib import cm
 from mpl_toolkits.mplot3d import Axes3D
-
+                
+import ico_propagation as sp
 def decision(p):
     if np.random.random()<p:
         return True
@@ -261,8 +262,7 @@ class run:
         
         return self.surf,
         
-                
-import ico_propagation as sp
+
 
 
 t = (1.0 + (5.0)**(0.5) )/ 2.0

### 2017-02-18 FinalAF.py
Merge remote-tracking branch 'origin/master'
# Conflicts:
#	Sphere Model/New Network With Proper Propagation/risk_curve_Sphere.py

@@ -347,6 +347,11 @@ class run: #Class to run code
                 if len(self.num_excited) > 440 and all(i <= 220 for i in self.num_excited[-441:]) :
                     self.stopfibrillation()
 
+    
+    
+    
+    
+    
     def fibrillation(self):
             
         if self.infibrillation==False:

### 2017-02-18 risk_curve.py

@@ -37,6 +37,7 @@ kisherror = kishmatrix[:,2]
 p_transversalconn= kishnu #np.arange(0,0.6,0.05) #nu
 
 
+
 time=100000
 number_of_systems=50
 
@@ -50,8 +51,10 @@ risk=[]
 riskstd=[]
 t_fib_data = []
 t_fib_temp = []
+
 number_fib = []
 n_fib_data = []
+
 n_fib_avg = []
 n_fib_err = []
 t_in_fib = []

### 2017-02-19 Sphere Model/New Network With Proper Propagation/Network_Projection_withrestitution.py
implemented restitution

@@ -1,0 +1,670 @@
+import numpy as np
+from scipy.sparse import csr_matrix
+import operator
+from itertools import chain
+import time
+from matplotlib import animation
+from matplotlib import pyplot as plt
+from matplotlib import cm
+from mpl_toolkits.mplot3d import Axes3D
+import pylab as pl
+from matplotlib import collections  as mc
+from mpl_toolkits.mplot3d.art3d import Line3DCollection
+import types
+import pickle
+import ico_prop_projection as sp
+
+def decision(p):
+    if np.random.random()<p:
+        return True
+    else:
+        return False
+
+
+
+def trial():
+    nodes=range(40000)
+    a=np.arange(40000)
+    b=np.arange(40000)
+    c=np.arange(40000)
+    np.random.shuffle(b)
+    np.random.shuffle(c)
+    l=[]
+    for e in np.arange(40000):
+        l.append((a[e],b[e]))
+        l.append((c[e],b[e]))
+        l.append((a[e],c[e]))
+    array_transv=[]
+    p=[0,1]
+    return nodes,l,array_transv,1,p,0.5,0.5
+        
+    
+
+class create_network:
+    def __init__(self,array_nodesindices,array_vertical,array_transv,p_transv,impulse_start,p_dysf,p_unexcitable, excitation, hbs,recursion=5):
+        """
+        this inputs array_nodesindices as a list of faces indices,
+        
+        array_vertical and array_transv as  list of tuples
+        """
+        t0=time.time()
+            
+        # watch it because here we assume that faces go from 0 to n
+        
+        self.p_dysf=p_dysf
+        self.p_transv=p_transv
+        self.p_unexcitable=p_unexcitable
+        self.excitation=excitation
+        self.heartbeatssteps=hbs
+        self.totalruns = 0
+        
+        self.recursion=recursion
+        print "recursion=",self.recursion,"make sure the restitution curve is set accordingly"
+        
+        
+        self.array_vertical=array_vertical
+        self.impulse_start=impulse_start
+        self.size=len(array_nodesindices)
+        self.nodes=np.zeros(self.size)
+       
+        self.array_transv=[]
+        self.excited=[]
+        self.dysf=np.random.rand(self.size)  #create array of dysf cells
+        self.dysf[self.dysf < self.p_dysf] = 1
+        self.dysf[self.dysf != 1] = 0
+        self.unexcited=np.random.rand(self.size)
+        self.array_alltransv=array_transv
+       
+        
[…]

### 2017-02-20 Literature Review/DraftReport.tex
POSTERTINGZZZZZZZZZZZZZ

\documentclass[twocolumn]{article}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{commath}
\usepackage{adjustbox}
%\usepackage[sc]{mathpazo}
\usepackage[margin=0.5in]{geometry}
%\usepackage{multicol}
\usepackage{mathtools}
\usepackage{abstract}
\usepackage[compact]{titlesec}
\DeclareMathOperator{\sech}{sech}
\newcommand{\matr}[1]{\mathbf{#1}}
%\usepackage{placeins}
%\usepackage{textcomp}
\graphicspath{ {H:/} }

\begin{document}

\title {\textbf{Literature Review: Complexity science approach to Atrial Fibrillation}}
\author{Tigany Zarrouk\\ CID: 00823932\\\textbf{Supervisor}: Kim Christensen \\\textbf{Word Count}: 2488.}

\date{}
\maketitle

\begin{abstract}

Modelling Atrial Fibrillation (AF) to ascertain its causes accurately, allowing for effective treatment, has been extensively studied over the past few decades. It is the largest cause of strokes linked with cardiac arrhythmia in aging populations and is due to impaired contractility of the atria in the heart, leading to blood clot formation. Action potential propagation, mediated by ionic species, are seen to be irregular in AF, which causes some atrial myocytes to contract at 300+ times per minute, much faster than healthy hearts which contract at 60-100 times per minute. This beating is regulated by the sinoatrial node which initiates action potentials that propagate through myocytes, causing contractions. Altered refractory period and conduction speed of signals has been seen to promote AF in various models and is caused physically by the fibrosis of cells, reducing the coupling of electrical signals transversally to some myocytes; giving rise to reentry circuits. The most prominent reentry model, the spiral wave model, has been seen to more closely follow clinical and experimental observations than other models of wavefront action. Cellular automata modelling paves the way for large scale simulations of tissue in the heart without the need to solve differential equations, which can be computationally costly. Moore neighbourhoods are seen to be more effective at producing isotropy of spiral waves, reducing flat edges, in rectilinear cellular automata models. Modelling scroll waves, using hybrid cellular automata models, using novel methods to find rotors and replicating the morphology of the heart with its heterogeneities more closely seem to be viable directions for further research.

\end{abstract}

\section{\textbf{\underline{Introduction}}}

Medical science and the study of disease has been a crucial marker of society's success and development through the ages. Understanding disease gives insights into how to treat and diagnose serious conditions, of which can be fatal. One such condition that has eluded explanation and effective treatment is Atrial Fibrillation (AF), which is the leading cause of Ischaemic stroke in people over 75---seemingly due to incomplete contraction which leads to the formation of blood clots that can travel to the brain \cite{Hart}. It has been known that there are cardiovascular diseases that are predisposed to initiate AF, such as Congestive Heart Failure, Pericarditis and Hypertension through a combination of atrial pressure and dilation, but the precise mechanics are unknown. AF and VF (ventricular fibrillation) are thought to be caused by the breakup of waves of electrical impulses, that induce contraction, into interfering wavelets that produce chaotic behaviour.

Healthy hearts have a regular, sinus  rhythm which is moderated by the sinoatrial node located in the right atrium of the heart. Fibrillatory activity in the atrium is characterised by a hastened activation of cells: atrial cells activate 300-600 times a minute in contrast to the expected sinus rhythm of 60-100 beats per minute.% Ectopic firing has been observed in mainly in the left atria.

Current treatment involves introducing antiarrhythmic drugs of which side effects include proarrhythmia, which can cause sudden death \cite{Nattel1}. Surgical methods primarily involve the Maze procedure, by which a maze-like pattern is cut into the atria such that there is only one path that the electrical impulse can take from the sinoatrial node to the atrioventricular node. The treatment is usually successful but still may lead to fibrillation \cite{Tsui}. The ablation of cells in has been shown to treat AF therapeutically. Catheter ablation has had suboptimal success rates. Electrogram-guided ablation is a burgeoning area of research. Complex-fractionated atrial electrograms (CFAEs) are proposed as sites for ablation as they seem to be critical for AF \cite{Greek}. % These sites seem to be mainly focussed around the pulmonary veins.

Electrograms are used for non-contact mapping and can be used to elucidate areas of fibrillatory activity. The most popular types are unipolar and bipolar. Unipolar electrograms have problems subtracting far-field signal and the wavefront direction affects the amplitude of the wave recorded---in contrast to bipolar electrograms---but the morphology of the surface can indicate the direction of the wavefront. Bipolar electrodes offer the advantage of a superior signal-to-noise ratio but with markedly less amplitude in sensing depolarization \cite{Vincent}. 
%` whereas bipolar electrograms do not experience these problems.

\section{\textbf{\underline{Electrophysiology of the heart}}}

Signals propagate through the heart by an action potential arising due to the interplay of ions with pacemaker cells, in the sinoatrial and atrioventricular nodes, and myocardial cells in cardiac muscle \cite{Augustus}. See figure 1 for a diagram of the action potentials involved.

 Myocardial cells in the heart have a potential difference and concentration gradient intrinsic to their structure: Na$^{+}$ and Cl$^{-}$ ions are higher in concentration outside of the cell, while K$^{+}$ and organic anions (A$^{-}$) are more concentrated on the inside. Ca$^{2+}$  ions are higher in concentration extracellularly but they are bound, mainly in the sarcoplasmic reticulum. 
This gives a resting membrane potential of $-75$mV, when no action potential is propagating, as the only ion that is allowed free movement is K$^{+}$ of which the concentration gradient and potential difference balance out the efflux (due to the concentration gradient) and influx (due to the negative resultant potential difference). This is phase 4 of the action potential. There are gates through which specific ions can flow into or out of the cell when a threshold voltage has been reached. After a sufficient stimulation, positive ions (mainly Na$^{+}$)  flood into the myocardial cell which causes rapid depolarisation (phase 0). Once depolarised enough, after a small efflux of K$^{+}$ and  Cl$^{-}$ ions (phase 1), Ca$^{2+}$ ion voltage-gated channels open, which induces Ca$^{2+}$ release from intracellular stores. This gives a plateau of potential difference (phase 2) where an influx of  K$^{+}$ ions sustains the voltage against the efflux of  Ca$^{2+}$ ions. The free calcium ions cause myocardial contraction. After a this long plateau phase,  K$^{+}$ ion channels reopen to repolarise the cell (phase 3) to get ready for another contraction, reverting to phase 4. The period where a cell cannot respond to stimuli to induce contraction is the refractory period.

%5 phases of the normal action potential:

%Phase 4, or the resting potential, is stable at ≈−90 mV in normal working myocardial cells.

%Phase 0 is the phase of rapid depolarization. The membrane potential shifts into positive voltage range. This phase is central to rapid propagation of the cardiac impulse (conduction velocity, θ=1 m/s).

%Phase 1 is a phase of rapid repolarization. This phase sets the potential for the next phase of the action potential.

%Phase 2, a plateau phase, is the longest phase. It is unique among excitable cells and marks the phase of calcium entry into the cell.

%Phase 3 is the phase of rapid repolarization that restores the membrane potential to its resting valu
%Through this electrical conduction, the cardiac muscle (myocardium) contracts.

In the sinoatrial node, the depolarising current is primarily fulfilled by Ca$^{2+}$ ion movement, which is slow in comparison to the fast Na$^{+}$ currents of the myocardial cells. SA node action potentials have only three phases. Phase 4 is the spontaneous depolarization (pacemaker potential) that triggers the action potential. Phase 0 is the depolarization phase of the action potential. This is followed by phase 3, repolarization. Once the cell is completely repolarized, the cycle is spontaneously repeated.

\begin{figure}
\caption[short title]{Diagram of the action potential in a myocardial cell (A) and in sinoatrial cell (B). Numbers indicate phases of the action potential.  ERP = Effective refractory period \cite{Mizzi}}
\centering
\includegraphics[width = 0.5\textwidth]{newactionpot}
\end{figure}

Ectopic firing---from a focus, not located at the Sinoatrial node, at which action potentials are initiated---can arise from 'afterdepolarisations': the free intracellular Ca$^{2+}$ concentration rises sharply during the depolarised phase of the action potential; when relaxation occurs, the concentration of Ca$^{2+}$  is reduced by uptake into the sarcoplasmic reticulum via Na$^{+}$/ Ca$^{2+}$  exchange. This is an electrogenic process: three Na$^{+}$ ions are exchanged for a single Ca$^{2+}$, thus making the potential difference more positive. This induces an inward current that can depolarise the cell which can lead to spontaneous action potentials firing \cite{Nattel2}.

Atrial cell fibrosis causes cells to be unresponsive. Current research suggests that this is due to excessive Ca$^{2+}$ loading, which occurs with tachycardia as many action potentials are enacted. Furthermore, this leads to electrical remodelling of the heart, making AF the more probable pathway of activation regardless of the mechanism for tachycardia \cite{Nattel2}. Due to fibrillation, the concentration of the intracellular Ca$^{2+}$ is altered, thus impairing effective contraction which may give rise to blood clots that can cause strokes. %REVISEEE

%MONODOMAIN AND BIDOMAIN STUFF??

\section{\textbf{\underline{Types of Models}}} %METHODS

Mechanisms thought to be responsible for active cardiac arrhythmias are either enhanced/abnormal pulse formation or re-entry. For the former, simulations have included modelling the ionic currents of action potentials such as the Fenton-Karma model \cite{Grandi} \cite{Fenton} and modelling the impulse of an action potential using generic reaction-diffusion equations, such as the FitzHugh-Nagumo Model \cite{FitzHugh}. Most of these models are based on the pioneering work of Hodgkin and Huxley who described the action potential of the giant squid axon, which bears similarities to cardiac cells \cite{Hodgkin}. %CHECKKKKKKKKKKKKKKKKKKK
 For all of these models bidomain or monodomain approaches can be used to model the electrical propagation in myocardial tissue. % intracellularly????
  The bidomain model takes into account the anisotropy of intracellular and extracellular spaces. The monodomain model is a reduction of the bidomain model by assuming anisotropy ratios of intracellular and extracellular domains are equal---this is the model most commonly used in simulations today.%ET CETERA?
% ,   is the Many types of models have been used to simulate fibrillatory behaviour in the heart, from, and signals for contraction, to cellular automata and

Re-entry---the mechanism by which a propagating impulse fails to die out after normal activation of the heart---has been studied in depth via various circus reentry formulations: the Ring Model, the Leading circle model, the Figure-of-Eight model and the Spiral Wave model \cite{Antze} \cite{Tusscher}. The Ring model models a wave of propagation that moves along one, unidirectional pathway of heart tissue and returns to the origin of excitation, thus simulating a central area of inexcitable tissue. This simple model implies that the conduction velocity and the refractory period are both factors: the length of the circuit (pathlength) must be greater or equal to the wavelength (the distance travelled by an impulse during the refractory period). The smaller the wavelength, the greater the number of reentry circuits that can be accommodated during the refractory period to promote fibrillatory behaviour. See figure 2(c).

The leading circle model found that there need not be a block of inexcitable tissue at the centre of circus movement for a reentry circuit to occur, using properly timed stimuli (see figure 2a). \cite{Allessie}. %Using a pathlength that is the size of the wavelength: decreased wavelength decreases minimum circuit size which increases number of circuits that can be accommodated.
 To rectify this reentry, it would be thought that the refractory period must increase to limit number of accommodated circuits, but incompatible observations have been found in antiarrhythmic drugs that block Na$^{+}$ channels, thus slowing down depolarisation. Such agents are effective in terminating AF, but according to leading circle theory should promote AF because they decrease conduction velocity and thereby decrease the wavelength. See figure 2(a) for diagram.

The Figure of Eight model has a centre where two concurrently propagating waves circulate in opposite directions that merge into a common, central pathway.

\begin{figure}
\caption[short title]{Diagram showing the leading circle and spiral wave models of reentry. PL = pathlength, RP = refractory period, CV = conduction velocity, WL = wavelength \cite{Nattel3}}
\centering
\includegraphics[width = 0.5\textwidth]{reeentry}
\end{figure}

The Spiral wave model---which consists of a single rotating wavefront around a core, simulating tachycardia---has been observed in systems that have sites that have excited, refractory and resting states in both homogeneous and heterogeneous media \cite{Greenburg} \cite{Bub}. This is due to anisotropies in conduction velocity in the cardiac tissue. See figure 2(b) and figure 3. % for comparison to a normal wavefront. 
 Experimental demonstration of spiral waves has been achieved by in research by Davidenko et al and the concept is now seen to be an underlying driver of fibrillation; the paradigm appears to more closely correspond with many clinical and experimental observations than other models \cite{Davidenko} \cite{Comtois}. In one such observation, the presence of rotors was seen in 96\% of patients with AF, in a 2012 study \cite{afstudy}. The curvature of the spiral wave is related to the speed of the wave, the more curvature the slower the speed of propagation. High curvature, and far-field instability, leads to spiral wave breakage into daughter wavelets. This agrees with
%popular classical modes of thought regarding the initiation of fibrillatory contraction, multiple circuit reentry, which in this case is due to spiral breakup, and 
recent opinions where single, small reentry circuits or local generators have been found to promote fibrillatory activity---the core of the spiral \cite{Skanes} \cite{Nattel2}. The tip of the spiral wave has been seen to drift around the core, thus moving the whole spiral wave. Such meandering can be a factor in promoting fibrillatory activity \cite{Tusscher}.
% Recent observations are in contradiction with the classical viewpoint,  . This is also concurrent 
 Extending the spiral wave model to 3D gives rise to scroll waves which propagate through multiple layers of cells and give rise to more complex wave dynamics and is therefore hard to implement computationally.  %It has been seen that the if a spiral wave is in homogenous media then it will always return to its resting state .

\begin{figure}
\caption[short title]{Diagram showing a spiral wave reentry circuit. (a), a normal wavefront in the heart. (b) spiral wave, which simulates tachycardia, as can be seen by multiple propagating wavefronts in a small area. (c), turbulent activity due to break up of spiral wave giving rise to fibrillation \cite{Alonso}}
\centering
\includegraphics[width = 0.5\textwidth]{spiralbreak2tachy}
\end{figure}

\section{\textbf{\underline{Cellular Automata Models and}\\ \underline{Further Developments}}}

Cellular automaton models have been formulated as there is no need to simulate the exact shape of the action potential, based on differential equations, and it is high speed in large scale simulations \cite{Felipe}. Only relative properties such as the excitation wavefront, refractory period and the morphology of the heart need be considered. Cells' states depend only on the states of their neighbours: they are either at rest, become excited, or are refractory. Refractory cells revert to rest after the refractory period has elapsed.

Mainly square lattices have been used for such cellular automata models. One problem is isotropy, especially when spiral waves are simulated. As seen in research by Pourhasanzande et al, Moore, rather than Von Neumann, neighbourhoods can be used with a Markus model reducing the square shape of the spiral wave by choosing an appropriate threshold value of neighbours \cite{Pour}. See figure 4.

\begin{figure}
\caption[short title]{Diagram showing spiral wave generated by using (a)---a Moore neighborhood---and (b)---a Von
Neumann neighborhood showing the reduction of flat edges using the Moore model with the Markus method
 \cite{Pour}}
\centering
\includegraphics[width = 0.5\textwidth]{flatedgered}
\end{figure}

Recent research has shown that rotating wavefronts can be initiated in regions where the proportion of transverse coupling between strands of tissue is reduced to a critical value, leading to spontaneous fibrillatory activity, in contrast to normal 'linear' wavefronts. This can be seen in the rectilinear cellular automata model developed by Christensen et al and reflects current opinion: the fibrosis of cells, which reduces the coupling of electrical signals to other cells, leads to AF\cite{Christensen}. This research also supports targeted ablation to stop AF. Other studies suggest similar mechanisms such as the coupling of fibroblasts to atrial myocytes heterogeneously gives rise to spiral wave breakups \cite{Ozawa}.

In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}. It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed. In addition, a multi-electrode system was created to study the localization of AF sources by simulated electrograms.

More cellular automata models have been researched by Siregar et al, where a 3D model of the heart, with simplified morphology is simulated and, in addition, a 2D CA model is compared qualitative model \cite{Siregar} \cite{Siregar2}. A hybrid I/O cellular automata approach has also been studied in research by Bartocci et al to describe cardiac tissue and improve the efficiency of the simulation process by parallelization \cite{Bartocci}.

 The impact of left and right atrial tissue electrical heterogeneities to promote AF has been studied in a biophysical model and it was seen that AF is more persistent in the right atria \cite{Luca}. Further development could be achieved by simulation from a cellular automata perspective. Modelling scroll waves and replicating the morphology of the heart with its heterogeneities more closely seem to be viable directions for further research. The effects of antiarrhythmic drugs on the heart during AF can be studied, along with the associated mechanism by which proarrhythmia occurs, and methods for its rectification. Developing ways to find CFAEs with novel electrogram methods via simulation could help targeted ablation, stopping AF more effectively and efficiently.

% Further developments could be obtained through this parallelization combined with 3D modelling, more accurately depicting the shape of the heart, making the code more efficient.

 %Many of these studies suggest targeted ablation can hinder or fully stop AF, therefore, electrogram-guided ablation to find CFAEs is a promising method it is concurrent with clinical observations. %electrogram-guided ablation to find CFAE.

%Most of these models only work one or two dimensionally, assuming that the heart is a 2D surface and not a 3D

%Cellular automaton models

%MAIN PAPER FROM KIM CHRISTIENSEN
%BIPOLAR ELECTROGRAM??
%METHODS TO POTENTIALLY STOP ATRIAL FIBRILLATION

%FUUUUUUUUUUUUURTHER METHODZZZZZZZZZZZZZZZZZZZZZZZZZZ

%\section{\textbf{\underline{Implications and Discussion}}}  %What our research project will be able to achieve, potentially

%\section{\textbf{\underline{Conclusion}}}

%In this literature review, possible underlying mechanisms of atrial fibrillation (AF) are described along with the relevant the electrophysiology of heart cells, specifically myocytes and  sinoatrial node cells. The integral role of these cells with regards to action potential propagation to contract the heart periodically is described. Models to simulate the action potentials of these cells to simulate reentry are seen to be a computationally costly way of modelling large scale cardiac tissue. Studied models of electrical wavefront propagation (reentry) that could cause AF, such as the ring, leading circle and spiral wave models are described; spiral wave reentry has been observed to be an underlying mechanism. Cellular automata---where cells' states change based on the states of their nearest neighbours---are viable models that reduce the amount of computation by not considering differential equations governing the action potentials. Therefore, reentry can be modelled on a large scale more easily. Moore neighbourhoods used with a Markus model can achieve smooth spiral waveforms that are isotropic and therefore model spiral wave reentry in a cellular automata models effectively. New directions of research can be taken to more accurately model AF such as: the morphology and heterogeneities of the heart, modelling 3D scroll wave dynamics and using hybrid I/O automata.

\section{\textbf{\underline{Bibliography}}}
\begin{thebibliography}{7}

%####################################################
\bibitem{Hart}
Hart, R. G. \& Halperin, J. L. (2001).
\emph{Atrial fibrillation and stroke: concepts and controversies.},
Stroke 32, 803–808,
2001

\bibitem{Nattel1}
S. Nattel, 
\emph{ Experimental evidence for proarrhythmic mechanisms of antiarrhythmic drugs.}
Cardiovasc Res. 1998 Mar;37(3):567-77.

\bibitem{Tsui}
Tsui, S. S. L. Grace, A. A. Ludman, P. F. Schofield, P. M. Page, A. J. P. Rowland, E.  and Large, S. R., 
\emph{ Maze 3 for Atrial Fibrillation: Two Cuts Too Few?.}
 Pacing and Clinical Electrophysiology,
17: 2163–2166. doi:10.1111/j.1540-8159.1994.tb03819.x
1994

\bibitem{Greek}
Demosthenes G Katritsis, Ioannis Pantos and Efstathios P Efstathopoulos, 
\emph{Catheter ablation of atrial fibrillation guided by electrogram fractionation and dominant frequency analysis},
Expert Review of Cardiovascular Therapy 
Volume 9, 2011 - Issue 5, 631-636, 2014

\bibitem{Vincent}
V. DeCaprio, P. Hurzeler snd S. Furman, 
\emph{A Comparison of Unipolar and Bipolar Electrograms for Cardiac Pacemaker Sensing}
Circulation, Vol 56, No 5, November 1977

\bibitem{Augustus}
Augustus O. Grant, MB, ChB, PhD
\emph{Basic Science for the Clinical Electrophysiologist: Cardiac Ion Channels}
Circulation: Arrhythmia and Electrophysiology.
2009; 2: 185-194
doi: 10.1161/CIRCEP.108.789081

\bibitem{Nattel3}
Nattel S, Carlsson L.
\emph{ Innovative approaches to anti-arrhythmic drug therapy.},
Nat Rev Drug Discov. 2006; 5: 1034–1049.

\bibitem{Mizzi}
\emph{Amiodarone supplants lidocaine in ACLS and CPR protocols}
A. Mizzi, T. Tran, D, Mangar, E. M. Camporesi
Anesthesiology Clinics 29(3):535-45 · September 2011

\bibitem{Nattel2}
S. Nattel, 
\emph{New Ideas about atrial fibrillation 50 years on},
 NATURE | VOL 415 | 10 JANUARY 2002 |

\bibitem{Grandi}
E. Grandi, S. V. Pandit, N. Voigt, M.D., A. J. Workman, D. Dobrev, J. Jalife, and D. M  Bers,
\emph{Human Atrial Action Potential and Ca$^{2+ }$ Model: Sinus Rhythm and Chronic Atrial Fibrillation},
Circ Res. 2011 October 14; 109(9): 1055–1066. doi:10.1161/CIRCRESAHA.111.253955.
November 2011

\bibitem{Fenton}
F. Fenton and A. Karma
\emph{Vortex dynamics in three-dimensional continuous myocardium with fiber
rotation: Filament instability and fibrillation},
1997

\bibitem{FitzHugh}
Richard FitzHugh, 
\emph{Impulses and Physiological States in Theoretical Models of Nerve Membrane},
Biophys J. 1961 Jul; 1(6): 445–466.

\bibitem{Hodgkin}
A. L. Hodgkin and A. F. Huxley, 
\emph{A quantitative description of membrane current and its application to conduction and excitation in nerve},
J. Physiol. (I952) I I7, 500-544

\bibitem{Antze}
 C. Antzelevitch
\emph{Basic mechanisms of re-entrant arrhythmias}
Curr Opin Cardiol. 2001 Jan;16(1):1-7.

\bibitem{Tusscher}
\emph{Spiral wave dynamics and ventricular arrhythmias},
 K. H. W. Johanna ten Tusscher,
 November 2004.

\bibitem{Allessie}
Allessie MA, Bonke FIM, Schopman JG,
\emph{ Circus movement in rabbit atrial muscle as a mechanism of tachycardia. III. The “leading circle” concept: a new model of circus movement in cardiac tissue without the involvement of an anatomical obstacle.},
 Circ Res 1977, 41:9–18.

\bibitem{Greenburg}
J. M. Greenberg and S. P. Hastings, 
\emph{Spatial Patterns for Discrete Models of Diffusion in Excitable Media}, 
SIAM J. APPL. MATH.
Vol. 34, No. 3, May 1978

\bibitem{Bub}
G. Bub, A. Shrier, and L. Glass,
\emph{Spiral Wave Generation in Heterogeneous Excitable Media},
 Phys. Rev. Lett.88, 058101 (2002)

\bibitem{Davidenko}
Davidenko JM, Kent PF, Chialvo DR, Michaels DC, Jalife J., 
\emph{Sustained vortex-like waves in normal isolated ventricular muscle.},
 Proc Natl Acad Sci USA. 1990;87:8785–8789

\bibitem{Comtois}
P. Comtois, J. Kneller, S. Nattel,
\emph{Of circles and spirals: Bridging the gap between the leading circle and spiral wave concepts of cardiac reentry}
Europace. 2005 Sep;7 Suppl 2:10-20.
DOI: http://dx.doi.org/10.1016/j.eupc.2005.05.011

\bibitem{afstudy}
S. M. Narayan, D. E. Krummen, and W.-J. RAPPEL, 
\emph{Clinical mapping approach to diagnose
electrical rotors and focal impulse sources for human atrial fibrillation},
 Journal of
cardiovascular electrophysiology, vol. 23, no. 5, pp. 447–454, 2012.

\bibitem{Skanes}
Mandapati, R., Skanes, A., Chen, J., Berenfeld, O. and  Jalife, J. 
\emph{Stable microreentrant sources as a mechanism of atrial fibrillation in the isolated sheep heart.}
Circulation 101, 194–199 (2000).

\bibitem{Alonso}
S. Alonso, M. Bar and B. Echebarria
\emph{Nonlinear physics of electrical wave
propagation in the heart: a review},
Rep. Prog. Phys. 79 (2016) 096601, 
doi:10.1088/0034-4885/79/9/096601

\bibitem{Christensen}
Kim Christensen, Kishan A. Manani, and Nicholas S. Peters
\emph{Simple Model for Identifying Critical Regions in Atrial Fibrillation}
Phys. Rev. Lett. 114, 028104 – Published 16 January 2015

\bibitem{Ozawa}
Ashihara T, Haraguchi R, Nakazawa K, Namba T, Ikeda T, Nakazawa Y, Ozawa T, Ito M, Horie M, Trayanova NA., 
\emph{The role of fibroblasts in complex fractionated electrograms during persistent/permanent atrial fibrillation: implications for electrogram-based catheter ablation.}
Circ Res. 2012 Jan 20;110(2):275-84. doi: 10.1161/CIRCRESAHA.111.255026

\bibitem{Felipe}
F.A. Atienza, J. R. Carrion, A. G. Alberola, J. L. R. Alvarez, J. J. S. Munoz, J. M. Sanchez, M. V. Chavarri,
\emph{A Probabilistic Model of Cardiac Electrical Activity Based on a Cellular Automata System},
Revista Española de Cardiología (English Edition), Volume 58, Issue 1, January 2005, Pages 41-47,
doi:10.1016/S1885-5857(06)60233-8

\bibitem{Pour}
F. Pourhasanzande, S. H. Sabzpoushan
\emph{A new Cellular Automata Model of Cardiac
Action Potential Propagation based on
Summation of Excited Neighbors}, 
International Journal of Medical, Health, Biomedical, Bioengineering and Pharmaceutical Engineering Vol:4, No:8, 2010

\bibitem{Fedotov}
N. M. Fedotov,  A. I. Oferkin, and S. V. Zhary,
\emph{Modeling Sources of Atrial Fibrillation on a Triangulated Sphere}
Biomedical Engineering, Vol. 49, No. 2, July, 2015, pp. 112115.

\bibitem{Siregar}
P. Siregar, J. P. Sinteff, N. Julen, P. Le Beux,
\emph{An Interactive 3D Anisotropic Cellular Automata Model of the Heart}
Computers and Biomedical Research
Volume 31, Issue 5, October 1998, Pages 323-347, 
doi:10.1006/cbmr.1998.1485

\bibitem{Siregar2}
P. Siregar, J.P. Sinteff, M. Chahine, P. Lebeux
\emph{A Cellular Automata Model of the Heart and Its Coupling with a Qualitative Model}, 
Computers and Biomedical Research
Volume 29, Issue 3, June 1996, Pages 222-246,
doi:10.1006/cbmr.1996.0017

\bibitem{Bartocci}
E. Bartocci, F. Corradini, M.R. Di Berardini, E. Entcheva and S.A. Smolka, R. Grosu, 
\emph{Modeling and simulation of cardiac tissue using hybrid I/O automata}
Theoretical Computer Science
Volume 410, Issues 33–34, 21 August 2009, Pages 3149-3165,
doi:10.1016/j.tcs.2009.02.042

\bibitem{Luca}
A. Luca, V. Jaquemet, N. Virag, J. M. Vesin
\emph{Influence of Right and Left Atrial Tissue Heterogeneity on
Atrial Fibrillation Perpetuation }
Computing in Cardiology 2015; 42:449-452.

%imref https://o.quizlet.com/i/1UZTM_D5PFefI8E5JOk8Ig_m.jpg

\end{thebibliography}

\end{document}

### 2017-02-20 Literature Review/Literature Review, Complexity Science approach to Atrial Fibrillation.tex
POSTERTINGZZZZZZZZZZZZZ

--- 2017-02-20 Literature Review/DraftReport.tex
@@
 \title {\textbf{Literature Review: Complexity science approach to Atrial Fibrillation}}
-\author{Tigany Zarrouk\\ CID: 00823932\\\textbf{Supervisor}: Kim Christensen \\\textbf{Word Count}: 2488.}
-
+\author{Tigany Zarrouk\\ CID: 00823932}
 \date{}
@@
 \begin{abstract}
-
-Modelling Atrial Fibrillation (AF) to ascertain its causes accurately, allowing for effective treatment, has been extensively studied over the past few decades.
-It is the largest cause of strokes linked with cardiac arrhythmia in aging populations and is due to impaired contractility of the atria in the heart, leading to blood clot formation.
-Action potential propagation, mediated by ionic species, are seen to be irregular in AF, which causes some atrial myocytes to contract at 300+ times per minute, much faster than healthy hearts which contract at 60-100 times per minute.
-This beating is regulated by the sinoatrial node which initiates action potentials that propagate through myocytes, causing contractions.
-Altered refractory period and conduction speed of signals has been seen to promote AF in various models and is caused physically by the fibrosis of cells, reducing the coupling of electrical signals transversally to some myocytes; giving rise to reentry circuits.
-The most prominent reentry model, the spiral wave model, has been seen to more closely follow clinical and experimental observations than other models of wavefront action.
-Cellular automata modelling paves the way for large scale simulations of tissue in the heart without the need to solve differential equations, which can be computationally costly.
-Moore neighbourhoods are seen to be more effective at producing isotropy of spiral waves, reducing flat edges, in rectilinear cellular automata models.
-Modelling scroll waves, using hybrid cellular automata models, using novel methods to find rotors and replicating the morphology of the heart with its heterogeneities more closely seem to be viable directions for further research.
+In this literature review, the electrophysiology of heart cells, specifically heart muscle cells (myocytes) and sinoatrial node cells, are elucidated and their relation to action potential propagation to contract the heart is described.
+Methods to simulate the action potentials of these cells are mentioned and are seen to be a computationally costly way of modelling large scale heart tissue.
+Studied models of electrical wavefront propagation that could cause AF, such as the ring model, leading circle model and spiral wave model are described; spiral wave reentry has been observed to be an underlying mechanism.
+Cellular automata---where cells' states change based on the states of their nearest neighbours---are viable models that reduce the amount of computation by not considering differential equations relating to the action potentials.
+Therefore, reentry can be modelled on a large scale more easily.
+Moore neighbourhoods used with a Markus model can achieve smooth spiral waveforms that are isotropic and therefore model spiral wave reentry in a cellular automata models effectively.
+New directions of research can be taken to more accurately model AF such as: the morphology of the heart, modelling 3D scroll wave dynamics and using hybrid I/O automata.
 
@@
 One such condition that has eluded explanation and effective treatment is Atrial Fibrillation (AF), which is the leading cause of Ischaemic stroke in people over 75---seemingly due to incomplete contraction which leads to the formation of blood clots that can travel to the brain \cite{Hart}.
-It has been known that there are cardiovascular diseases that are predisposed to initiate AF, such as Congestive Heart Failure, Pericarditis and Hypertension through a combination of atrial pressure and dilation, but the precise mechanics are unknown.
+It has been known that there are cardiovascular diseases that are predisposed to initiate AF, such as Congestive Heart Failure, Pericarditis, Coronary Heart Disease and Hypertension through a combination of atrial pressure and dilation, but the precise mechanics are unknown.
 AF and VF (ventricular fibrillation) are thought to be caused by the breakup of waves of electrical impulses, that induce contraction, into interfering wavelets that produce chaotic behaviour.
@@
 Healthy hearts have a regular, sinus  rhythm which is moderated by the sinoatrial node located in the right atrium of the heart.
-Fibrillatory activity in the atrium is characterised by a hastened activation of cells: atrial cells activate 300-600 times a minute in contrast to the expected sinus rhythm of 60-100 beats per minute.% Ectopic firing has been observed in mainly in the left atria.
+Fibrillatory activity in the atrium is characterised by a hastened activation of cells: atrial cells activate 300-600 times a minute in contrast to the expected sinus rhythm of 60 to 100 beats per minute.% Ectopic firing has been observed in mainly in the left atria.
 
-Current treatment involves introducing antiarrhythmic drugs of which side effects include proarrhythmia, which can cause sudden death \cite{Nattel1}.
-Surgical methods primarily involve the Maze procedure, by which a maze-like pattern is cut into the atria such that there is only one path that the electrical impulse can take from the sinoatrial node to the atrioventricular node.
+Current treatment involves introducing antiarrhythmic drugs of which side effects include proarrhythmia, a form of cardiac arrhythmia, which can cause sudden death \cite{Nattel1}.
+Surgical methods primarily involves the Maze procedure, by which a children's maze-like pattern is cut into the atria such that there is only one path that the electrical impulse can take from the sinoatrial node to the atrioventricular node.
 The treatment is usually successful but still may lead to fibrillation \cite{Tsui}.
@@
 Catheter ablation has had suboptimal success rates.
-Electrogram-guided ablation is a burgeoning area of research.
-Complex-fractionated atrial electrograms (CFAEs) are proposed as sites for ablation as they seem to be critical for AF \cite{Greek}.
+Electrogram-guided ablation is a burgeoning area of research: complex fractionated atrial electrograms are proposed as sites for ablation; they seem to be critical for AF \cite{Greek}.
 % These sites seem to be mainly focussed around the pulmonary veins.
@@
 Once depolarised enough, after a small efflux of K$^{+}$ and  Cl$^{-}$ ions (phase 1), Ca$^{2+}$ ion voltage-gated channels open, which induces Ca$^{2+}$ release from intracellular stores.
-This gives a plateau of potential difference (phase 2) where an influx of  K$^{+}$ ions sustains the voltage against the efflux of  Ca$^{2+}$ ions.
+This gives a plateau of potential difference (phase 2) where an influx of  K$^{+}$ ions sustains the voltage plateau against the efflux of  Ca$^{2+}$ ions.
 The free calcium ions cause myocardial contraction.
 After a this long plateau phase,  K$^{+}$ ion channels reopen to repolarise the cell (phase 3) to get ready for another contraction, reverting to phase 4.
-The period where a cell cannot respond to stimuli to induce contraction is the refractory period.
+The period where a cell cannot contract and undergoes repolarisation ready for contraction is the refractory period.
 
@@
 In the sinoatrial node, the depolarising current is primarily fulfilled by Ca$^{2+}$ ion movement, which is slow in comparison to the fast Na$^{+}$ currents of the myocardial cells.
-SA node action potentials have only three phases.
+SA node action potentials have only three of the phases found in the myocardial action potential.
 Phase 4 is the spontaneous depolarization (pacemaker potential) that triggers the action potential.
@@
 This is followed by phase 3, repolarization.
-Once the cell is completely repolarized, the cycle is spontaneously repeated.
+Once the cell is completely repolarized at about -60 mV, the cycle is spontaneously repeated.
 
@@
 \caption[short title]{Diagram of the action potential in a myocardial cell (A) and in sinoatrial cell (B).
-Numbers indicate phases of the action potential.
+Numbers on the diagram indicate phases of the action potential.
 ERP = Effective refractory period \cite{Mizzi}}
@@
 
-Mechanisms thought to be responsible for active cardiac arrhythmias are either enhanced/abnormal pulse formation or re-entry.
-For the former, simulations have included modelling the ionic currents of action potentials such as the Fenton-Karma model \cite{Grandi} \cite{Fenton} and modelling the impulse of an action potential using generic reaction-diffusion equations, such as the FitzHugh-Nagumo Model \cite{FitzHugh}.
+Mechanisms thought to be responsible for active cardia arrhythmias are either enhanced/abnormal pulse formation or re-entry.
+For the former, simulations have included modelling the ionic currents to propagate action potentials such as the Fenton-Karma model \cite{Grandi} \cite{Fenton} and modelling the impulse of an action potential using generic reaction-diffusion equations, such as the FitzHugh-Nagumo Model \cite{FitzHugh}.
 Most of these models are based on the pioneering work of Hodgkin and Huxley who described the action potential of the giant squid axon, which bears similarities to cardiac cells \cite{Hodgkin}.
@@
   The bidomain model takes into account the anisotropy of intracellular and extracellular spaces.
-The monodomain model is a reduction of the bidomain model by assuming anisotropy ratios of intracellular and extracellular domains are equal---this is the model most commonly used in simulations today.%ET CETERA?
+The monodomain model is a reduction of the bidomain model by assuming anisotropy ratios of intracellular and extracellular domains are equal---this is the most commonly used model.%ET CETERA?
 % ,   is the Many types of models have been used to simulate fibrillatory behaviour in the heart, from, and signals for contraction, to cellular automata and
@@
 Re-entry---the mechanism by which a propagating impulse fails to die out after normal activation of the heart---has been studied in depth via various circus reentry formulations: the Ring Model, the Leading circle model, the Figure-of-Eight model and the Spiral Wave model \cite{Antze} \cite{Tusscher}.
-The Ring model models a wave of propagation that moves along one, unidirectional pathway of heart tissue and returns to the origin of excitation, thus simulating a central area of inexcitable tissue.
-This simple model implies that the conduction velocity and the refractory period are both factors: the length of the circuit (pathlength) must be greater or equal to the wavelength (the distance travelled by an impulse during the refractory period).
+The Ring model models a wave of propagation that moves along one, unidirectional pathway of heart tissue and returns to the origin of excitation, thus simulating a central anatomical obstacle (inexcitable tissue) in the centre.
+This implies that the conduction velocity and the refractory period are both factors: the length of the circuit (pathlength) must be greater or equal to the wavelength (the distance travelled by an impulse during the refractory period---the product of conduction velocity and refractory period).
 The smaller the wavelength, the greater the number of reentry circuits that can be accommodated during the refractory period to promote fibrillatory behaviour.
@@
 The Spiral wave model---which consists of a single rotating wavefront around a core, simulating tachycardia---has been observed in systems that have sites that have excited, refractory and resting states in both homogeneous and heterogeneous media \cite{Greenburg} \cite{Bub}.
-This is due to anisotropies in conduction velocity in the cardiac tissue.
-See figure 2(b) and figure 3.
+See figure 2(b) and and figure 3.
 % for comparison to a normal wavefront.
  Experimental demonstration of spiral waves has been achieved by in research by Davidenko et al and the concept is now seen to be an underlying driver of fibrillation; the paradigm appears to more closely correspond with many clinical and experimental observations than other models \cite{Davidenko} \cite{Comtois}.
-In one such observation, the presence of rotors was seen in 96\% of patients with AF, in a 2012 study \cite{afstudy}.
-The curvature of the spiral wave is related to the speed of the wave, the more curvature the slower the speed of propagation.
+The curvature of the wave is related to the speed of the wave, the more curvature the slower the speed of propagation.
 High curvature, and  far-field instability, leads to spiral wave breakage into daughter wavelets.
-This agrees with
-%popular classical modes of thought regarding the initiation of fibrillatory contraction, multiple circuit reentry, which in this case is due to spiral breakup, and 
-recent opinions where single, small reentry circuits or local generators have been found to promote fibrillatory activity---the core of the spiral \cite{Skanes} \cite{Nattel2}.
+This agrees with popular classical modes of thought regarding the initiation of fibrillatory contraction, multiple circuit reentry, and recent opinions where single, small reentry circuits or local generators have been found to promote fibrillatory activity---the core of the spiral \cite{Skanes} \cite{Nattel2}.
 The tip of the spiral wave has been seen to drift around the core, thus moving the whole spiral wave.
-Such meandering can be a factor in promoting fibrillatory activity \cite{Tusscher}.
+This mechanic can be a factor in promoting fibrillatory activity.
 % Recent observations are in contradiction with the classical viewpoint,  . This is also concurrent 
- Extending the spiral wave model to 3D gives rise to scroll waves which propagate through multiple layers of cells and give rise to more complex wave dynamics and is therefore hard to implement computationally.
+ Extending the spiral wave model to 3D gives rise to scroll waves which propagate through multiple layers of cells and give rise to more complex dynamics and is therefore hard to implement computationally.
 %It has been seen that the if a spiral wave is in homogenous media then it will always return to its resting state .
@@
 
-\section{\textbf{\underline{Cellular Automata Models and}\\ \underline{Further Developments}}}
+\section{\textbf{\underline{Cellular Automata Models}}}
+
+Recent research has shown that rotating wavefronts can be initiated in regions where the proportion of transverse coupling between strands of tissue is reduced, leading to fibrillatory activity, in contrast to normal 'linear' wavefronts.
+This can be seen in the cellular automata model developed by Christensen et al.
+and reflects current opinion: the fibrosis of cells, which reduces the coupling of electrical signals to other cells, leads to AF\cite{Christensen}.
+Other studies suggest that the coupling of fibroblasts to atrial myocytes heterogeneously gives rise to spiral wave breakups \cite{Ozawa}.
 
@@
 Cells' states depend only on the states of their neighbours: they are either at rest, become excited, or are refractory.
-Refractory cells revert to rest after the refractory period has elapsed.
+This is seen in research by Fedotov, using a cellular automaton model on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
+It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed.
+In addition, a multi-electrode system was created to study the localization of AF sources.
 
@@
 
-Recent research has shown that rotating wavefronts can be initiated in regions where the proportion of transverse coupling between strands of tissue is reduced to a critical value, leading to spontaneous fibrillatory activity, in contrast to normal 'linear' wavefronts.
-This can be seen in the rectilinear cellular automata model developed by Christensen et al and reflects current opinion: the fibrosis of cells, which reduces the coupling of electrical signals to other cells, leads to AF\cite{Christensen}.
-This research also supports targeted ablation to stop AF.
-Other studies suggest similar mechanisms such as the coupling of fibroblasts to atrial myocytes heterogeneously gives rise to spiral wave breakups \cite{Ozawa}.
-
-In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
-It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed.
-In addition, a multi-electrode system was created to study the localization of AF sources by simulated electrograms.
-
 More cellular automata models have been researched by Siregar et al, where a 3D model of the heart, with simplified morphology is simulated and, in addition, a 2D CA model is compared qualitative model \cite{Siregar} \cite{Siregar2}.
-A hybrid I/O cellular automata approach has also been studied in research by Bartocci et al to describe cardiac tissue and improve the efficiency of the simulation process by parallelization \cite{Bartocci}.
-
- The impact of left and right atrial tissue electrical heterogeneities to promote AF has been studied in a biophysical model and it was seen that AF is more persistent in the right atria \cite{Luca}.
-Further development could be achieved by simulation from a cellular automata perspective.
-Modelling scroll waves and replicating the morphology of the heart with its heterogeneities more closely seem to be viable directions for further research.
-The effects of antiarrhythmic drugs on the heart during AF can be studied, along with the associated mechanism by which proarrhythmia occurs, and methods for its rectification.
-Developing ways to find CFAEs with novel electrogram methods via simulation could help targeted ablation, stopping AF more effectively and efficiently.
-
-% Further developments could be obtained through this parallelization combined with 3D modelling, more accurately depicting the shape of the heart, making the code more efficient.
-
- %Many of these studies suggest targeted ablation can hinder or fully stop AF, therefore, electrogram-guided ablation to find CFAEs is a promising method it is concurrent with clinical observations.
-%electrogram-guided ablation to find CFAE.
+A hybrid I/O cellular automata approach has also been studied in research by Bartocci et al to describe cardiac tissue and improve the efficiency of the simulation process \cite{Bartocci}.
 
@@
 
-%\section{\textbf{\underline{Conclusion}}}
+\section{\textbf{\underline{Conclusion}}}
 
-%In this literature review, possible underlying mechanisms of atrial fibrillation (AF) are described along with the relevant the electrophysiology of heart cells, specifically myocytes and  sinoatrial node cells.
-The integral role of these cells with regards to action potential propagation to contract the heart periodically is described.
-Models to simulate the action potentials of these cells to simulate reentry are seen to be a computationally costly way of modelling large scale cardiac tissue.
-Studied models of electrical wavefront propagation (reentry) that could cause AF, such as the ring, leading circle and spiral wave models are described; spiral wave reentry has been observed to be an underlying mechanism.
-Cellular automata---where cells' states change based on the states of their nearest neighbours---are viable models that reduce the amount of computation by not considering differential equations governing the action potentials.
-Therefore, reentry can be modelled on a large scale more easily.
-Moore neighbourhoods used with a Markus model can achieve smooth spiral waveforms that are isotropic and therefore model spiral wave reentry in a cellular automata models effectively.
-New directions of research can be taken to more accurately model AF such as: the morphology and heterogeneities of the heart, modelling 3D scroll wave dynamics and using hybrid I/O automata.
+Modelling Atrial Fibrillation (AF) to ascertain its causes accurately, allowing for effective treatment, has been extensively studied over the past few decades.
+It is the largest cause of strokes linked with cardiac arrhythmia in aging populations; is due to impaired contractility of the atria in the heart, leading to blood clot formation.
+Irregular action potentials, mediated by ionic species, observed in AF causes some atrial heart muscle cells, myocytes, to contract at 300+ times per minute, much faster than healthy hearts which contract at 60 to 100 times per minute.
+This beating is regulated by the sinoatrial node which propagates periodic signals through myocytes, causing contractions.
+Altered refractory period and conduction speed of signals has been seen to promote AF in various models and is caused physically by the fibrosis of cells, reducing the coupling of electrical signals to some myocytes; giving rise to reentry circuits.
+The most prominent reentry model, the spiral wave model, has been seen to more closely follow clinical and experimental observations than other models of wavefront action.
+Cellular automata modelling paves the way for large scale simulations of tissue in the heart without the need to solve differential equations, that govern the action potentials, which can be computationally costly.
+Moore neighbourhoods are seen to be more effective at producing isotropy of spiral waves, reducing flat edges, in rectilinear cellular automata models.
+Modelling scroll waves, using hybrid cellular automata models and replicating the morphology of the heart more closely seem to be viable directions for further research.
 
@@
 
-\bibitem{afstudy}
-S. M. Narayan, D. E. Krummen, and W.-J. RAPPEL, 
-\emph{Clinical mapping approach to diagnose
-electrical rotors and focal impulse sources for human atrial fibrillation},
- Journal of
-cardiovascular electrophysiology, vol. 23, no. 5, pp. 447–454, 2012.
+\bibitem{Christensen}
+Kim Christensen, Kishan A. Manani, and Nicholas S. Peters
+\emph{Simple Model for Identifying Critical Regions in Atrial Fibrillation}
+Phys. Rev. Lett. 114, 028104 – Published 16 January 2015
 
@@
 doi:10.1088/0034-4885/79/9/096601
-
-\bibitem{Christensen}
-Kim Christensen, Kishan A. Manani, and Nicholas S. Peters
-\emph{Simple Model for Identifying Critical Regions in Atrial Fibrillation}
-Phys. Rev. Lett. 114, 028104 – Published 16 January 2015
 
@@
 
-\bibitem{Luca}
-A. Luca, V. Jaquemet, N. Virag, J. M. Vesin
-\emph{Influence of Right and Left Atrial Tissue Heterogeneity on
-Atrial Fibrillation Perpetuation }
-Computing in Cardiology 2015; 42:449-452.
-
 %imref https://o.quizlet.com/i/1UZTM_D5PFefI8E5JOk8Ig_m.jpg

### 2017-02-20 Literature Review/MSCI POSTER draft.tex
POSTERTINGZZZZZZZZZZZZZ

--- 2017-02-20 Literature Review/DraftReport.tex
@@
-\documentclass[twocolumn]{article}
+\documentclass[twocolumn, a1paper]{article}
+ %\usepackage[scale=0.78,size=a1]{beamerposter}
+%\setlength{\paperheight}{33.1in}
+%\setlength{\paperwidth}{23.4in}
 \usepackage{cite}
@@
 
-\title {\textbf{Literature Review: Complexity science approach to Atrial Fibrillation}}
-\author{Tigany Zarrouk\\ CID: 00823932\\\textbf{Supervisor}: Kim Christensen \\\textbf{Word Count}: 2488.}
+\title {\textbf{ Complexity science approach to Atrial Fibrillation}}
+\author{Tigany Zarrouk\\ CID: 00823932\\\textbf{Supervisor}: Kim Christensen \\ Condensed Matter Theory Group---Imperial College London}
 
@@
 
-The leading circle model found that there need not be a block of inexcitable tissue at the centre of circus movement for a reentry circuit to occur, using properly timed stimuli (see figure 2a).
-\cite{Allessie}.
-%Using a pathlength that is the size of the wavelength: decreased wavelength decreases minimum circuit size which increases number of circuits that can be accommodated.
- To rectify this reentry, it would be thought that the refractory period must increase to limit number of accommodated circuits, but incompatible observations have been found in antiarrhythmic drugs that block Na$^{+}$ channels, thus slowing down depolarisation.
-Such agents are effective in terminating AF, but according to leading circle theory should promote AF because they decrease conduction velocity and thereby decrease the wavelength.
-See figure 2(a) for diagram.
-
-The Figure of Eight model has a centre where two concurrently propagating waves circulate in opposite directions that merge into a common, central pathway.
-
-\begin{figure}
-\caption[short title]{Diagram showing the leading circle and spiral wave models of reentry. PL = pathlength, RP = refractory period, CV = conduction velocity, WL = wavelength \cite{Nattel3}}
-\centering
-\includegraphics[width = 0.5\textwidth]{reeentry}
-\end{figure}
-
 The Spiral wave model---which consists of a single rotating wavefront around a core, simulating tachycardia---has been observed in systems that have sites that have excited, refractory and resting states in both homogeneous and heterogeneous media \cite{Greenburg} \cite{Bub}.
@@
 \centering
-\includegraphics[width = 0.5\textwidth]{spiralbreak2tachy}
+\includegraphics[width = 0.25\textwidth]{spiralbreak2tachy}
 \end{figure}
@@
 \section{\textbf{\underline{Cellular Automata Models and}\\ \underline{Further Developments}}}
-
-Cellular automaton models have been formulated as there is no need to simulate the exact shape of the action potential, based on differential equations, and it is high speed in large scale simulations \cite{Felipe}.
-Only relative properties such as the excitation wavefront, refractory period and the morphology of the heart need be considered.
-Cells' states depend only on the states of their neighbours: they are either at rest, become excited, or are refractory.
-Refractory cells revert to rest after the refractory period has elapsed.
-
-Mainly square lattices have been used for such cellular automata models.
-One problem is isotropy, especially when spiral waves are simulated.
-As seen in research by Pourhasanzande et al, Moore, rather than Von Neumann, neighbourhoods can be used with a Markus model reducing the square shape of the spiral wave by choosing an appropriate threshold value of neighbours \cite{Pour}.
-See figure 4.
-
-\begin{figure}
-\caption[short title]{Diagram showing spiral wave generated by using (a)---a Moore neighborhood---and (b)---a Von
-Neumann neighborhood showing the reduction of flat edges using the Moore model with the Markus method
- \cite{Pour}}
-\centering
-\includegraphics[width = 0.5\textwidth]{flatedgered}
-\end{figure}
 

### 2017-02-20 Literature Review/PosterTemplateDraftAF.tex
POSTERTINGZZZZZZZZZZZZZ

% --------------------------------------------------------------------------- %
% Poster for the ECCS 2011 Conference about Elementary Dynamic Networks.      %
% --------------------------------------------------------------------------- %
% Created with Brian Amberg's LaTeX Poster Template. Please refer for the     %
% attached README.md file for the details how to compile with `pdflatex`.     %
% --------------------------------------------------------------------------- %
% $LastChangedDate:: 2011-09-11 10:57:12 +0200 (V, 11 szept. 2011)          $ %
% $LastChangedRevision:: 128                                                $ %
% $LastChangedBy:: rlegendi                                                 $ %
% $Id:: poster.tex 128 2011-09-11 08:57:12Z rlegendi                        $ %
% --------------------------------------------------------------------------- %
\documentclass[a0paper,portrait]{baposter}

\usepackage{relsize}		% For \smaller
\usepackage{url}			% For \url
% \usepackage{epstopdf}	% Included EPS files automatically converted to PDF to include with pdflatex

%%% Global Settings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\graphicspath{{pix/}}	% Root directory of the pictures 
% \tracingstats=2			% Enabled LaTeX logging with conditionals

%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{bordercol}{RGB}{40,40,40}
\definecolor{headercol1}{RGB}{186,215,230}
\definecolor{headercol2}{RGB}{80,80,80}
\definecolor{headerfontcol}{RGB}{0,0,0}
\definecolor{boxcolor}{RGB}{186,215,230}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Utility functions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Save space in lists. Use this after the opening of the list %%%%%%%%%%%%%%%%
\newcommand{\compresslist}{
	\setlength{\itemsep}{1pt}
	\setlength{\parskip}{0pt}
	\setlength{\parsep}{0pt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Document Start %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\typeout{Poster rendering started}

%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\background{
	\begin{tikzpicture}[remember picture,overlay]%
	\draw (current page.north west)+(-2em,2em) node[anchor=north west]
	{\includegraphics[height=1.1\textheight]{background}};
	\end{tikzpicture}
}

%%% General Poster Settings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%% Eye Catcher, Title, Authors and University Images %%%%%%%%%%%%%%%%%%%%%%
\begin{poster}{
	grid=false,
	% Option is left on true though the eyecatcher is not used. The reason is
	% that we have a bit nicer looking title and author formatting in the headercol
	% this way
	%eyecatcher=false, 
	borderColor=bordercol,
	headerColorOne=headercol1,
	headerColorTwo=headercol2,
	headerFontColor=headerfontcol,
	% Only simple background color used, no shading, so boxColorTwo isn't necessary
	boxColorOne=boxcolor,
	headershape=roundedright,
	headerfont=\Large\sf\bf,
	textborder=rectangle,
	background=user,
	headerborder=open,
  boxshade=plain
}
%%% Eye Cacther %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
	Eye Catcher, empty if option eyecatcher=false - unused
}
%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\sf\bf
	Complexity Science Approach To Atrial Fibrillation
}
%%% Authors %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
	\vspace{1em} Tigany Zarrouk \& Mattia Gaggi\\
	%{\smaller legendi@inf.elte.hu, lgulyas@colbud.hu, gkampis@colbud.hu}
}
%%% Logo %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%{
% The logos are compressed a bit into a simple box to make them smaller on the result
% (Wasn't able to find any bigger of them.)
%\setlength\fboxsep{0pt}
%\setlength\fboxrule{0.5pt}
%	\fbox{
%		\begin{minipage}{14em}
%			\includegraphics[width=10em,height=4em]{colbud_logo}
%			\includegraphics[width=4em,height=4em]{elte_logo} \\
%			\includegraphics[width=10em,height=4em]{dynanets_logo}%
%			\includegraphics[width=4em,height=4em]{aitia_logo}
%		\end{minipage}
%	}
%}

\headerbox{Problem}{name=problem,column=0,row=0}{
Modelling Atrial Fibrillation (AF) to ascertain its causes accurately, allowing for effective treatment, has been extensively studied over the past few decades. It is the largest cause of strokes linked with cardiac arrhythmia in aging populations and is due to impaired contractility of the atria in the heart, leading to blood clot formation. Action potential propagation, mediated by ionic species, are seen to be irregular in AF, which causes some atrial myocytes to contract at 300+ times per minute, much faster than healthy hearts which contract at 60-100 times per minute.

This beating is regulated by the sinoatrial node which initiates action potentials that propagate through myocytes, causing contractions. Altered refractory period and conduction speed of signals has been seen to promote AF in various models and is caused physically by the fibrosis of cells, reducing the coupling of electrical signals transversally to some myocytes; giving rise to reentry circuits. The most prominent reentry model, the spiral wave model, has been seen to more closely follow clinical and experimental observations than other models of wavefront action. Cellular automata modelling paves the way for large scale simulations of tissue in the heart without the need to solve differential equations, which can be computationally costly.
%\begin{figure}
%\caption[short title]{Diagram showing a spiral wave reentry circuit. (a), a normal wavefront in the heart. (b) spiral wave, which simulates tachycardia, as can be seen by multiple propagating wavefronts in a small area. (c), turbulent activity due to break up of spiral wave giving rise to fibrillation \cite{Alonso}}
%\centering
%\includegraphics[width = 0.5\textwidth]{spiralbreak2tachy}
%\end{figure}
}

\headerbox{Basic Concepts}{name=definitions,column=0,below=problem}{
Originally, a model devised by Christensen \emph(et al.), modelled atrial tissue of the heart using a 200x200 lattice of cells, whereby cells along rows were fully connected to each other---replicating myocardial fibres---and cells in columns were connected to each other with a probability $\nu$  Furthermore, a proportion of the cells were deigned to be dysfunctional with a probability $\delta$ which had a probability $\epsilon$ of not being excited by an adjacent excited cell connected to it: this simulated the effects of fibrosis of cells in the heart. Tachycardia was simulated by using a period of 220 timesteps for a heartbeat and cells, once excited, had a refractory period of 50 timesteps. As can be seen in the figures below, chaotic, fibrillatory activity arised spontaneously from the model. 
}

\headerbox{Models}{name=models,column=0,below=definitions}{
\textbf{ER1} $G_0$ is a random graph. Add each non-existing edge with $p_A$, delete each existing edge with $p_D$ probability. \\
\textbf{ER2} $G_0$ is a random graph. Add $k_A$ uniformly selected random new edges and delete $k_D$ existing edges. \\
\textbf{ER3} $G_0$ is a random graph. Rewire $k_{RW}$ edges. \\
%\textbf{SPA} (\emph{Snapshot preferential}) $G_0$ is a scale free network. Add $k_A$ edges from a random node with preferential attachment based on the snapshot network. Delete $k_D$ existing edges. \\
%\textbf{CPA} (\emph{Cumulative preferential}) $G_0$ is a scale free network. Add $k_A$ edges from a random node with preferential attachment based on the cumulative network. Delete $k_D$ existing edges.
}

\headerbox{References}{name=references,column=0,below=models}{
\smaller													% Make the whole text smaller
\vspace{-0.4em} 										% Save some space at the beginning
\bibliographystyle{plain}							% Use plain style
\renewcommand{\section}[2]{\vskip 0.05em}		% Omit "References" title
\begin{thebibliography}{1}							% Simple bibliography with widest label of 1

\end{thebibliography}
}

\headerbox{Acknowledgements}{name=acknowledgements,column=0,below=references, above=bottom}{
\smaller						% Make the whole text smaller
\vspace{-0.4em}			% Save some space at the beginning
This research was partially supported by the Hungarian Government (KMOP-1.1.2-08/1-2008-0002 ) and the European Union's Seventh Framework Programme: DynaNets, FET-Open project no. FET-233847 (\url{http://www.dynanets.org}). The supports are gratefully acknowledged.
}

\headerbox{Dynamic Networks are Sensitive to Aggregation}{name=density,span=2,column=1,row=0}{
Network characteristics are extremely sensitive to minor changes in aggregation length. In our previous work \cite{prevWork1} \cite{prevWork2}, we studied the cumulative properties of Elementary Dynamic Network models over the complete time period (i.e., until they reach the stable point of a full network). Here we focus on the more realistc domain of sparse (cumulative) networks. We find that even when snapshot networks are stationary, \textbf{important network characteristics}  (average path lenght, clustering, betwenness centrality) \textbf{are extremely sensitive to aggregation} (window length).

%\includegraphics[angle=-90,width=0.98\linewidth]{PA_and_ER_Models_statisticalMeasures}

}

\headerbox{Degree Distribution Radically Changes}
{name=degreeDistribution,span=2,column=1,below=density,above=bottom}{
Degree distributions are exceptionally sensitive to the length of the aggregation window. \textbf{The same dynamic network may produce a normal, lognormal or even power law distribution for different aggregation lenghts.} The digree distribution of the snapshot and cumulative network is inherently different. The following surfaces show the CPA model until it approaches the complete network.
\vspace{-0.2em}
\begin{center}
%	\includegraphics[angle=-90,width=0.49\linewidth]{CPA_3d_snapshot}
%	\includegraphics[angle=-90,width=0.49\linewidth]{CPA_3d_cumulative}
\end{center}
\vspace{-0.2em}
Taking slices of the cumulative 3D charts shows us how the degree distribution changes. The log-log charts below show the progression of these changes as the aggregation window gets larger.
\vspace{-0.2em}
\begin{center}
%	\includegraphics[angle=-90,width=0.49\linewidth]{ER1_cumulativeDegrees}
%	\includegraphics[angle=-90,width=0.49\linewidth]{CPA_cumulativeDegrees}
\end{center}
}

\end{poster}
\end{document}

### 2017-02-20 Literature Review/MSCI POSTER draft.tex
poster

@@
 
+\usepackage{smartdiagram}
+\usepackage{metalogo}
+\usepackage{dtklogos}
+
 \begin{document}
 
-\title {\textbf{ Complexity science approach to Atrial Fibrillation}}
-\author{Tigany Zarrouk\\ CID: 00823932\\\textbf{Supervisor}: Kim Christensen \\ Condensed Matter Theory Group---Imperial College London}
+\title {\textbf{ A Sphere Model for Atrial Fibrillation}}
+\author{Tigany Zarrouk \& Mattia Gaggi \textbf{Supervisor}: Kim Christensen \\ Condensed Matter Theory Group---Imperial College London}
 
@@
 
-\begin{abstract}
-
-Modelling Atrial Fibrillation (AF) to ascertain its causes accurately, allowing for effective treatment, has been extensively studied over the past few decades.
-It is the largest cause of strokes linked with cardiac arrhythmia in aging populations and is due to impaired contractility of the atria in the heart, leading to blood clot formation.
-Action potential propagation, mediated by ionic species, are seen to be irregular in AF, which causes some atrial myocytes to contract at 300+ times per minute, much faster than healthy hearts which contract at 60-100 times per minute.
-This beating is regulated by the sinoatrial node which initiates action potentials that propagate through myocytes, causing contractions.
-Altered refractory period and conduction speed of signals has been seen to promote AF in various models and is caused physically by the fibrosis of cells, reducing the coupling of electrical signals transversally to some myocytes; giving rise to reentry circuits.
-The most prominent reentry model, the spiral wave model, has been seen to more closely follow clinical and experimental observations than other models of wavefront action.
-Cellular automata modelling paves the way for large scale simulations of tissue in the heart without the need to solve differential equations, which can be computationally costly.
-Moore neighbourhoods are seen to be more effective at producing isotropy of spiral waves, reducing flat edges, in rectilinear cellular automata models.
-Modelling scroll waves, using hybrid cellular automata models, using novel methods to find rotors and replicating the morphology of the heart with its heterogeneities more closely seem to be viable directions for further research.
-
-\end{abstract}
-
 \section{\textbf{\underline{Introduction}}}
 
-Medical science and the study of disease has been a crucial marker of society's success and development through the ages.
-Understanding disease gives insights into how to treat and diagnose serious conditions, of which can be fatal.
-One such condition that has eluded explanation and effective treatment is Atrial Fibrillation (AF), which is the leading cause of Ischaemic stroke in people over 75---seemingly due to incomplete contraction which leads to the formation of blood clots that can travel to the brain \cite{Hart}.
-It has been known that there are cardiovascular diseases that are predisposed to initiate AF, such as Congestive Heart Failure, Pericarditis and Hypertension through a combination of atrial pressure and dilation, but the precise mechanics are unknown.
-AF and VF (ventricular fibrillation) are thought to be caused by the breakup of waves of electrical impulses, that induce contraction, into interfering wavelets that produce chaotic behaviour.
+Atrial Fibrillation (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
+AF is also the most widespread cardiac condition, with over 30 millions of people worldwide suffering from it.
+In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ million pounds.Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time (also called fibrillation episodes).\\
 
-Healthy hearts have a regular, sinus  rhythm which is moderated by the sinoatrial node located in the right atrium of the heart.
-Fibrillatory activity in the atrium is characterised by a hastened activation of cells: atrial cells activate 300-600 times a minute in contrast to the expected sinus rhythm of 60-100 beats per minute.% Ectopic firing has been observed in mainly in the left atria.
+Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres and finally from the ventricular endocardium to the ventricular epicardium.
+The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell membrane.
+When a muscle cell (myocyte) is at rest, there is a built up potential difference between the outside and the inside of the cell (polarisation).
 
-Current treatment involves introducing antiarrhythmic drugs of which side effects include proarrhythmia, which can cause sudden death \cite{Nattel1}.
-Surgical methods primarily involve the Maze procedure, by which a maze-like pattern is cut into the atria such that there is only one path that the electrical impulse can take from the sinoatrial node to the atrioventricular node.
-The treatment is usually successful but still may lead to fibrillation \cite{Tsui}.
-The ablation of cells in has been shown to treat AF therapeutically.
-Catheter ablation has had suboptimal success rates.
-Electrogram-guided ablation is a burgeoning area of research.
-Complex-fractionated atrial electrograms (CFAEs) are proposed as sites for ablation as they seem to be critical for AF \cite{Greek}.
-% These sites seem to be mainly focussed around the pulmonary veins.
+%picture of the heart here
 
-Electrograms are used for non-contact mapping and can be used to elucidate areas of fibrillatory activity.
-The most popular types are unipolar and bipolar.
-Unipolar electrograms have problems subtracting far-field signal and the wavefront direction affects the amplitude of the wave recorded---in contrast to bipolar electrograms---but the morphology of the surface can indicate the direction of the wavefront.
-Bipolar electrodes offer the advantage of a superior signal-to-noise ratio but with markedly less amplitude in sensing depolarization \cite{Vincent}.
-%` whereas bipolar electrograms do not experience these problems.
+AF is usually produced  by an calcified muscle cells (fibrosis) in the cardiac tissue which generates a process called reentry.
+During reentry the electric signal wavefront propagation breaks.
+It has been shown a clear link between fibrosis and AF
+Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research.
+The current research revolves around three possible approaches:
 
-\section{\textbf{\underline{Electrophysiology of the heart}}}
+\smartdiagram[bubble diagram]{Current research,
+ Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) t}
 
-Signals propagate through the heart by an action potential arising due to the interplay of ions with pacemaker cells, in the sinoatrial and atrioventricular nodes, and myocardial cells in cardiac muscle \cite{Augustus}.
-See figure 1 for a diagram of the action potentials involved.
+ Our research is based on expanding on the wrok done by Kim Christensen on CA (Cellular Automata) models \cite{Christensen}.
 
- Myocardial cells in the heart have a potential difference and concentration gradient intrinsic to their structure: Na$^{+}$ and Cl$^{-}$ ions are higher in concentration outside of the cell, while K$^{+}$ and organic anions (A$^{-}$) are more concentrated on the inside.
-Ca$^{2+}$  ions are higher in concentration extracellularly but they are bound, mainly in the sarcoplasmic reticulum.
-This gives a resting membrane potential of $-75$mV, when no action potential is propagating, as the only ion that is allowed free movement is K$^{+}$ of which the concentration gradient and potential difference balance out the efflux (due to the concentration gradient) and influx (due to the negative resultant potential difference).
-This is phase 4 of the action potential.
-There are gates through which specific ions can flow into or out of the cell when a threshold voltage has been reached.
-After a sufficient stimulation, positive ions (mainly Na$^{+}$)  flood into the myocardial cell which causes rapid depolarisation (phase 0).
-Once depolarised enough, after a small efflux of K$^{+}$ and  Cl$^{-}$ ions (phase 1), Ca$^{2+}$ ion voltage-gated channels open, which induces Ca$^{2+}$ release from intracellular stores.
-This gives a plateau of potential difference (phase 2) where an influx of  K$^{+}$ ions sustains the voltage against the efflux of  Ca$^{2+}$ ions.
-The free calcium ions cause myocardial contraction.
-After a this long plateau phase,  K$^{+}$ ion channels reopen to repolarise the cell (phase 3) to get ready for another contraction, reverting to phase 4.
-The period where a cell cannot respond to stimuli to induce contraction is the refractory period.
+\section{\textbf{Kim Christensen Model}}
 
-%5 phases of the normal action potential:
+"Simple Model for Identifying Critical Regions in Atrial Fibrillation"~\cite{Christensen}produced a simple 2D model for AF.
+The cardiac cells were arranged in a squared grid with horizontal connections, with a percentage of transversal connections (to reproduce the effect of fibrosis) and a percentage of unexcitable cells.
 
-%Phase 4, or the resting potential, is stable at ≈−90 mV in normal working myocardial cells.
+Electric impulse propagation rules:
 
-%Phase 0 is the phase of rapid depolarization.
-The membrane potential shifts into positive voltage range.
-This phase is central to rapid propagation of the cardiac impulse (conduction velocity, θ=1 m/s).
+%insert fucking rules
 
-%Phase 1 is a phase of rapid repolarization. This phase sets the potential for the next phase of the action potential.
+The results from this model show that there is a threshold value of vertical connections beyond which AF is produced spontaneously (without changing refractory period or conduction speed).
+This matches real data according to which AF is correlated with fibrosis.
+The model also shows that burning the area affected by AF effectively stops the arrhythmia, which has been recently discovered to be the case .
 
-%Phase 2, a plateau phase, is the longest phase. It is unique among excitable cells and marks the phase of calcium entry into the cell.
-
-%Phase 3 is the phase of rapid repolarization that restores the membrane potential to its resting valu
-%Through this electrical conduction, the cardiac muscle (myocardium) contracts.
-
-In the sinoatrial node, the depolarising current is primarily fulfilled by Ca$^{2+}$ ion movement, which is slow in comparison to the fast Na$^{+}$ currents of the myocardial cells.
-SA node action potentials have only three phases.
-Phase 4 is the spontaneous depolarization (pacemaker potential) that triggers the action potential.
-Phase 0 is the depolarization phase of the action potential.
-This is followed by phase 3, repolarization.
-Once the cell is completely repolarized, the cycle is spontaneously repeated.
-
-\begin{figure}
-\caption[short title]{Diagram of the action potential in a myocardial cell (A) and in sinoatrial cell (B).
-Numbers indicate phases of the action potential.
-ERP = Effective refractory period \cite{Mizzi}}
-\centering
-\includegraphics[width = 0.5\textwidth]{newactionpot}
-\end{figure}
-
-Ectopic firing---from a focus, not located at the Sinoatrial node, at which action potentials are initiated---can arise from 'afterdepolarisations': the free intracellular Ca$^{2+}$ concentration rises sharply during the depolarised phase of the action potential; when relaxation occurs, the concentration of Ca$^{2+}$  is reduced by uptake into the sarcoplasmic reticulum via Na$^{+}$/ Ca$^{2+}$  exchange.
-This is an electrogenic process: three Na$^{+}$ ions are exchanged for a single Ca$^{2+}$, thus making the potential difference more positive.
-This induces an inward current that can depolarise the cell which can lead to spontaneous action potentials firing \cite{Nattel2}.
-
-Atrial cell fibrosis causes cells to be unresponsive.
-Current research suggests that this is due to excessive Ca$^{2+}$ loading, which occurs with tachycardia as many action potentials are enacted.
-Furthermore, this leads to electrical remodelling of the heart, making AF the more probable pathway of activation regardless of the mechanism for tachycardia \cite{Nattel2}.
-Due to fibrillation, the concentration of the intracellular Ca$^{2+}$ is altered, thus impairing effective contraction which may give rise to blood clots that can cause strokes.
-%REVISEEE
-
-%MONODOMAIN AND BIDOMAIN STUFF??
-
-\section{\textbf{\underline{Types of Models}}} %METHODS
+\section{\textbf{Implementing Restitution in Kim Christensen Model}} %METHODS
 
@@
 
-\section{\textbf{\underline{Cellular Automata Models and}\\ \underline{Further Developments}}}
+\section{\textbf{Kim Christensen model on a Sphere}}
 
@@
 %####################################################
-\bibitem{Hart}
-Hart, R. G. \& Halperin, J. L. (2001).
-\emph{Atrial fibrillation and stroke: concepts and controversies.},
-Stroke 32, 803–808,
-2001
-
-\bibitem{Nattel1}
-S. Nattel, 
-\emph{ Experimental evidence for proarrhythmic mechanisms of antiarrhythmic drugs.}
-Cardiovasc Res. 1998 Mar;37(3):567-77.
-
-\bibitem{Tsui}
-Tsui, S. S. L. Grace, A. A. Ludman, P. F. Schofield, P. M. Page, A. J. P. Rowland, E.  and Large, S. R., 
-\emph{ Maze 3 for Atrial Fibrillation: Two Cuts Too Few?.}
- Pacing and Clinical Electrophysiology,
-17: 2163–2166. doi:10.1111/j.1540-8159.1994.tb03819.x
-1994
-
-\bibitem{Greek}
-Demosthenes G Katritsis, Ioannis Pantos and Efstathios P Efstathopoulos, 
-\emph{Catheter ablation of atrial fibrillation guided by electrogram fractionation and dominant frequency analysis},
-Expert Review of Cardiovascular Therapy 
-Volume 9, 2011 - Issue 5, 631-636, 2014
-
-\bibitem{Vincent}
-V. DeCaprio, P. Hurzeler snd S. Furman, 
-\emph{A Comparison of Unipolar and Bipolar Electrograms for Cardiac Pacemaker Sensing}
-Circulation, Vol 56, No 5, November 1977
-
-\bibitem{Augustus}
-Augustus O. Grant, MB, ChB, PhD
-\emph{Basic Science for the Clinical Electrophysiologist: Cardiac Ion Channels}
-Circulation: Arrhythmia and Electrophysiology.
-2009; 2: 185-194
-doi: 10.1161/CIRCEP.108.789081
-
-\bibitem{Nattel3}
-Nattel S, Carlsson L.
-\emph{ Innovative approaches to anti-arrhythmic drug therapy.},
-Nat Rev Drug Discov. 2006; 5: 1034–1049.
-
-\bibitem{Mizzi}
-\emph{Amiodarone supplants lidocaine in ACLS and CPR protocols}
-A. Mizzi, T. Tran, D, Mangar, E. M. Camporesi
-Anesthesiology Clinics 29(3):535-45 · September 2011
-
-\bibitem{Nattel2}
-S. Nattel, 
-\emph{New Ideas about atrial fibrillation 50 years on},
- NATURE | VOL 415 | 10 JANUARY 2002 |
-
-\bibitem{Grandi}
-E. Grandi, S. V. Pandit, N. Voigt, M.D., A. J. Workman, D. Dobrev, J. Jalife, and D. M  Bers,
-\emph{Human Atrial Action Potential and Ca$^{2+ }$ Model: Sinus Rhythm and Chronic Atrial Fibrillation},
-Circ Res. 2011 October 14; 109(9): 1055–1066. doi:10.1161/CIRCRESAHA.111.253955.
-November 2011
-
-\bibitem{Fenton}
-F. Fenton and A. Karma
-\emph{Vortex dynamics in three-dimensional continuous myocardium with fiber
-rotation: Filament instability and fibrillation},
-1997
-
-\bibitem{FitzHugh}
-Richard FitzHugh, 
-\emph{Impulses and Physiological States in Theoretical Models of Nerve Membrane},
-Biophys J. 1961 Jul; 1(6): 445–466.
-
-\bibitem{Hodgkin}
-A. L. Hodgkin and A. F. Huxley, 
-\emph{A quantitative description of membrane current and its application to conduction and excitation in nerve},
-J. Physiol. (I952) I I7, 500-544
-
-\bibitem{Antze}
- C. Antzelevitch
-\emph{Basic mechanisms of re-entrant arrhythmias}
-Curr Opin Cardiol. 2001 Jan;16(1):1-7.
-
-\bibitem{Tusscher}
-\emph{Spiral wave dynamics and ventricular arrhythmias},
- K. H. W. Johanna ten Tusscher,
- November 2004.
-
-\bibitem{Allessie}
-Allessie MA, Bonke FIM, Schopman JG,
-\emph{ Circus movement in rabbit atrial muscle as a mechanism of tachycardia.
-III.
-The “leading circle” concept: a new model of circus movement in cardiac tissue without the involvement of an anatomical obstacle.},
- Circ Res 1977, 41:9–18.
-
-\bibitem{Greenburg}
-J. M. Greenberg and S. P. Hastings, 
-\emph{Spatial Patterns for Discrete Models of Diffusion in Excitable Media}, 
-SIAM J. APPL. MATH.
-Vol. 34, No. 3, May 1978
-
-\bibitem{Bub}
-G. Bub, A. Shrier, and L. Glass,
-\emph{Spiral Wave Generation in Heterogeneous Excitable Media},
- Phys. Rev. Lett.88, 058101 (2002)
-
-\bibitem{Davidenko}
-Davidenko JM, Kent PF, Chialvo DR, Michaels DC, Jalife J., 
-\emph{Sustained vortex-like waves in normal isolated ventricular muscle.},
- Proc Natl Acad Sci USA. 1990;87:8785–8789
 
@@
 DOI: http://dx.doi.org/10.1016/j.eupc.2005.05.011
-
-\bibitem{afstudy}
-S. M. Narayan, D. E. Krummen, and W.-J. RAPPEL, 
-\emph{Clinical mapping approach to diagnose
-electrical rotors and focal impulse sources for human atrial fibrillation},
- Journal of
-cardiovascular electrophysiology, vol. 23, no. 5, pp. 447–454, 2012.
-
-\bibitem{Skanes}
-Mandapati, R., Skanes, A., Chen, J., Berenfeld, O. and  Jalife, J. 
-\emph{Stable microreentrant sources as a mechanism of atrial fibrillation in the isolated sheep heart.}
-Circulation 101, 194–199 (2000).
 
@@
 
-\bibitem{Ozawa}
-Ashihara T, Haraguchi R, Nakazawa K, Namba T, Ikeda T, Nakazawa Y, Ozawa T, Ito M, Horie M, Trayanova NA., 
-\emph{The role of fibroblasts in complex fractionated electrograms during persistent/permanent atrial fibrillation: implications for electrogram-based catheter ablation.}
-Circ Res. 2012 Jan 20;110(2):275-84. doi: 10.1161/CIRCRESAHA.111.255026
-
-\bibitem{Felipe}
-F.A. Atienza, J. R. Carrion, A. G. Alberola, J. L. R. Alvarez, J. J. S. Munoz, J. M. Sanchez, M. V. Chavarri,
-\emph{A Probabilistic Model of Cardiac Electrical Activity Based on a Cellular Automata System},
-Revista Española de Cardiología (English Edition), Volume 58, Issue 1, January 2005, Pages 41-47,
-doi:10.1016/S1885-5857(06)60233-8
-
-\bibitem{Pour}
-F. Pourhasanzande, S. H. Sabzpoushan
-\emph{A new Cellular Automata Model of Cardiac
-Action Potential Propagation based on
-Summation of Excited Neighbors}, 
-International Journal of Medical, Health, Biomedical, Bioengineering and Pharmaceutical Engineering Vol:4, No:8, 2010
-
 \bibitem{Fedotov}
@@
 Biomedical Engineering, Vol. 49, No. 2, July, 2015, pp. 112115.
-
-\bibitem{Siregar}
-P. Siregar, J. P. Sinteff, N. Julen, P. Le Beux,
-\emph{An Interactive 3D Anisotropic Cellular Automata Model of the Heart}
-Computers and Biomedical Research
-Volume 31, Issue 5, October 1998, Pages 323-347, 
-doi:10.1006/cbmr.1998.1485
-
-\bibitem{Siregar2}
-P. Siregar, J.P. Sinteff, M. Chahine, P. Lebeux
-\emph{A Cellular Automata Model of the Heart and Its Coupling with a Qualitative Model}, 
-Computers and Biomedical Research
-Volume 29, Issue 3, June 1996, Pages 222-246,
-doi:10.1006/cbmr.1996.0017
-
-\bibitem{Bartocci}
-E. Bartocci, F. Corradini, M.R. Di Berardini, E. Entcheva and S.A. Smolka, R. Grosu, 
-\emph{Modeling and simulation of cardiac tissue using hybrid I/O automata}
-Theoretical Computer Science
-Volume 410, Issues 33–34, 21 August 2009, Pages 3149-3165,
-doi:10.1016/j.tcs.2009.02.042
-
-\bibitem{Luca}
-A. Luca, V. Jaquemet, N. Virag, J. M. Vesin
-\emph{Influence of Right and Left Atrial Tissue Heterogeneity on
-Atrial Fibrillation Perpetuation }
-Computing in Cardiology 2015; 42:449-452.
 

### 2017-02-21 Literature Review/MSCI POSTER draft.tex
poster stuff

@@
-\documentclass[twocolumn, a1paper]{article}
+\documentclass[twocolumn, a1paper, 20pt]{article}
  %\usepackage[scale=0.78,size=a1]{beamerposter}
@@
 The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell membrane.
-When a muscle cell (myocyte) is at rest, there is a built up potential difference between the outside and the inside of the cell (polarisation).
 
@@
 
-AF is usually produced  by an calcified muscle cells (fibrosis) in the cardiac tissue which generates a process called reentry.
-During reentry the electric signal wavefront propagation breaks.
-It has been shown a clear link between fibrosis and AF
-Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research.
+When the electric impulse propagates a single muscle cell goes through three stages:
+
+\begin{itemize}
+  \item \textit{Excitable state}: the muscle cell is at rest with a negative built up potential.
+
+  \item \textit{Excited state}: the muscle cell is excited by its excited neighbours and the Voltage of the cell is at its peak. 
+  \item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again/
+\end{itemize}
+
+Where the refractory period depends on the last time a cell was excited.\\
+
+AF is correlated to the amount of \textbf{ fibrosis} in the cardiac tissue which generates a process called reentry. \\
+During reentry the electric signal wavefront propagation breaks. It has been shown a clear link between fibrosis and AF
+Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research. \\
 The current research revolves around three possible approaches:
@@
 \smartdiagram[bubble diagram]{Current research,
- Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) t}
+ Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
 
@@
 
-"Simple Model for Identifying Critical Regions in Atrial Fibrillation"~\cite{Christensen}produced a simple 2D model for AF.
-The cardiac cells were arranged in a squared grid with horizontal connections, with a percentage of transversal connections (to reproduce the effect of fibrosis) and a percentage of unexcitable cells.
+\subsection{Description of the model}
 
-Electric impulse propagation rules:
+In the publication "a Simple Model for Identifying Critical Regions in Atrial Fibrillation"~\cite{Christensen}, Prof.
+Christensen produced a simple 2D model for AF.
+The cardiac cells were arranged in a squared grid with horizontal connections, with a percentage of \textbf{ transversal connections} (to reproduce the effect of fibrosis) and a percentage of \textbf{dysfunctional cells}.
+Dysfunctional cells are excited by neighbours with a fixed probability.\\
 
-%insert fucking rules
+\textbf{\textit{Propagation Rules the model}}\\
 
-The results from this model show that there is a threshold value of vertical connections beyond which AF is produced spontaneously (without changing refractory period or conduction speed).
+\begin{itemize}
+  \item \textit{Excitable state}: at time step t=0, each cell at rest is an element in the grid
+
+  \item \textit{Excited state}: at t=1, the cell is excited by one of its \textit{ 4 neighbour cells}, and it can excite one the excitable neighbour cells at the next time-step.
+  \item \textit{Refractory state} :from t=2 onwards, the cell goes through the refractory state for t=T time-steps before becoming excitable again/
+\end{itemize}
+
+\subsection{Results}
+
+The result show that AF naturally arises from the model itself.
+\textbf{ It is important to note that this model is the only model ever created where arrhythmia naturally occurs from the structure of the heart} instead of being put in by hand in some way (as for instance adding by hand a new source of electric impulses.
+
+ % insert picture here of the model with wavefronf broken etc
+
+ In addition the simulations produced by the model show a correlation between fibrillation and fibrosis .
+There is a threshold value of vertical connections beyond which AF is produced spontaneously (without changing refractory period or conduction speed).
 This matches real data according to which AF is correlated with fibrosis.
-The model also shows that burning the area affected by AF effectively stops the arrhythmia, which has been recently discovered to be the case .
+The model also shows that burning the area affected by AF effectively stops the arrhythmia, which has been recently discovered to be the case.
 
-\section{\textbf{Implementing Restitution in Kim Christensen Model}} %METHODS
+%insert picture
 
-Mechanisms thought to be responsible for active cardiac arrhythmias are either enhanced/abnormal pulse formation or re-entry.
-For the former, simulations have included modelling the ionic currents of action potentials such as the Fenton-Karma model \cite{Grandi} \cite{Fenton} and modelling the impulse of an action potential using generic reaction-diffusion equations, such as the FitzHugh-Nagumo Model \cite{FitzHugh}.
-Most of these models are based on the pioneering work of Hodgkin and Huxley who described the action potential of the giant squid axon, which bears similarities to cardiac cells \cite{Hodgkin}.
-%CHECKKKKKKKKKKKKKKKKKKK
- For all of these models bidomain or monodomain approaches can be used to model the electrical propagation in myocardial tissue. % intracellularly????
-  The bidomain model takes into account the anisotropy of intracellular and extracellular spaces.
-The monodomain model is a reduction of the bidomain model by assuming anisotropy ratios of intracellular and extracellular domains are equal---this is the model most commonly used in simulations today.%ET CETERA?
-% ,   is the Many types of models have been used to simulate fibrillatory behaviour in the heart, from, and signals for contraction, to cellular automata and
-
-Re-entry---the mechanism by which a propagating impulse fails to die out after normal activation of the heart---has been studied in depth via various circus reentry formulations: the Ring Model, the Leading circle model, the Figure-of-Eight model and the Spiral Wave model \cite{Antze} \cite{Tusscher}.
-The Ring model models a wave of propagation that moves along one, unidirectional pathway of heart tissue and returns to the origin of excitation, thus simulating a central area of inexcitable tissue.
-This simple model implies that the conduction velocity and the refractory period are both factors: the length of the circuit (pathlength) must be greater or equal to the wavelength (the distance travelled by an impulse during the refractory period).
-The smaller the wavelength, the greater the number of reentry circuits that can be accommodated during the refractory period to promote fibrillatory behaviour.
-See figure 2(c).
-
-The Spiral wave model---which consists of a single rotating wavefront around a core, simulating tachycardia---has been observed in systems that have sites that have excited, refractory and resting states in both homogeneous and heterogeneous media \cite{Greenburg} \cite{Bub}.
-This is due to anisotropies in conduction velocity in the cardiac tissue.
-See figure 2(b) and figure 3.
-% for comparison to a normal wavefront.
- Experimental demonstration of spiral waves has been achieved by in research by Davidenko et al and the concept is now seen to be an underlying driver of fibrillation; the paradigm appears to more closely correspond with many clinical and experimental observations than other models \cite{Davidenko} \cite{Comtois}.
-In one such observation, the presence of rotors was seen in 96\% of patients with AF, in a 2012 study \cite{afstudy}.
-The curvature of the spiral wave is related to the speed of the wave, the more curvature the slower the speed of propagation.
-High curvature, and far-field instability, leads to spiral wave breakage into daughter wavelets.
-This agrees with
-%popular classical modes of thought regarding the initiation of fibrillatory contraction, multiple circuit reentry, which in this case is due to spiral breakup, and 
-recent opinions where single, small reentry circuits or local generators have been found to promote fibrillatory activity---the core of the spiral \cite{Skanes} \cite{Nattel2}.
-The tip of the spiral wave has been seen to drift around the core, thus moving the whole spiral wave.
-Such meandering can be a factor in promoting fibrillatory activity \cite{Tusscher}.
-% Recent observations are in contradiction with the classical viewpoint,  . This is also concurrent 
- Extending the spiral wave model to 3D gives rise to scroll waves which propagate through multiple layers of cells and give rise to more complex wave dynamics and is therefore hard to implement computationally.
-%It has been seen that the if a spiral wave is in homogenous media then it will always return to its resting state .
-
-\begin{figure}
-\caption[short title]{Diagram showing a spiral wave reentry circuit.
-(a), a normal wavefront in the heart.
-(b) spiral wave, which simulates tachycardia, as can be seen by multiple propagating wavefronts in a small area.
-(c), turbulent activity due to break up of spiral wave giving rise to fibrillation \cite{Alonso}}
-\centering
-\includegraphics[width = 0.25\textwidth]{spiralbreak2tachy}
-\end{figure}
+\section{\textbf{Implementing Restitution in Kim Christensen Model}}
 
 \section{\textbf{Kim Christensen model on a Sphere}}
-
-Recent research has shown that rotating wavefronts can be initiated in regions where the proportion of transverse coupling between strands of tissue is reduced to a critical value, leading to spontaneous fibrillatory activity, in contrast to normal 'linear' wavefronts.
-This can be seen in the rectilinear cellular automata model developed by Christensen et al and reflects current opinion: the fibrosis of cells, which reduces the coupling of electrical signals to other cells, leads to AF\cite{Christensen}.
-This research also supports targeted ablation to stop AF.
-Other studies suggest similar mechanisms such as the coupling of fibroblasts to atrial myocytes heterogeneously gives rise to spiral wave breakups \cite{Ozawa}.
 
@@
 In addition, a multi-electrode system was created to study the localization of AF sources by simulated electrograms.
-
-More cellular automata models have been researched by Siregar et al, where a 3D model of the heart, with simplified morphology is simulated and, in addition, a 2D CA model is compared qualitative model \cite{Siregar} \cite{Siregar2}.
-A hybrid I/O cellular automata approach has also been studied in research by Bartocci et al to describe cardiac tissue and improve the efficiency of the simulation process by parallelization \cite{Bartocci}.
-
- The impact of left and right atrial tissue electrical heterogeneities to promote AF has been studied in a biophysical model and it was seen that AF is more persistent in the right atria \cite{Luca}.
-Further development could be achieved by simulation from a cellular automata perspective.
-Modelling scroll waves and replicating the morphology of the heart with its heterogeneities more closely seem to be viable directions for further research.
-The effects of antiarrhythmic drugs on the heart during AF can be studied, along with the associated mechanism by which proarrhythmia occurs, and methods for its rectification.
-Developing ways to find CFAEs with novel electrogram methods via simulation could help targeted ablation, stopping AF more effectively and efficiently.
-
-% Further developments could be obtained through this parallelization combined with 3D modelling, more accurately depicting the shape of the heart, making the code more efficient.
-
- %Many of these studies suggest targeted ablation can hinder or fully stop AF, therefore, electrogram-guided ablation to find CFAEs is a promising method it is concurrent with clinical observations.
-%electrogram-guided ablation to find CFAE.
-
-%Most of these models only work one or two dimensionally, assuming that the heart is a 2D surface and not a 3D
-
-%Cellular automaton models
-
-%MAIN PAPER FROM KIM CHRISTIENSEN
-%BIPOLAR ELECTROGRAM??
-%METHODS TO POTENTIALLY STOP ATRIAL FIBRILLATION
-
-%FUUUUUUUUUUUUURTHER METHODZZZZZZZZZZZZZZZZZZZZZZZZZZ
-
-%\section{\textbf{\underline{Implications and Discussion}}}  %What our research project will be able to achieve, potentially
 

### 2017-02-21 Literature Review/MSCI POSTER draft.tex
postert

@@
-\documentclass[twocolumn, a1paper, 20pt]{article}
+\documentclass[twocolumn, a1paper, 12pt]{article}
  %\usepackage[scale=0.78,size=a1]{beamerposter}
@@
   \item \textit{Excited state}: the muscle cell is excited by its excited neighbours and the Voltage of the cell is at its peak. 
-  \item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again/
+  \item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again - this period is called \textit{refractory period)}.
 \end{itemize}
 
-Where the refractory period depends on the last time a cell was excited.\\
+Where the refractory period depends on the last time a cell was excited and is always fixed to a constant.\\
 
@@
 
-\section{\textbf{Kim Christensen Model}}
+\section{\textbf{Kim Christensen's Model}}
 
@@
 
-\section{\textbf{Implementing Restitution in Kim Christensen Model}}
+\section{\textbf{Implementing Restitution }}
 
-\section{\textbf{Kim Christensen model on a Sphere}}
+Even though in Prof.Christensen model, the refractory period of each muscle cell was always fixed, this is not what happens in a real heart.
+In a real heart the refractory period depends on the heart beat rate and on the last time each cell it was excited.
+The relationship between refractory period and heart beat rate is called restitution.
+\\
+In the first part of our project, we implemented restitution in Kim Christensen model in order
+
+\section{\textbf{Kim Christensen's model on a Sphere}}
 

### 2017-02-21 Sphere Model/New Network With Proper Propagation/Network_Projection_Appended.py
some extra analysis

@@ -308,7 +308,7 @@ class run:
         if self.store==True:
                 self.excitedhistory.append(self.network.excited)
                 self.num_excited.append(len(self.network.excited))
-        plt.show()
+        #plt.show()
 
           
             
@@ -353,16 +353,16 @@ class run:
                 
     def plot_sphere_a(self, sph):
         self.fig = plt.figure()
-        #self.ax = self.fig.add_subplot(111, projection='3d')#.gca(projection='3d')projection = 'mollweide'
+        self.ax = self.fig.add_subplot(111, projection='3d')#.gca(projection='3d')projection = 'mollweide'
         #self.ax = self.fig.add_subplot(111)#, projection = '3d') #'mollweide')#.gca(projection='3d')projection = 'mollweide'
-        #self.ax.view_init(elev=90., azim=0)
+        self.ax.view_init(elev=40., azim=0)
         
         #self.ax.axis([-1,1,-1,1, -1, 1])
-        """
+        
         self.ax.set_xlim(-0.55, 0.55)
         self.ax.set_ylim(-0.55, 0.55)
         self.ax.set_zlim(-0.55, 0.55)
-        """
+        
         #plt.gca().set_aspect('equal')
         #axes = plt.gca()
         #axes.set_xlim([-4,4])
@@ -371,7 +371,7 @@ class run:
         #print("self.x", self.x)
         #self.avg = (self.xc+ self.yc+ self.zc)/3
         #self.x, self.y = sph.Mercator_Projection(-2.3*np.pi/8.)#sph.Mercator_Projection()#s#Mollewide
-        self.z = np.zeros(self.x.shape)
+        #self.z = np.zeros(self.x.shape)
         self.triangles = sph.ch.simplices
         #self.xy=np.vstack((self.x, self.y)).T
         #tri = Delaunay(self.xy)
@@ -381,13 +381,13 @@ class run:
         #self.surf = self.ax.
         #self.triangles = tri.simplices
         #mask = [s]
-        self.surf =plt.tripcolor(self.x, self.y,  self.triangles, facecolors = self.triangles[:,0])#facecolours = self.triangles[:,0], edgecolors = 'k')#, cmap=plt.cm.Greys_r, antialiased=False)
-        #self.surf = self.ax.plot_trisurf(self.x,self.y,self.z, triangles=sph.ch.simplices, cmap=plt.cm.Greys_r, alpha = 1)
[…]

### 2017-02-21 Sphere Model/New Network With Proper Propagation/risk_curve_Sphere_dysf001.py

@@ -51,15 +51,15 @@ p_transversalconn.append(0.01)
 p_transversalconn.append(0.005)
 np.asarray(p_transversalconn)
 
-extra_nu = [0.195, 0.205,0.215, 0.225, 0.235, 0.245,0.32, 0.34]
+extra_nu = [0.195, 0.205,0.215, 0.225, 0.235, 0.245,0.32, 0.34,0.33, 0.35, 0.36, 0.37, 0.38,0.39, 0.4]
 m_trans_conn = p_transversalconn.copy()
 for nu in extra_nu:
     m_trans_conn.append(nu)
 
 timet=100000
-number_of_systems=50
-prob_dysf=0.05
-prob_unexcitable=0.05
+number_of_systems=12
+prob_dysf=0.005
+prob_unexcitable=0.005
 excitationvalue=25
 heart=110
 threshold=350
@@ -143,13 +143,13 @@ for elements in m_trans_conn:
 
 
 
-fileobj0 = open('connections_sph_run1.pkl', 'wb')
-fileobj1 = open('dysf_grids_sph_run1.pkl', 'wb')
-fileobj2 = open('t_fib_data_sph_run1.pkl', 'wb')
-fileobj3 = open('t_in_fib_sph_run1.pkl', 'wb')
-fileobj6 = open('risk_fib_sph_run1.pkl', 'wb')
-fileobj4 = open('risk_sph_run1.pkl', 'wb')
-fileobj5 = open('riskerror_sph_run1.pkl', 'wb')
+fileobj0 = open('connections_sph_run3_dysf0005.pkl', 'wb')
+fileobj1 = open('dysf_grids_sph_run3_dysf0005.pkl', 'wb')
+fileobj2 = open('t_fib_data_sph_run3_dysf0005.pkl', 'wb')
+fileobj3 = open('t_in_fib_sph_run3_dysf0005.pkl', 'wb')
+fileobj6 = open('risk_fib_sph_run3_dysf0005.pkl', 'wb')
+fileobj4 = open('risk_sph_run3_dysf0005.pkl', 'wb')
+fileobj5 = open('riskerror_sph_run3_dysf0005.pkl', 'wb')
 
 
 pickle.dump(conn, fileobj0, -1)
@@ -215,6 +215,6 @@ ax.errorbar(m_trans_conn,risk, yerr=riskerror, fmt='cd')
 
 
 ax.legend()
-ax.set_title('Risk Curve')
+ax.set_title('Risk Curve--eps/delta=0.005')
 plt.show()    
 

### 2017-02-21 Sphere Model/New Network With Proper Propagation/risk_curve_Sphere_extra_nu.py

@@ -51,10 +51,15 @@ p_transversalconn.append(0.01)
 p_transversalconn.append(0.005)
 np.asarray(p_transversalconn)
 
-extra_nu = [0.195, 0.205,0.215, 0.225, 0.235, 0.245,0.32, 0.34]
+extra_nu1 = [0.195, 0.205,0.215, 0.225, 0.235, 0.245,0.32, 0.34]
+extra_nu2 = [0.33, 0.35, 0.36, 0.37, 0.38,0.39, 0.4 ]
 m_trans_conn = p_transversalconn.copy()
-for nu in extra_nu:
+for nu in extra_nu1:
     m_trans_conn.append(nu)
+for nu in extra_nu2:
+    m_trans_conn.append(nu)
+
+
 
 timet=100000
 number_of_systems=50
@@ -95,7 +100,7 @@ risk_fib = []
 risk=[]      #risk final output
 riskerror=[]   #risk std
 
-for elements in m_trans_conn:
+for elements in extra_nu2:
     n = Network.create_network(array_nodesindices = np.arange(len(colours)),
                    array_vertical = vconn,
                    array_transv = hconn,
@@ -143,13 +148,13 @@ for elements in m_trans_conn:
 
 
 
-fileobj0 = open('connections_sph_run1.pkl', 'wb')
-fileobj1 = open('dysf_grids_sph_run1.pkl', 'wb')
-fileobj2 = open('t_fib_data_sph_run1.pkl', 'wb')
-fileobj3 = open('t_in_fib_sph_run1.pkl', 'wb')
-fileobj6 = open('risk_fib_sph_run1.pkl', 'wb')
-fileobj4 = open('risk_sph_run1.pkl', 'wb')
-fileobj5 = open('riskerror_sph_run1.pkl', 'wb')
+fileobj0 = open('connections_sph_run1_extra_data.pkl', 'wb')
+fileobj1 = open('dysf_grids_sph_run1_extra_data.pkl', 'wb')
+fileobj2 = open('t_fib_data_sph_run1_extra_data.pkl', 'wb')
+fileobj3 = open('t_in_fib_sph_run1_extra_data.pkl', 'wb')
+fileobj6 = open('risk_fib_sph_run1_extra_data.pkl', 'wb')
+fileobj4 = open('risk_sph_run1_extra_data.pkl', 'wb')
+fileobj5 = open('riskerror_sph_run1_extra_data.pkl', 'wb')
 
 
[…]

### 2017-02-21 Sphere Model/New Network With Proper Propagation/risk_curve_restitution_datafromAF2.py

@@ -1,0 +1,185 @@
+import AF_restitution
+import numpy as np
+from matplotlib import pyplot as plt
+import pickle
+
+#Kishan's Datae
+kishmatrix = np.array([[0.02,0.99981,4.3015e-06],
+[0.04,0.99983,3.8088e-06],
+[0.06,0.9998,1.0454e-05],
+[0.08,0.99968,3.0663e-05],
+[0.11,0.99772,0.00044859],
+[0.13,0.96099,0.018246],
+[0.15,0.60984,0.054379],
+[0.17,0.16381,0.041092],
+[0.19,0.017807,0.0080603],
+[0.21,0.020737,0.016513],
+[0.23,4.922e-05,4.8685e-05],
+[0.25,0.0001084,8.4968e-05],
+[0.27,0,0],
+[0.29,0,0],
+[0.12,0.99152,0.0027053],
+[0.14,0.86184,0.028043],
+[0.16,0.29714,0.055185],
+[0.18,0.039206,0.013863],
+[0.2,0.0056277,0.0028284],
+[0.22,4.834e-05,3.6005e-05],
+[0.24,0.00082172,0.00081342],
+[0.26,0,0],
+[0.28,0,0],
+[0.3,9.406e-05,9.3115e-05],
+[0.1,0.99919,0.00010423]])
+
+
+kishnu = kishmatrix[:,0]
+kishrisk = kishmatrix[:,1]
+kisherror = kishmatrix[:,2]
+p_transversalconn= kishnu #np.arange(0,0.6,0.05) #nu
+p_transversalconn= list(kishnu) #np.arange(0,0.6,0.05) #nu
+p_transversalconn.append(0.01)
+p_transversalconn.append(0.005)
+np.asarray(p_transversalconn)
+
+time=100000
+number_of_systems=50
+risk=[]
+riskstd=[]
+t_fib_data = []
+t_fib_temp = []
+number_fib = []
+n_fib_data = []
+n_fib_avg = []
+n_fib_err = []
+t_in_fib = []
+
+#fileobj0 = open('realizationdata', 'wb')
+
+d=open('dysfgrids.pkl','rb')
+dysfgrids=pickle.load(d)
+f=open('fibrosisgrids.pkl','rb')
+fibrosisgrids=pickle.load(f)
+er=open('risk.pkl','rb')
+risk_o=pickle.load(er)
+fr=open('riskerrorr.pkl','rb')
+risk_o_std=pickle.load(fr)
+counter=0
+
+"""
[…]

### 2017-02-22 Literature Review/Atrial Fibrillation Poster 22feb.tex
Poster things
Added new file and template, if you could add in the stuff and text you
think is pertinent then that is great. I couldn't get the smartdiagram
to work with the beamerposter package, so we will need to find another
way to do it. Also the attenuation of text is necessary

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dreuw & Deselaer's Poster
% LaTeX Template
% Version 1.0 (11/04/13)
%
% Created by:
% Philippe Dreuw and Thomas Deselaers
% http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}

\usepackage[orientation=portrait,size=a1,scale=1.3]{beamerposter} % Use the beamerposter package for laying out the poster with a portrait orientation and an a0 paper size

\usetheme{I6pd2} % Use the I6pd2 theme supplied with this template

\usepackage[english]{babel} % English language/hyphenation

\usepackage{amsmath,amsthm,amssymb,latexsym} % For including math equations, theorems, symbols, etc

%\usepackage{times}\usefonttheme{professionalfonts}  % Uncomment to use Times as the main font
%\usefonttheme[onlymath]{serif} % Uncomment to use a Serif font within math environments

\boldmath % Use bold for everything within the math environment

\usepackage{booktabs} % Top and bottom rules for tables

\graphicspath{{figures/}} % Location of the graphics files

\usecaptiontemplate{\small\structure{\insertcaptionname~\insertcaptionnumber: }\insertcaption} % A fix for figure numbering

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{\huge A Sphere Model for Atrial Fibrillation} % Poster title

\author{Tigany Zarrouk \& Mattia Gaggi } % Author(s)

\institute{Supervisor: Kim Christensen.  Condensed Matter Theory Group---Imperial College London} % Institution(s)

%----------------------------------------------------------------------------------------
%	FOOTER TEXT
%----------------------------------------------------------------------------------------

\newcommand{\leftfoot}{} % Left footer text

\newcommand{\rightfoot}{} % Right footer text

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of two major columns, each of which can be subdivided further with another \begin{columns} block - the [t] argument aligns each column's content to the top

\begin{column}{.02\textwidth}\end{column} % Empty spacer column

\begin{column}{.465\textwidth} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{block}{Objectives}

\begin{enumerate}
\item Replicate original model of Atrial Fibrillation (AF) from Christensen \emph{et al.}
\item Incorporate restitution of heart cells into the model. 
\item Model on more realistic morphology via translating model onto a sphere and adding restitution. 
\item Analyse to see if there is enhanced risk with the new model, using new morphology and model basis, to create a more accurate AF model.
\end{enumerate}

\end{block}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}

\begin{itemize}
\item Atrial Fibrillation (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure. AF is also the most widespread cardiac condition, with over 30 millions of people worldwide suffering from it. In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ million pounds. Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time (also called fibrillation episodes).\\
\end{itemize}

\end{block}

\begin{block}{Phenomenology of Atrial Fibrillation}

	\begin{itemize}
	\item Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres 		and finally from the ventricular endocardium to the ventricular epicardium. The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell 		membrane.

	\item When the electric impulse propagates a single muscle cell goes through three stages:

		\begin{itemize}
 		 \item \textit{Excitable state}: the muscle cell is at rest with a negative built up potential.

  		\item \textit{Excited state}: the muscle cell is excited by its excited neighbours and the Voltage of the cell is at its peak. 
  		\item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again - this period is called \textit{refractory period)}.
		\end{itemize}

	\item Where the refractory period depends on the last time a cell was excited and is always fixed to a constant.\\

	\item AF is correlated to the amount of \textbf{ fibrosis} in the cardiac tissue which generates a process called reentry. \\
	During reentry the electric signal wavefront propagation breaks. It has been shown a clear link between fibrosis and AF
	Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research. \\
	The current research revolves around three possible approaches:
	\end{itemize}

%\begin{frame}
%\smartdiagram[bubble diagram]{Current research,
 %Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
% \end{frame}
 Our research is based on expanding on the work done by Kim Christensen on CA (Cellular Automata) models \cite{Christensen}.

\end{block}

\begin{block}{Results: Connectome}

%\centering
\begin{figure}
\includegraphics[width=0.7\linewidth]{connectome}
\caption{Hemisphere of possible connections defined on the sphere model. Each vertex corresponds to a face. Longitudinal connections are in red, transversal are in blue. Starts from a single pentagon on the icosphere and then can be seen to propagate outwards. }
\end{figure}

\end{block}

%\end{column}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{.03\textwidth}\end{column} % Empty spacer column

\begin{column}{.465\textwidth} % The second column

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

%------------------------------------------------

\begin{block}{Results: Sphere and Previous model}

%\centering
\begin{figure}
\includegraphics[width=0.8\linewidth]{combchristensenspherefib}
\caption{Figures of: Christensen \emph{et al.}'s model at times of t=1, t=60 and t=470,for a, b and c, respectively, and the sphere model at times of t=1, t=41 and t=241 for d, e and f respectively. Normal wavefront propagation seen starting from pentagonal node in the leftmost figures.In figure e, the excitation wavefront propagates around the sphere, mimicking Christensen \emph{et al.}'s model, b. On the right, chaotic, fibrillatory activity has arisen from reentry circuit spontaneously manifesting itself in c and f.}
\end{figure}

\end{block}

\begin{block}{Implementing Restitution }

In Christensen \emph{et al.}'s model, the refractory period of each muscle cell was always fixed, this is not what happens in a real heart.
In a real heart the refractory period depends on the heart beat rate and on the last time each cell it was excited. The relationship between refractory period and heart beat rate is called restitution.
\end{block}

\begin{block}{Model on a Sphere}

\begin{itemize}
\item In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}. It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed. In addition, a multi-electrode system was created to study the localization of AF sources by simulated electrograms. 
\end{itemize}

\end{block}
%------------------------------------------------

\begin{block}{Results: Figure}

%\centering
\begin{figure}
\includegraphics[width=0.65\linewidth]{xriskcurvesphere}
\caption{Figure of the Risk Curve of all the systems analysed. Y-axis: probablility of a realisation being in AF, x-axis: value of $\nu$, the  probability of a transversal connection between cells}
\end{figure}

\end{block}

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}
\begin{itemize}
\item A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of heart muscle fibre connectivity while incorporating the effects of fibrosis. This model was improved upon by the additon of the restitution process, making AF more risky. By translating the model to a sphere, the steady growth of heart muscle fibres from the sino-atrial node could be replicated, while simulating a more realisitc morphology of the atria. The risk of AF was seen to be higher, than with the original model. Verification of this model can be achieved with the analysis of the proportion of fibrosis in atrial heart tissue. 
\end{itemize}
\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\begin{block}{References}

\nocite{*} % Insert publications even if they are not cited in the poster
\small{\bibliographystyle{unsrt}
\bibliography{sample1}}

\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

%\setbeamercolor{block title}{fg=black,bg=orange!70} % Change the block title color

%\begin{block}{Contact Information}

%\begin{itemize}
%\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
%\item Email: \href{mailto:john@smith.com}{john@smith.com}
%\item Phone: +1 (000) 111 1111
%\end{itemize}

%\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the second column

\begin{column}{.015\textwidth}\end{column} % Empty spacer column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2017-02-22 Literature Review/NewAFPoster.tex
Poster things
Added new file and template, if you could add in the stuff and text you
think is pertinent then that is great. I couldn't get the smartdiagram
to work with the beamerposter package, so we will need to find another
way to do it. Also the attenuation of text is necessary

--- 2017-02-22 Literature Review/Atrial Fibrillation Poster 22feb.tex
@@
 
+\usepackage{smartdiagram}
+\usepackage{metalogo}
+\usepackage{dtklogos}
+
 \usecaptiontemplate{\small\structure{\insertcaptionname~\insertcaptionnumber: }\insertcaption} % A fix for figure numbering
@@
 
-\author{Tigany Zarrouk \& Mattia Gaggi } % Author(s)
+\author{Tigany Zarrouk and Mattia Gaggi } % Author(s)
 
@@
 \item Incorporate restitution of heart cells into the model. 
-\item Model on more realistic morphology via translating model onto a sphere and adding restitution. 
-\item Analyse to see if there is enhanced risk with the new model, using new morphology and model basis, to create a more accurate AF model.
+\item Add more features, such as a modified connectivity, new dynamics via modelling on a sphere. 
+\item Analyse to see if there is enhanced risk with the model, using new morphology and model basis, to create a more accurate AF model.
 \end{enumerate}
@@
 %\smartdiagram[bubble diagram]{Current research,
- %Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
+% Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
 % \end{frame}
@@
 
-\begin{block}{Results: Connectome}
+\end{column}
 
-%\centering
-\begin{figure}
-\includegraphics[width=0.7\linewidth]{connectome}
-\caption{Hemisphere of possible connections defined on the sphere model.
-Each vertex corresponds to a face.
-Longitudinal connections are in red, transversal are in blue.
-Starts from a single pentagon on the icosphere and then can be seen to propagate outwards.
-}
-\end{figure}
-
-\end{block}
+%\begin{column}{.43\textwidth} % The second subdivided column within the first main column
 
 %\end{column}
+%\end{columns} % End of the subdivision
+
+%\begin{itemize}
+%\item Curabitur sapien ligula, faucibus in feugiat quis, vestibulum a turpis.
+%\begin{itemize}
+%\item Phasellus quis nunc neque. Suspendisse mauris diam, suscipit non gravida in, placerat id enim. Ut nec ipsum in lectus ultrices sagittis.
+%\item Ut nec ipsum in lectus ultrices sagittis.
+%\item Phasellus quis nunc neque.
+%\end{itemize}
+%\end{itemize}
+
+%\end{block}
+
+%----------------------------------------------------------------------------------------
+%	METHODS
+%----------------------------------------------------------------------------------------
+
+%----------------------------------------------------------------------------------------
+%	MATHEMATICAL SECTION
+%----------------------------------------------------------------------------------------
 
@@
 
-%------------------------------------------------
-
-\begin{block}{Results: Sphere and Previous model}
-
-%\centering
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{combchristensenspherefib}
-\caption{Figures of: Christensen \emph{et al.}'s model at times of t=1, t=60 and t=470,for a, b and c, respectively, and the sphere model at times of t=1, t=41 and t=241 for d, e and f respectively.
-Normal wavefront propagation seen starting from pentagonal node in the leftmost figures.In figure e, the excitation wavefront propagates around the sphere, mimicking Christensen \emph{et al.}'s model, b.
-On the right, chaotic, fibrillatory activity has arisen from reentry circuit spontaneously manifesting itself in c and f.}
-\end{figure}
-
-\end{block}
-
 \begin{block}{Implementing Restitution }
@@
 The relationship between refractory period and heart beat rate is called restitution.
+\\
+In the first part of our project, we implemented restitution in Kim Christensen model in order
+
 \end{block}
@@
 \begin{figure}
-\includegraphics[width=0.65\linewidth]{xriskcurvesphere}
+\includegraphics[width=0.8\linewidth]{xriskcurvesphere}
 \caption{Figure of the Risk Curve of all the systems analysed. Y-axis: probablility of a realisation being in AF, x-axis: value of $\nu$, the  probability of a transversal connection between cells}
@@
 \end{block}
+%\begin{block}{References}
+
+%\nocite{*} % Insert publications even if they are not cited in the poster%
+%%\small{\bibliographystyle{unsrt}
+%\bibliography{sample}}
+
+%\begin{frame}[allowframebreaks]
+
+% \frametitle<presentation>{Further Reading}
+
+%\begin{thebibliography}{1}  
+  % \beamertemplatearticlebibitems
+ %\bibitem{Christensen}
+%\newblock Christensen \emph{et al.}, 
+%\newblock \emph{Simple Model To Identify Critical Regions in Atrial Fibrillation}
+%Phys. Rev. Lett. 114, 028104
+
+  %\end{thebibliography}
+%\end{frame}
+
+%\end{block}
+
+%----------------------------------------------------------------------------------------
+%	ACKNOWLEDGEMENTS
+%----------------------------------------------------------------------------------------
+
+\begin{block}{Acknowledgments}
+
+\begin{itemize}
+\item Thanks to Professor Kim Christensen with his guidance througout the research project. 
+\end{itemize}
+
+\end{block}
 
@@
 
-%\setbeamercolor{block title}{fg=black,bg=orange!70} % Change the block title color
+\setbeamercolor{block title}{fg=black,bg=orange!70} % Change the block title color
 
-%\begin{block}{Contact Information}
+\begin{block}{Contact Information}
 
-%\begin{itemize}
-%\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
-%\item Email: \href{mailto:john@smith.com}{john@smith.com}
-%\item Phone: +1 (000) 111 1111
-%\end{itemize}
+\begin{itemize}
 
-%\end{block}
+\item Email: \href{mailto:john@smith.com}{john@smith.com}
+\item Phone: +1 (000) 111 1111
+\end{itemize}
+
+\end{block}
 

### 2017-02-22 Literature Review/conference_poster_2.tex
Poster things
Added new file and template, if you could add in the stuff and text you
think is pertinent then that is great. I couldn't get the smartdiagram
to work with the beamerposter package, so we will need to find another
way to do it. Also the attenuation of text is necessary

--- 2017-02-22 Literature Review/NewAFPoster.tex
@@
 
-\usepackage{smartdiagram}
-\usepackage{metalogo}
-\usepackage{dtklogos}
-
 \usecaptiontemplate{\small\structure{\insertcaptionname~\insertcaptionnumber: }\insertcaption} % A fix for figure numbering
@@
 
-\title{\huge A Sphere Model for Atrial Fibrillation} % Poster title
+\title{\huge Unnecessarily Complicated Research Title} % Poster title
 
-\author{Tigany Zarrouk and Mattia Gaggi } % Author(s)
+\author{John Smith, James Smith and Jane Smith} % Author(s)
 
-\institute{Supervisor: Kim Christensen.  Condensed Matter Theory Group---Imperial College London} % Institution(s)
+\institute{Department and University Name} % Institution(s)
 
@@
 
-\newcommand{\leftfoot}{} % Left footer text
+\newcommand{\leftfoot}{http://www.LaTeXTemplates.com} % Left footer text
 
-\newcommand{\rightfoot}{} % Right footer text
+\newcommand{\rightfoot}{john@smith.com} % Right footer text
 
@@
 \begin{enumerate}
-\item Replicate original model of Atrial Fibrillation (AF) from Christensen \emph{et al.}
-\item Incorporate restitution of heart cells into the model. 
-\item Add more features, such as a modified connectivity, new dynamics via modelling on a sphere. 
-\item Analyse to see if there is enhanced risk with the model, using new morphology and model basis, to create a more accurate AF model.
+\item Donec fringilla, velit id lobortis commodo, eros dui consectetur mi, ut interdum lorem dui sed mauris.
+\item Nulla ac nulla rhoncus est bibendum ullamcorper:
+\item Quisque vestibulum, nisl sit amet gravida ultricies dis parturient montes, nascetur ridiculus musobortis commodo, eros dui consectetur mi.
 \end{enumerate}
@@
 \begin{itemize}
-\item Atrial Fibrillation (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
-AF is also the most widespread cardiac condition, with over 30 millions of people worldwide suffering from it.
-In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ million pounds.
-Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time (also called fibrillation episodes).\\
+\item Donec fringilla, velit id lobortis commodo, eros dui consectetur mi, ut interdum lorem dui sed mauris.
+Duis id sem nunc, a pharetra odio.
+Phasellus posuere \alert{semper massa}, id bibendum ligula tristique at.
+Integer sit amet vulputate turpis.
+Sed erat lacus, faucibus at viverra et, mattis nec sem.
+Cras faucibus \alert{scelerisque} cursus.
+Opet volutpat ligula.
+Duis semper lorem eget dui dignissim porttitor.
+Nulla facilisi.
+In ullamcorper lorem quis dolor iaculis nec egestas enim ultricies.
+Cras ut mauris elit, ut lacinia dui.
+Proin in ante et libero hendrerit iaculis.
+Nulla eu erat a urna laoreet auctor id a turpis.
+Nam mollis tristique neque eu luctus.
+Suspendisse rutrum congue nisi sed convallis.
+Aenean id neque dolor.
 \end{itemize}
@@
 
-\begin{block}{Phenomenology of Atrial Fibrillation}
+%----------------------------------------------------------------------------------------
+%	MATERIALS
+%----------------------------------------------------------------------------------------
+
+\begin{block}{Materials}
+
+\begin{columns} % Subdivide the first main column
+\begin{column}{.54\textwidth} % The first subdivided column within the first main column
+\begin{itemize}
+\item Vestibulum nisl, quis euismod velit eros in ligula.
+\begin{itemize}
+\item Cras rhoncus quam et augue convallis in elementum urna tincidunt.
+\end{itemize}
+\item Proin ut vestibulum augue.
+\begin{itemize}
+\item Donec dapibus sagittis neque eu ultrices.
+\end{itemize}
+\end{itemize}
+\end{column}
+
+\begin{column}{.43\textwidth} % The second subdivided column within the first main column
+\centering
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{placeholder.jpg}
+\caption{Figure caption}
+\end{figure}
+\end{column}
+\end{columns} % End of the subdivision
 
 \begin{itemize}
-	\item Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres 		and finally from the ventricular endocardium to the ventricular epicardium.
-The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell 		membrane.
-
-	\item When the electric impulse propagates a single muscle cell goes through three stages:
-
+\item Curabitur sapien ligula, faucibus in feugiat quis, vestibulum a turpis.
 \begin{itemize}
- 		 \item \textit{Excitable state}: the muscle cell is at rest with a negative built up potential.
-
-  		\item \textit{Excited state}: the muscle cell is excited by its excited neighbours and the Voltage of the cell is at its peak. 
-  		\item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again - this period is called \textit{refractory period)}.
+\item Phasellus quis nunc neque. Suspendisse mauris diam, suscipit non gravida in, placerat id enim. Ut nec ipsum in lectus ultrices sagittis.
+\item Ut nec ipsum in lectus ultrices sagittis.
+\item Phasellus quis nunc neque.
+\end{itemize}
 \end{itemize}
 
-	\item Where the refractory period depends on the last time a cell was excited and is always fixed to a constant.\\
-
-	\item AF is correlated to the amount of \textbf{ fibrosis} in the cardiac tissue which generates a process called reentry. \\
-	During reentry the electric signal wavefront propagation breaks. It has been shown a clear link between fibrosis and AF
-	Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research. \\
-	The current research revolves around three possible approaches:
-	\end{itemize}
-
-%\begin{frame}
-%\smartdiagram[bubble diagram]{Current research,
-% Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
-% \end{frame}
- Our research is based on expanding on the work done by Kim Christensen on CA (Cellular Automata) models \cite{Christensen}.
-
 \end{block}
-
-\end{column}
-
-%\begin{column}{.43\textwidth} % The second subdivided column within the first main column
-
-%\end{column}
-%\end{columns} % End of the subdivision
-
-%\begin{itemize}
-%\item Curabitur sapien ligula, faucibus in feugiat quis, vestibulum a turpis.
-%\begin{itemize}
-%\item Phasellus quis nunc neque. Suspendisse mauris diam, suscipit non gravida in, placerat id enim. Ut nec ipsum in lectus ultrices sagittis.
-%\item Ut nec ipsum in lectus ultrices sagittis.
-%\item Phasellus quis nunc neque.
-%\end{itemize}
-%\end{itemize}
-
-%\end{block}
 
@@
 
+\begin{block}{Methods}
+
+\begin{itemize}
+\item Maecenas Vel Nisl Elit
+\begin{itemize}
+\item Suspendisse potenti. Fusce a est eget turpis rhoncus varius sed sed dui. Cras justo nibh, bibendum a cursus eget, consequat et dui. Maecenas vel nisl elit, sed dignissim dolor. 
+\item In hac habitasse platea dictumst.
+\end{itemize}
+
+\item Viewpoint Matching Constraints
+\begin{itemize}
+\item Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. 
+\item Proin in nisi diam.
+\item Nam ultricies pellentesque nunc, ultrices volutpat nisl ultrices a.
+\end{itemize}
+
+\item Volutpat 
+\begin{itemize}
+\item Duis semper lorem eget dui dignissim porttitor.
+\item Nulla facilisi. In ullamcorper lorem quis dolor.
+\end{itemize}
+\end{itemize}
+
+\end{block}
+
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+
+\begin{block}{Mathematical Section}
+
+\begin{itemize}
+\item Maecenas Ultricies Feugiat Velit Non Mattis.
+\begin{itemize}
+\item Duis ante erat, bibendum nec tempus nec, interdum quis est.
+Nulla at mollis tortor.
+Phasellus quis leo dolor, aliquam laoreet orci $X$ Donec dapibus sagittis neque eu nec, interdum quis est.
+$Y_n, n=1,\cdots,N$ ndum nec tempus nec, interd
+\begin{align*}
+X \rightarrow r(X) & = \arg \max_{c} \Big\{ \max_n \big\{ \sum_{x_i \in X} \delta(x_i,Y_{n,c})\big\} \Big\} 
+\end{align*}
+\item Cras faucibus scelerisque cursus. Proin ut vestibulum augue. $\delta(x_i,Y_{n,c})$
+\end{itemize}
+\item Fusce tempus arcu id ligula varius dictum.
+Donec ut nisl dui, ac consectetur elit.
+In nec enim porta augue venenatis sollicitudin.
+Phasellus quis nunc neque.
+Suspendisse mauris diam, suscipit non gravida in, placerat id enim.
+Ut nec ipsum in lectus ultrices sagittis.
+\end{itemize}
+
+\end{block}
 
@@
 
-\begin{block}{Implementing Restitution }
+\begin{block}{Results: Table}
 
-In Christensen \emph{et al.}'s model, the refractory period of each muscle cell was always fixed, this is not what happens in a real heart.
-In a real heart the refractory period depends on the heart beat rate and on the last time each cell it was excited.
-The relationship between refractory period and heart beat rate is called restitution.
-\\
-In the first part of our project, we implemented restitution in Kim Christensen model in order
+\begin{itemize}
+\item Ased Aliquet Luctus Lectus
+\end{itemize}
+
+\begin{table}
+\begin{tabular}{l l l}
+\toprule
+\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
+\midrule
+Treatment 1 & 0.0003262 & 0.562 \\
+Treatment 2 & 0.0015681 & 0.910 \\
+Treatment 3 & 0.0009271 & 0.296 \\
+\bottomrule
+\end{tabular}
+\caption{Table caption}
+\end{table}
+
+\begin{itemize}
+\item Sollicitudin Vel Orci
+\item Maecenas Ultricies Feugiat Velit Non Mattis.
+\end{itemize}
+
+\begin{table}
+\begin{tabular}{l l l}
+\toprule
+\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
+\midrule
+Treatment 1 & 0.0003262 & 0.562 \\
+Treatment 2 & 0.0015681 & 0.910 \\
+Treatment 3 & 0.0009271 & 0.296 \\
+\bottomrule
+\end{tabular}
+\caption{Table caption}
+\end{table}
 
@@
 
-\begin{block}{Model on a Sphere}
-
-\begin{itemize}
-\item In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
-It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed.
-In addition, a multi-electrode system was created to study the localization of AF sources by simulated electrograms.
-\end{itemize}
-
-\end{block}
 %------------------------------------------------
@@
 
-%\centering
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{xriskcurvesphere}
-\caption{Figure of the Risk Curve of all the systems analysed. Y-axis: probablility of a realisation being in AF, x-axis: value of $\nu$, the  probability of a transversal connection between cells}
+\includegraphics[width=0.8\linewidth]{placeholder.jpg}
+\caption{Figure caption}
 \end{figure}
@@
 \begin{block}{Conclusion}
+
 \begin{itemize}
-\item A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of heart muscle fibre connectivity while incorporating the effects of fibrosis.
-This model was improved upon by the additon of the restitution process, making AF more risky.
-By translating the model to a sphere, the steady growth of heart muscle fibres from the sino-atrial node could be replicated, while simulating a more realisitc morphology of the atria.
-The risk of AF was seen to be higher, than with the original model.
-Verification of this model can be achieved with the analysis of the proportion of fibrosis in atrial heart tissue.
+\item Opet volutpat ligula.
+Duis semper lorem eget dui dignissim porttitor.
+Nulla facilisi.
+In ullamcorper lorem quis dolor iaculis nec egestas enim ultricies.
+Cras ut mauris elit, ut lacinia dui.
+Proin in ante et libero hendrerit iaculis.
+\item Nulla eu erat a urna laoreet auctor id a turpis. Nam mollis tristique neque eu luctus. Suspendisse rutrum congue nisi sed convallis. 
+\item Aenean id neque dolor.
+\item Opet volutpat ligula.
+Duis semper lorem eget dui dignissim porttitor.
+Nulla facilisi.
+In ullamcorper lorem quis dolor iaculis nec egestas enim ultricies.
+Cras ut mauris elit, ut lacinia dui.
+Proin in ante et libero hendrerit iaculis.
 \end{itemize}
+
 \end{block}
@@
 \small{\bibliographystyle{unsrt}
-\bibliography{sample1}}
+\bibliography{sample}}
 
 \end{block}
-%\begin{block}{References}
-
-%\nocite{*} % Insert publications even if they are not cited in the poster%
-%%\small{\bibliographystyle{unsrt}
-%\bibliography{sample}}
-
-%\begin{frame}[allowframebreaks]
-
-% \frametitle<presentation>{Further Reading}
-
-%\begin{thebibliography}{1}  
-  % \beamertemplatearticlebibitems
- %\bibitem{Christensen}
-%\newblock Christensen \emph{et al.}, 
-%\newblock \emph{Simple Model To Identify Critical Regions in Atrial Fibrillation}
-%Phys. Rev. Lett. 114, 028104
-
-  %\end{thebibliography}
-%\end{frame}
-
-%\end{block}
 
@@
 \begin{itemize}
-\item Thanks to Professor Kim Christensen with his guidance througout the research project. 
+\item Nam mollis tristique neque eu luctus.
+Suspendisse rutrum congue nisi sed convallis.
+Aenean id neque dolor.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
 \end{itemize}
@@
 \begin{itemize}
-
+\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
 \item Email: \href{mailto:john@smith.com}{john@smith.com}

### 2017-02-22 Literature Review/somedraft.tex
Poster things
Added new file and template, if you could add in the stuff and text you
think is pertinent then that is great. I couldn't get the smartdiagram
to work with the beamerposter package, so we will need to find another
way to do it. Also the attenuation of text is necessary

--- 2017-02-22 Literature Review/NewAFPoster.tex
@@
 
-\end{column}
-
-%\begin{column}{.43\textwidth} % The second subdivided column within the first main column
-
 %\end{column}
-%\end{columns} % End of the subdivision
-
-%\begin{itemize}
-%\item Curabitur sapien ligula, faucibus in feugiat quis, vestibulum a turpis.
-%\begin{itemize}
-%\item Phasellus quis nunc neque. Suspendisse mauris diam, suscipit non gravida in, placerat id enim. Ut nec ipsum in lectus ultrices sagittis.
-%\item Ut nec ipsum in lectus ultrices sagittis.
-%\item Phasellus quis nunc neque.
-%\end{itemize}
-%\end{itemize}
-
-%\end{block}
-
-%----------------------------------------------------------------------------------------
-%	METHODS
-%----------------------------------------------------------------------------------------
-
-%----------------------------------------------------------------------------------------
-%	MATHEMATICAL SECTION
-%----------------------------------------------------------------------------------------
 
@@
 %----------------------------------------------------------------------------------------
+
+%------------------------------------------------
 
@@
 \end{block}
-%\begin{block}{References}
-
-%\nocite{*} % Insert publications even if they are not cited in the poster%
-%%\small{\bibliographystyle{unsrt}
-%\bibliography{sample}}
-
-%\begin{frame}[allowframebreaks]
-
-% \frametitle<presentation>{Further Reading}
-
-%\begin{thebibliography}{1}  
-  % \beamertemplatearticlebibitems
- %\bibitem{Christensen}
-%\newblock Christensen \emph{et al.}, 
-%\newblock \emph{Simple Model To Identify Critical Regions in Atrial Fibrillation}
-%Phys. Rev. Lett. 114, 028104
-
-  %\end{thebibliography}
-%\end{frame}
-
-%\end{block}
 

### 2017-02-22 Literature Review/somedraft2.tex
Poster things
Added new file and template, if you could add in the stuff and text you
think is pertinent then that is great. I couldn't get the smartdiagram
to work with the beamerposter package, so we will need to find another
way to do it. Also the attenuation of text is necessary

--- 2017-02-22 Literature Review/NewAFPoster.tex
@@
 
-%\begin{frame}
 %\smartdiagram[bubble diagram]{Current research,
 % Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
-% \end{frame}
- Our research is based on expanding on the work done by Kim Christensen on CA (Cellular Automata) models \cite{Christensen}.
+
+ %Our research is based on expanding on the work done by Kim Christensen on CA (Cellular Automata) models \cite{Christensen}.
 
@@
 %\end{column}
-%\end{columns} % End of the subdivision
+\end{columns} % End of the subdivision
 
@@
 \begin{itemize}
-\item In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
+In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
 It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed.
@@
 \begin{itemize}
-\item A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of heart muscle fibre connectivity while incorporating the effects of fibrosis.
+A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of heart muscle fibre connectivity while incorporating the effects of fibrosis.
 This model was improved upon by the additon of the restitution process, making AF more risky.

### 2017-02-22 Risk Curve Data/Risk Curve 13-11/FinalAFplot.py

@@ -1,0 +1,415 @@
+import numpy as np
+from numpy import vectorize
+import pickle
+#import dill
+import time
+import types
+import random
+from matplotlib import pyplot as plt
+from matplotlib import animation
+from matplotlib import gridspec
+from itertools import compress
+import math
+import copy
+def decision(probability):
+    return random.random() < probability
+
+gridintime=[] #List for storage of the grid
+
+class heart:
+    def __init__(self,L=200,p_unexcitable=0.05,p_fibrosis= 0.2,p_dysf=0.05):
+
+        """#########################PLAN######################
+        
+      
+
+        1) Don't use a list of refractory cells, therefore they just update the grid with excited cells and then subtract from the array
+
+        2) Use vectors for the excited cells
+    
+
+	"""
+        self.L=L 
+        self.p_dysf=p_dysf #The fraction of a dysfunctional cells 
+        self.p_fibrosis=p_fibrosis #The fraction of missing transversal connections ( Nu)
+        self.p_unexcitable=p_unexcitable #Probablility of a dysfunctional cell being unexcitable
+        self.excitation=50 #Value of excitation on the lattice
+        self.heartbeatsteps=220 #Time period between excitation wavefronts
+        self.grid = np.zeros((self.L,self.L))
+        self.gridofexcite = copy.deepcopy(self.grid)
+        
+        
+        self.electrocardiovertpos=np.zeros((self.L,self.L))
+        self.electrocardiohorizpos=np.zeros((self.L,self.L))
+        self.volt=0
+                                        #list of edges True or False for every element self.grid[a,b] the element self.edges[a][b] is the edge below the elements
+
+        self.time=0
+        self.tcounter=[0]
+   
+
+
+        self.dysfgrid = np.random.rand(self.L, self.L)    #grid of random numbers 
+        self.dysfgrid[self.dysfgrid < self.p_dysf] = 1
+        self.dysfgrid[self.dysfgrid != 1] = 0
+		
+	
+        self.edgegrid = np.random.rand(self.L, self.L) #grid of random numbers 
+        self.edgegrid[self.edgegrid < self.p_fibrosis] = 1
+        self.edgegrid[self.edgegrid != 1] = 0
+
+      
+    def reinitialise(self):
+        
+        self.time=0
+        self.tcounter=[0]
+        self.grid = np.zeros((self.L,self.L))
+        self.gridofexcite = copy.deepcopy(self.grid)
+
[…]

### 2017-02-22 Sphere Model/New Network With Proper Propagation/Network_Projection_Appended.py

@@ -517,16 +517,19 @@ class Define_Connections:
         fig2 = plt.figure()
         ax2 = fig2.gca(projection='3d')
         #ax2.plot(phi_avg[:,0], phi_avg[:,1], phi_avg[:,2], 'r', markersize=1)
+        ax2.set_xlim(-0.55, 0.55)
+        ax2.set_ylim(-0.55, 0.55)
+        ax2.set_zlim(-0.55, 0.55)
         ax2.add_collection(lc1)
         ax2.add_collection(lc2)
         plt.show()
 
 
-"""
-s = sp.Sphere( recursion_level = 6 )
-#opening pickled files for reinstatement of connections
-s.construct_icosphere()
 
+s = sp.Sphere( recursion_level = 5 )
+#opening pickled files for reinstatement of connections
+#'s.construct_icosphere()
+"""
 d=open('horiz_conn_rec_6xconn.pkl', 'rb')
 hconn=pickle.load(d)
 
@@ -539,28 +542,37 @@ pent_ind=pickle.load(e)
 g=open('colours_rec_6xconn.pkl', 'rb')
 colours=pickle.load(g)
 
+#closing pickle files
+d.close()
+e.close()
+f.close()
+g.close()
+#h.close()
 #h=open('sph_rec_6xconn.pkl', 'rb')
 #s=pickle.load(h)
-
 """
 
 
-#conn = Define_Connections(s)
-#colours, vconn, hconn, pent_ind = conn.define_connections() #not needed if using pickled data
+
+conn = Define_Connections(s)
+colours, vconn, hconn, pent_ind = conn.define_connections() #not needed if using pickled data
 
 
 
-#n = create_network(array_nodesindices = np.arange(len(colours)),
-#                   array_vertical = vconn,
-#                   array_transv = hconn,
-#                   p_transv = 0.27,
-#                   impulse_start = pent_ind,
-#                   p_dysf = 0.05,
- #                  p_unexcitable = 0.05,
-#                   excitation = 25, 
- #                  hbs = 110)
+n = create_network(array_nodesindices = np.arange(len(colours)),
+                   array_vertical = vconn,
+                   array_transv = hconn,
+                   p_transv = 0.1,
+                   impulse_start = pent_ind,
+                   p_dysf = 0.05,
+                   p_unexcitable = 0.05,
+                   excitation = 25, 
+                   hbs = 110)
 
[…]

### 2017-02-22 Sphere Model/New Network With Proper Propagation/ico_prop_projection.py

@@ -351,8 +351,9 @@ class Sphere:
         '''Makes the icosphere using convex hull'''
         self.icosahedron_vertices = self.normalize_v3(np.array(self.icosahedron_vertices)) #normalizing icosahedron vertices of unit sphere
         self.icosahedron_vertices.tolist()
-        
+        #print("self.faces", self.faces)
         for face in self.faces:
+            #print("face", face)
             new_face_vert = [self.icosahedron_vertices[face[0]], self.icosahedron_vertices[face[1]], self.icosahedron_vertices[face[2]]]
             self.face_cache.append(new_face_vert)
         self.faces = np.asarray(self.face_cache ) #this is an array of points that correspond the icosahedron faces
@@ -371,7 +372,14 @@ class Sphere:
         surf.set_array(colours)
         self.icosahedron_vertices=np.asarray(self.icosahedron_vertices)
         ax.scatter(self.icosahedron_vertices[:,0],self.icosahedron_vertices[:,1], self.icosahedron_vertices[:,2], c='red')
+        ax.view_init(elev=3, azim=-169)
         
+        #self.ax.axis([-1,1,-1,1, -1, 1])
+        
+        ax.set_xlim(-0.55, 0.55)
+        ax.set_ylim(-0.55, 0.55)
+        ax.set_zlim(-0.55, 0.55)
+        ax.set_title('Sphere Plot, t=0')
         plt.show()
         return surf
     """

### 2017-02-22 Sphere Model/New Network With Proper Propagation/risk_curve_Sphere_dysf001.py

@@ -58,8 +58,8 @@ for nu in extra_nu:
 
 timet=100000
 number_of_systems=12
-prob_dysf=0.005
-prob_unexcitable=0.005
+prob_dysf=0.1
+prob_unexcitable=0.1
 excitationvalue=25
 heart=110
 threshold=350
@@ -143,13 +143,13 @@ for elements in m_trans_conn:
 
 
 
-fileobj0 = open('connections_sph_run3_dysf0005.pkl', 'wb')
-fileobj1 = open('dysf_grids_sph_run3_dysf0005.pkl', 'wb')
-fileobj2 = open('t_fib_data_sph_run3_dysf0005.pkl', 'wb')
-fileobj3 = open('t_in_fib_sph_run3_dysf0005.pkl', 'wb')
-fileobj6 = open('risk_fib_sph_run3_dysf0005.pkl', 'wb')
-fileobj4 = open('risk_sph_run3_dysf0005.pkl', 'wb')
-fileobj5 = open('riskerror_sph_run3_dysf0005.pkl', 'wb')
+fileobj0 = open('connections_sph_run4_dysf01.pkl', 'wb')
+fileobj1 = open('dysf_grids_sph_run4_dysf01.pkl', 'wb')
+fileobj2 = open('t_fib_data_sph_run4_dysf01.pkl', 'wb')
+fileobj3 = open('t_in_fib_sph_run4_dysf01.pkl', 'wb')
+fileobj6 = open('risk_fib_sph_run4_dysf01.pkl', 'wb')
+fileobj4 = open('risk_sph_run4_dysf01.pkl', 'wb')
+fileobj5 = open('riskerror_sph_run4_dysf01.pkl', 'wb')
 
 
 pickle.dump(conn, fileobj0, -1)
@@ -215,6 +215,6 @@ ax.errorbar(m_trans_conn,risk, yerr=riskerror, fmt='cd')
 
 
 ax.legend()
-ax.set_title('Risk Curve--eps/delta=0.005')
+ax.set_title('Risk Curve--eps/delta=0.1')
 plt.show()    
 

### 2017-02-23 Sphere Model/New Network With Proper Propagation/risk_curve_restitution_sphere_with_old_sphere_data.py
Restitution Sphere Model
Made a file to create the graph necessary to make the restitution risk
curve for the sphere. There is an error with the
Network_Projection_withrestitution file thought. line 192, index error.
Can you sort this out please?

@@ -1,21 +1,9 @@
+import Network_Projection_withrestitution as Network
 import numpy as np
-from scipy.sparse import csr_matrix
-import operator
-from itertools import chain
-import time
-from matplotlib import animation
 from matplotlib import pyplot as plt
-from matplotlib import cm
-from mpl_toolkits.mplot3d import Axes3D
-import pylab as pl
-from matplotlib import collections  as mc
-from mpl_toolkits.mplot3d.art3d import Line3DCollection
-import types
 import pickle
-import ico_prop_projection as sp
-import Network_Projection_Appended as Network
-import copy
-
+import time
+#Kishan's Datae
 kishmatrix = np.array([[0.02,0.99981,4.3015e-06],
 [0.04,0.99983,3.8088e-06],
 [0.06,0.9998,1.0454e-05],
@@ -42,6 +30,7 @@ kishmatrix = np.array([[0.02,0.99981,4.3015e-06],
 [0.3,9.406e-05,9.3115e-05],
 [0.1,0.99919,0.00010423]])
 
+
 kishnu = kishmatrix[:,0]
 kishrisk = kishmatrix[:,1]
 kisherror = kishmatrix[:,2]
@@ -58,15 +47,61 @@ for nu in extra_nu:
 
 timet=100000
 number_of_systems=12
-prob_dysf=0.1
-prob_unexcitable=0.1
+prob_dysf=0.05
+prob_unexcitable=0.05
 excitationvalue=25
 heart=110
 threshold=350
 
+risk=[]
+riskstd=[]
+t_fib_data = []
+t_fib_temp = []
+number_fib = []
+n_fib_data = []
+n_fib_avg = []
+n_fib_err = []
+t_in_fib = []
 
+#fileobj0 = open('realizationdata', 'wb')
 
-#opening pickled files for reinstatement of connections
+d=open('dysf_grids_sph_run1.pkl','rb')
+dysfgrids=pickle.load(d)
+d.close()
+
+dx=open('dysf_grids_sph_run1_extra_data.pkl','rb')
+dysfgrids_x=pickle.load(dx)
+dx.close()
+
+c=open('connections_sph_run1.pkl','rb')
+connections=pickle.load(c)
+c.close()
+
+cx=open('connections_sph_run1_extra_data.pkl','rb')
+connections_x=pickle.load(cx)
+cx.close()
+
+#f=open('fibrosisgrids.pkl','rb')
+#fibrosisgrids=pickle.load(f)
+#f.close()
[…]

### 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
poster work

@@
 \graphicspath{{figures/}} % Location of the graphics files
+\usepackage{graphicx}
+\usepackage{wrapfig}
 
@@
 
-\begin{block}{Objectives}
-
-\begin{enumerate}
-\item Replicate original model of Atrial Fibrillation (AF) from Christensen \emph{et al.}
-\item Incorporate restitution of heart cells into the model. 
-\item Model on more realistic morphology via translating model onto a sphere and adding restitution. 
-\item Analyse to see if there is enhanced risk with the new model, using new morphology and model basis, to create a more accurate AF model.
-\end{enumerate}
-
-\end{block}
-
 %----------------------------------------------------------------------------------------
@@
 AF is also the most widespread cardiac condition, with over 30 millions of people worldwide suffering from it.
-In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ million pounds.
+In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ \textbf{ million pounds} .\\*
  Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time (also called fibrillation episodes).\\
@@
 
-\begin{block}{Phenomenology of Atrial Fibrillation}
+\begin{block}{Phenomenology of the Heart}
 
 	\begin{itemize}
-	\item Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres 		and finally from the ventricular endocardium to the ventricular epicardium.
+	\item \textbf{\textit{Heart Beats Propagation}}: Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres 		and finally from the ventricular endocardium to the ventricular epicardium.
 The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell membrane.
 
-	\item When the electric impulse propagates a single muscle cell goes through three stages:
+\begin{wrapfigure}{R}{0.3\textwidth}
+\centering
+\includegraphics[width=0.25\textwidth]{heart}
+\caption{\label{fig:frog1}This is a figure caption.}
+\end{wrapfigure}
+
+	\item \textbf{ \textit{Cells Excitation Rules}}: As the electric impulse propagates a single muscle cell goes through three stages:
 
@@
   		\item \textit{Excited state}: the muscle cell is excited by its excited neighbours and the Voltage of the cell is at its peak. 
-  		\item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again - this period is called \textit{refractory period)}.
+  		\item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again - this period is called \textit{refractory period}.
+The \textit{refractory period} duration is related to the heart beat rate and this relationship is called \textit{restitution relationship}.
 		\end{itemize}
 
-	\item Where the refractory period depends on the last time a cell was excited and is always fixed to a constant.\\
-
-	\item AF is correlated to the amount of \textbf{ fibrosis} in the cardiac tissue which generates a process called reentry. \\
-	During reentry the electric signal wavefront propagation breaks. It has been shown a clear link between fibrosis and AF
+	\item \textbf{\textit{ How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis} in the cardiac tissue which generates a process called \textit{ reentry}. \\
+	During \textit{ reentry} the electric signal wavefront propagation breaks. 
 	Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research. \\
-	The current research revolves around three possible approaches:
-	\end{itemize}
+	%The current research revolves around three possible approaches:
 
@@
 % \end{frame}
- Our research is based on expanding on the work done by Kim Christensen on CA (Cellular Automata) models \cite{Christensen}.
+ \item \textbf{\textit{Base for our Research}}: Our research is based on expanding on the work done by Kim Christensen which models the left atrium on a 2D surface.
+	\end{itemize}
+
+\end{block}
+\begin{block}{Our Objectives}
+
+\begin{enumerate}
+\item Replicate the model of a 2D cardiac atrium from Christensen \emph{et al.}
+\item  Translating the model onto a sphere in order to simulate the effects of a more realistic morphology. 
+\item Incorporate restitution of heart cells into the original model (where the \textit{refractory period} is set to a constant).
+\item Observe if there is a higher risk of Atrial fibrillation in the original model or in the new ones we developed.
+\end{enumerate}
 
@@
 
-\begin{block}{Results: Connectome}
+\begin{block}{Our Sphere model's Morphology}
 
@@
 \includegraphics[width=0.7\linewidth]{connectome}
-\caption{Hemisphere of possible connections defined on the sphere model.
-Each vertex corresponds to a face.
-Longitudinal connections are in red, transversal are in blue.
-Starts from a single pentagon on the icosphere and then can be seen to propagate outwards.
+\caption{The structure of our sphere model on which we implemented electric impulse propagation using a network.
+The connections in red are differentiated from the connection in blue in order to introduce fibrosis in the model.
+Fibrosis is indeed modelled by removing a percentage of the red connections.
 }
@@
 
-\begin{block}{Results: Sphere and Previous model}
+\begin{block}{Model on a Sphere}
+
+In research by Fedotov, a model of the atrium was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
+Inspired by this research, we built a new model which translates Kim Christensen model of a two dimensional atrium on a sphere
+
+%\end{block}
+%------------------------------------------------
+%\begin{block}{Results: Sphere and Previous model}
 
@@
 \includegraphics[width=0.8\linewidth]{combchristensenspherefib}
-\caption{Figures of: Christensen \emph{et al.}'s model at times of t=1, t=60 and t=470,for a, b and c, respectively, and the sphere model at times of t=1, t=41 and t=241 for d, e and f respectively.
-Normal wavefront propagation seen starting from pentagonal node in the leftmost figures.In figure e, the excitation wavefront propagates around the sphere, mimicking Christensen \emph{et al.}'s model, b.
-On the right, chaotic, fibrillatory activity has arisen from reentry circuit spontaneously manifesting itself in c and f.}
+\caption{A comparison between Christensen \emph{et al.}'s model  (a, b and c)  and the sphere model (for d, e and f ) evolving in time in a not-healty atrium.
+Normal wavefront propagation is seen starting in the leftmost figures .In figures b,e the excitation wavefront propagates unevenly due to fibrosis, starting reentry.
+On the right AF has fully arisen from both models generating chaotic fibrillatory activity.}
 \end{figure}
@@
 In a real heart the refractory period depends on the heart beat rate and on the last time each cell it was excited.
-The relationship between refractory period and heart beat rate is called restitution.
+As already mentioned before he relationship between refractory period and heart beat rate is called restitution.
+We introduced restitution in In Christensen \emph{et al.}'s model in order to study its effects.
 \end{block}
 
-\begin{block}{Model on a Sphere}
-
-\begin{itemize}
-\item In research by Fedotov, a cellular automaton model was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
-It was shown that both a decrease in refractory period and conduction velocity from a "healthy" myocardium state, where no irregular excitation waves above 300bpm are observed, resulted in an area where spiral waves were formed.
-In addition, a multi-electrode system was created to study the localization of AF sources by simulated electrograms.
-\end{itemize}
-
-\end{block}
-%------------------------------------------------
-
-\begin{block}{Results: Figure}
+\begin{block}{Results: Risk of AF for the three models}
 
@@
 \includegraphics[width=0.65\linewidth]{xriskcurvesphere}
-\caption{Figure of the Risk Curve of all the systems analysed. Y-axis: probablility of a realisation being in AF, x-axis: value of $\nu$, the  probability of a transversal connection between cells}
+\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied (Prof Christensen's model, restitution curve model, sphere model) averaged over 100 simulations against percentage of transversal connections($\nu$) in the model.
+A healthy heart is expected to have $\nu$ close to 1 while a heart with fibrosis low $\nu$.}
 \end{figure}
@@
 \begin{block}{Conclusion}
-\begin{itemize}
-\item A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of heart muscle fibre connectivity while incorporating the effects of fibrosis.
-This model was improved upon by the additon of the restitution process, making AF more risky.
-By translating the model to a sphere, the steady growth of heart muscle fibres from the sino-atrial node could be replicated, while simulating a more realisitc morphology of the atria.
-The risk of AF was seen to be higher, than with the original model.
-Verification of this model can be achieved with the analysis of the proportion of fibrosis in atrial heart tissue.
-\end{itemize}
+
+A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of the atrium muscle tissue while incorporating the effects of fibrosis.
+We implemented this model by adding more realistic features by using a more realistic heart morphology and restitution.
+Our results show a higher risk of AF for the new models created.
+
 \end{block}

### 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
poster

@@
 
-\begin{itemize}
-\item Atrial Fibrillation (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
+Atrial Fibrillation (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
 AF is also the most widespread cardiac condition, with over 30 millions of people worldwide suffering from it.
-In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ \textbf{ million pounds} .\\*
- Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time (also called fibrillation episodes).\\
-\end{itemize}
+In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ \textbf{million pounds}.\\*
+ Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time.\\
 
@@
 
+\begin{columns}
+\begin{column}{.54\textwidth}
 	\begin{itemize}
-	\item \textbf{\textit{Heart Beats Propagation}}: Each heart beat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres 		and finally from the ventricular endocardium to the ventricular epicardium.
-The electric signal is conducted in the cardiac muscle cells thanks to the polarisation mechanism of the cell membrane.
+	\item \textbf{\textit{Heartbeat Propagation}}: Each heartbeat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres, and finally from the ventricular endocardium to the ventricular epicardium.
+The electric signal is conducted in the cardiac muscle cells thanks to polarisation (change in voltage) of the cell membrane.
+	\end{itemize}
+\end{column}
 
-\begin{wrapfigure}{R}{0.3\textwidth}
-\centering
-\includegraphics[width=0.25\textwidth]{heart}
-\caption{\label{fig:frog1}This is a figure caption.}
-\end{wrapfigure}
+\begin{column}{.43\textwidth}
+	%\begin{wrapfigure}[10]{R}{0.3\textwidth}
+	\begin{figure}
 
-	\item \textbf{ \textit{Cells Excitation Rules}}: As the electric impulse propagates a single muscle cell goes through three stages:
+	\includegraphics[width=0.95\textwidth]{heart}
+	\caption{\label{fig:frog1}This is the structure of the heart.}
+	%\end{wrapfigure}
+	\end{figure}
+\end{column}
+\end{columns}
+\begin{itemize}
+
+	\item \textbf{ \textit{Cells Excitation Rules}}: As the electric impulse propagates, each cardiac single muscle cell can be in one of three stages:
 
@@
 
-  		\item \textit{Excited state}: the muscle cell is excited by its excited neighbours and the Voltage of the cell is at its peak. 
-  		\item \textit{Refractory state} :the muscle cell goes through a phase during which it can't be excited again - this period is called \textit{refractory period}.
-The \textit{refractory period} duration is related to the heart beat rate and this relationship is called \textit{restitution relationship}.
+  		\item \textit{Excited state}: the muscle cell is excited by the excited cells it is connected to and its voltage is at its peak. 
+  		\item \textit{Refractory state}: the muscle cell goes through a phase during which it can't be excited again, a decrease in voltage---this period is called \textit{refractory period}.
+The refractory period duration is related to the heartbeat rate and this relationship is called \textit{restitution}.
 		\end{itemize}
 
-	\item \textbf{\textit{ How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis} in the cardiac tissue which generates a process called \textit{ reentry}. \\
-	During \textit{ reentry} the electric signal wavefront propagation breaks. 
-	Even though many treatments have been developed, the mechanisms behind this condition are not completely understood and AF remains a major topic in medical research. \\
-	%The current research revolves around three possible approaches:
-
-%\begin{frame}
-%\smartdiagram[bubble diagram]{Current research,
- %Physiological \\ models, Image-Based \\ Models, \textbf{ CA Models}, antiarrhythmic \\ drugs) }
-% \end{frame}
- \item \textbf{\textit{Base for our Research}}: Our research is based on expanding on the work done by Kim Christensen which models the left atrium on a 2D surface.
 \end{itemize}
 
+\begin{columns}
+\begin{column}{.43\textwidth}
+\begin{figure}
+	\includegraphics[width=0.95\textwidth]{spiralbreak2tachy}
+	\caption{\label{fig:frog1}Idealised diagram of: (a), normal heatbeat propagation; (b), tachycardia and (c), fibrillation. }
+	%\end{wrapfigure}
+	\end{figure}
+
+\end{column}
+\begin{column}{.54\textwidth}
+\begin{itemize}
+	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---which is the scarring of cells.
+This creates dysfunctional/inactive cells, decreasing the number of connections between the cells in the cardiac tissue.
+This promotes a process called \textit{ reentry}.,\\
+	during which, the electric signal wavefront propagation breaks. 
+	Even though many treatments have been developed, the reentry mechanism is not completely understood and , as such, AF remains a major topic in medical research. \\
+	\end{itemize}
+
+\end{column}
+
+\end{columns}
+
+	\begin{itemize}
+
+	%The current research revolves around three possible approaches:
+ \item \textbf{\textit{Base for our Research}}: Our research is based on expanding on the work done by Christensen \emph{et al.} which models the left atrium on a 2D surface \cite{Christensen}.
+	\end{itemize}
 \end{block}
@@
 \item  Translating the model onto a sphere in order to simulate the effects of a more realistic morphology. 
-\item Incorporate restitution of heart cells into the original model (where the \textit{refractory period} is set to a constant).
-\item Observe if there is a higher risk of Atrial fibrillation in the original model or in the new ones we developed.
+\item Incorporate restitution of heart cells into the original model (where the \textit{refractory period} is instead set to a constant).
+\item Observe the risk of AF in the new models and compare it with the original model.
 \end{enumerate}
@@
 \begin{figure}
-\includegraphics[width=0.7\linewidth]{connectome}
+\includegraphics[width=0.6\linewidth]{connectome}
 \caption{The structure of our sphere model on which we implemented electric impulse propagation using a network.
@@
 
-\begin{block}{Model on a Sphere}
+\begin{block}{Implementing Our Sphere Model}
 
-In research by Fedotov, a model of the atrium was constructed on the surface of a triangulated sphere to model spiral rotor waves on a closed, heterogeneous surface \cite{Fedotov}.
-Inspired by this research, we built a new model which translates Kim Christensen model of a two dimensional atrium on a sphere
+In research by Fedotov, a model of the atrium was constructed on the surface of a triangulated sphere to AF on a closed, heterogeneous surface \cite{Fedotov}.
+Inspired by this research, we built a new model which translates Kim Christensen model of a two dimensional atrium onto a sphere.
 
@@
 
-\begin{block}{Implementing Restitution }
+\begin{block}{Implementing the Restitution Model}
 
-In Christensen \emph{et al.}'s model, the refractory period of each muscle cell was always fixed, this is not what happens in a real heart.
-In a real heart the refractory period depends on the heart beat rate and on the last time each cell it was excited.
-As already mentioned before he relationship between refractory period and heart beat rate is called restitution.
-We introduced restitution in In Christensen \emph{et al.}'s model in order to study its effects.
+In Christensen \emph{et al.}'s model, the refractory period of each muscle cell was always fixed to a constant.
+In a real heart, the refractory period depends on the heart beat rate and on the last time each cell it was excited.
+The relationship between refractory period and heart beat rate is called restitution.
+We introduced a linear restitution function in  Christensen \emph{et al.}'s model in order to study its effects.
 \end{block}
@@
 \includegraphics[width=0.65\linewidth]{xriskcurvesphere}
-\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied (Prof Christensen's model, restitution curve model, sphere model) averaged over 100 simulations against percentage of transversal connections($\nu$) in the model.
+\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied (Prof Christensen's model, restitution curve model, sphere model) averaged over 100 simulations against percentage of transversal connections ($\nu$) in the model.
 A healthy heart is expected to have $\nu$ close to 1 while a heart with fibrosis low $\nu$.}
@@
 \begin{block}{References}
+\begin{thebibliography}{1}
 
-\nocite{*} % Insert publications even if they are not cited in the poster
-\small{\bibliographystyle{unsrt}
-\bibliography{sample1}}
+%\nocite{*} % Insert publications even if they are not cited in the poster
+%\small{\bibliographystyle{unsrt}
+%\bibliography{sample1}}
+\bibitem{Comtois}
+P. Comtois, J. Kneller, S. Nattel,
+\emph{Of circles and spirals: Bridging the gap between the leading circle and spiral wave concepts of cardiac reentry}
+Europace. 2005 Sep;7 Suppl 2:10-20.
+DOI: http://dx.doi.org/10.1016/j.eupc.2005.05.011
+
+\bibitem{Alonso}
+S. Alonso, M. Bar and B. Echebarria
+\emph{Nonlinear physics of electrical wave
+propagation in the heart: a review},
+Rep. Prog. Phys. 79 (2016) 096601, 
+doi:10.1088/0034-4885/79/9/096601
+
+\bibitem{Christensen}
+Kim Christensen, Kishan A. Manani, and Nicholas S. Peters
+\emph{Simple Model for Identifying Critical Regions in Atrial Fibrillation}
+Phys. Rev. Lett. 114, 028104 – Published 16 January 2015
+
+\bibitem{Fedotov}
+N. M. Fedotov,  A. I. Oferkin, and S. V. Zhary,
+\emph{Modeling Sources of Atrial Fibrillation on a Triangulated Sphere}
+Biomedical Engineering, Vol. 49, No. 2, July, 2015, pp. 112115. 
+\end{thebibliography}
 

### 2017-02-23 Sphere Model/New Network With Proper Propagation/Network_Projection_withrestitution.py

@@ -189,6 +189,7 @@ class create_network:
         self.nodes[self.nodes==-1]=0
         
         #maybe switch these two
+        self.tempgrid[self.tempgrid>20000] = 220
         self.refr=self.maparray[self.tempgrid.astype(int)]
         
       

### 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
r

@@
 	\begin{itemize}
-	\item \textbf{\textit{Heartbeat Propagation}}: Each heartbeat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres, and finally from the ventricular endocardium to the ventricular epicardium.
-The electric signal is conducted in the cardiac muscle cells thanks to polarisation (change in voltage) of the cell membrane.
+	\item \textbf{\textit{Heartbeat Propagation}}: Each heartbeat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres, and finally from the ventricular endocardium to the ventricular epicardium (Figure \ref{fig:heart}).
+The electric signal is conducted in the cardiac muscle cells thanks to polarisation (change in voltage) of the cell membrane \cite{Comtois}.
 	\end{itemize}
@@
 	\includegraphics[width=0.95\textwidth]{heart}
-	\caption{\label{fig:frog1}This is the structure of the heart.}
+	\caption{\label{fig:heart}This is the structure of the heart.}
 	%\end{wrapfigure}
@@
 	\includegraphics[width=0.95\textwidth]{spiralbreak2tachy}
-	\caption{\label{fig:frog1}Idealised diagram of: (a), normal heatbeat propagation; (b), tachycardia and (c), fibrillation. }
+	\caption{\label{fig:fib}Idealised diagram of: (a), normal heatbeat propagation; (b), tachycardia and (c), fibrillation. }
 	%\end{wrapfigure}
@@
 This promotes a process called \textit{ reentry}.,\\
-	during which, the electric signal wavefront propagation breaks. 
+	during which, the electric signal wavefront propagation breaks \cite{Alonso}. 
 	Even though many treatments have been developed, the reentry mechanism is not completely understood and , as such, AF remains a major topic in medical research. \\
@@
 
-In Christensen \emph{et al.}'s model, the refractory period of each muscle cell was always fixed to a constant.
+In Christensen \emph{et al.}'s model \cite{Christensen}, the refractory period of each muscle cell was always fixed to a constant.
 In a real heart, the refractory period depends on the heart beat rate and on the last time each cell it was excited.

### 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
ss

@@
 \usepackage{wrapfig}
+%\usepackage{floatrow}
 
@@
 
-Atrial Fibrillation (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
-AF is also the most widespread cardiac condition, with over 30 millions of people worldwide suffering from it.
-In the UK, expenses related to this disease account for more than $1\%$ of the total budget of the National Health Service, which is more than $800$ \textbf{million pounds}.\\*
+\textbf{Atrial Fibrillation} (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
+AF is also the most widespread cardiac condition, with over 30 million people worldwide suffering from it.
+In the UK, expenses related to this disease account for more than 1 percent of the total budget of the National Health Service, which is more than 800 million pounds.\\*
  Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time.\\
@@
 	\begin{itemize}
-	\item \textbf{\textit{Heartbeat Propagation}}: Each heartbeat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres, and finally from the ventricular endocardium to the ventricular epicardium (Figure \ref{fig:heart}).
+	\item \textbf{\textit{Heartbeat Propagation}}: Each heartbeat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres, into the lower reaches of the heart (Figure \ref{fig:heart}).
 The electric signal is conducted in the cardiac muscle cells thanks to polarisation (change in voltage) of the cell membrane \cite{Comtois}.
@@
 
-	\includegraphics[width=0.95\textwidth]{heart}
+	\includegraphics[width=0.85\textwidth]{heart}
 	\caption{\label{fig:heart}This is the structure of the heart.}
@@
 		\begin{itemize}
- 		 \item \textit{Excitable state}: the muscle cell is at rest with a negative built up potential.
+ 		 \item \textit{Rest/Excitable state}: the muscle cell is at rest with a negative built up potential.
 
-  		\item \textit{Excited state}: the muscle cell is excited by the excited cells it is connected to and its voltage is at its peak. 
-  		\item \textit{Refractory state}: the muscle cell goes through a phase during which it can't be excited again, a decrease in voltage---this period is called \textit{refractory period}.
-The refractory period duration is related to the heartbeat rate and this relationship is called \textit{restitution}.
+  		\item \textit{Excited state}: the muscle cell is excited by the excited cells it is connected to and its voltage is at a maximum. 
+  		\item \textit{Refractory state}: the muscle cell goes through a phase during which it can't be excited again, a decrease in voltage until its rest state voltage---this period is called \textit{refractory period}.
+The refractory period duration is modulated by the cell excitation rate and this relationship is called \textit{restitution}.
 		\end{itemize}
@@
 	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---which is the scarring of cells.
-This creates dysfunctional/inactive cells, decreasing the number of connections between the cells in the cardiac tissue.
-This promotes a process called \textit{ reentry}.,\\
-	during which, the electric signal wavefront propagation breaks \cite{Alonso}. 
-	Even though many treatments have been developed, the reentry mechanism is not completely understood and , as such, AF remains a major topic in medical research. \\
+This creates dysfunctional/inactive cells, decreasing the number of connections between the cells in cardiac tissue.
+This promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation breaks \cite{Alonso}.
+Even though many treatments have been developed, the reentry mechanism is not completely understood and, as such, AF remains a major topic in medical research.
+\\
 	\end{itemize}
@@
 	%The current research revolves around three possible approaches:
- \item \textbf{\textit{Base for our Research}}: Our research is based on expanding on the work done by Christensen \emph{et al.} which models the left atrium on a 2D surface \cite{Christensen}.
+ \item \textbf{\textit{Base for our Research}}: Our research is based on the work done by Christensen \emph{et al.}, which models the left atrium of the heart on a 2D surface \cite{Christensen}.
 	\end{itemize}
@@
 \item Replicate the model of a 2D cardiac atrium from Christensen \emph{et al.}
-\item  Translating the model onto a sphere in order to simulate the effects of a more realistic morphology. 
-\item Incorporate restitution of heart cells into the original model (where the \textit{refractory period} is instead set to a constant).
-\item Observe the risk of AF in the new models and compare it with the original model.
+\item  Translate the model onto a sphere in order to simulate the effects of a more realistic morphology. 
+\item Incorporate restitution of heart cells into the original model (such that the \textit{refractory period} of cells is modulated).
+\item Observe the risk of AF in the new models and compare with the original model.
 \end{enumerate}
@@
 
-\begin{block}{Our Sphere model's Morphology}
+\begin{block}{Sphere Model Morphology}
 
@@
 \begin{figure}
-\includegraphics[width=0.6\linewidth]{connectome}
-\caption{The structure of our sphere model on which we implemented electric impulse propagation using a network.
-The connections in red are differentiated from the connection in blue in order to introduce fibrosis in the model.
-Fibrosis is indeed modelled by removing a percentage of the red connections.
+
+\begin{columns}
+\begin{column}{.67\textwidth}
+\includegraphics[width=0.8\linewidth]{connectome}
+
+\end{column}
+\hspace{-6em}
+\begin{column}{.3\textwidth}
+\caption{The structure of the sphere model upon which electric impulse propagation was implemented using a network.
+The connections in red are differentiated from the connection in blue in order to introduce fibrosis to the model.
+An effect of fibrosis is modelled by removing a percentage of the red connections.
 }
+
+\end{column}
+\end{columns}
 \end{figure}
@@
 
-\begin{block}{Implementing Our Sphere Model}
+\begin{block}{Implementing The Sphere Model}
 
-In research by Fedotov, a model of the atrium was constructed on the surface of a triangulated sphere to AF on a closed, heterogeneous surface \cite{Fedotov}.
-Inspired by this research, we built a new model which translates Kim Christensen model of a two dimensional atrium onto a sphere.
+In research by Fedotov, a model of the atrium was constructed on the surface of a triangulated sphere to model AF on a closed, heterogeneous surface \cite{Fedotov}.
+Inspired by this research, we built a new model which translates Christensen \emph{et al.}'s model of a two dimensional atrium onto a sphere.
 
@@
 \includegraphics[width=0.8\linewidth]{combchristensenspherefib}
-\caption{A comparison between Christensen \emph{et al.}'s model  (a, b and c)  and the sphere model (for d, e and f ) evolving in time in a not-healty atrium.
-Normal wavefront propagation is seen starting in the leftmost figures .In figures b,e the excitation wavefront propagates unevenly due to fibrosis, starting reentry.
-On the right AF has fully arisen from both models generating chaotic fibrillatory activity.}
+\caption{A comparison between Christensen \emph{et al.}'s model  (a, b and c)  and the sphere model (for d, e and f ) evolving in time in a fibrotic atrium.
+Normal wavefront propagation is seen starting in the leftmost figures .In figures b,e the excitation wavefront propagates unevenly due to fibrosis, initiating reentry.
+On the right, AF has spontaneously arisen from both models, generating fibrillation.}
 \end{figure}
@@
 
-In Christensen \emph{et al.}'s model \cite{Christensen}, the refractory period of each muscle cell was always fixed to a constant.
-In a real heart, the refractory period depends on the heart beat rate and on the last time each cell it was excited.
+In Christensen \emph{et al.}'s model \cite{Christensen}, the refractory period of each muscle cell was fixed to be constant.
+In a real heart, the refractory period depends on the heart cell activation rate.
 The relationship between refractory period and heart beat rate is called restitution.
@@
 \includegraphics[width=0.65\linewidth]{xriskcurvesphere}
-\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied (Prof Christensen's model, restitution curve model, sphere model) averaged over 100 simulations against percentage of transversal connections ($\nu$) in the model.
-A healthy heart is expected to have $\nu$ close to 1 while a heart with fibrosis low $\nu$.}
+\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied (Christensen /emph{et al.}'s model, restitution curve model and sphere model) averaged over 50 simulations, against the percentage of transversal connections ($\nu$) in the models.
+A healthy heart is expected to have $\nu$ close to 1 while a fibrotic heart has low $\nu$.}
 \end{figure}
@@
 A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of the atrium muscle tissue while incorporating the effects of fibrosis.
-We implemented this model by adding more realistic features by using a more realistic heart morphology and restitution.
+We implemented this model by adding more features: a more realistic heart morphology, using a sphere, and restitution.
 Our results show a higher risk of AF for the new models created.

### 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
ss

@@
 In the UK, expenses related to this disease account for more than 1 percent of the total budget of the National Health Service, which is more than 800 million pounds.\\*
- Atrial fibrillation usually manifests itself in patients affected through alterations of the normal cardiac rhythm lasting short periods of time.\\
+ Atrial fibrillation usually manifests itself in patients with abnormal cardiac rhythm lasting short periods of time.\\
 

### 2017-02-23 Literature Review/Atrial Fibrillation Poster FINAL.tex
final tingz

--- 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
@@
 \textbf{Atrial Fibrillation} (AF) is a type of cardiac arrhythmia and one of the major causes of stroke and heart failure.
-AF is also the most widespread cardiac condition, with over 30 million people worldwide suffering from it.
-In the UK, expenses related to this disease account for more than 1 percent of the total budget of the National Health Service, which is more than 800 million pounds.\\*
- Atrial fibrillation usually manifests itself in patients with abnormal cardiac rhythm lasting short periods of time.\\
+AF is also the most widespread cardiac condition, with over 30 million people suffering from it worldwide.
+In the UK, expenses related to this disease account for more than 1 percent of the total National Health Service budget, costing more than 800 million pounds.\\*
+ Atrial fibrillation usually manifests itself in patients with tachycardia, an abnormal hastening of cardiac beating.\\
 
@@
 	\item \textbf{\textit{Heartbeat Propagation}}: Each heartbeat originates as an electric signal in the sinoatrial (SA) node that propagates first into the atria, then through the atrioventricular (AV) node, through Purkinje fibres, into the lower reaches of the heart (Figure \ref{fig:heart}).
-The electric signal is conducted in the cardiac muscle cells thanks to polarisation (change in voltage) of the cell membrane \cite{Comtois}.
+The electric signal is conducted through the cardiac muscle cells due to polarisation (a change in voltage of the cell membrane) which contracts the cell \cite{Comtois}.
 	\end{itemize}
@@
 
-	\item \textbf{ \textit{Cells Excitation Rules}}: As the electric impulse propagates, each cardiac single muscle cell can be in one of three stages:
+	\item \textbf{ \textit{Cells Excitation Rules}}: As the electric impulse propagates, each cardiac muscle cell can be in one of three stages:
 
 		\begin{itemize}
- 		 \item \textit{Rest/Excitable state}: the muscle cell is at rest with a negative built up potential.
-
-  		\item \textit{Excited state}: the muscle cell is excited by the excited cells it is connected to and its voltage is at a maximum. 
-  		\item \textit{Refractory state}: the muscle cell goes through a phase during which it can't be excited again, a decrease in voltage until its rest state voltage---this period is called \textit{refractory period}.
-The refractory period duration is modulated by the cell excitation rate and this relationship is called \textit{restitution}.
+ 		 \item \textit{Rest/Excitable state}: the muscle cell is at rest with constant negative potential.								
+  		\item \textit{Excited state}: the muscle cell is excited by adjacent excited cells it is connected to, making the voltage is at a maximum. 
+  		\item \textit{Refractory state}: the muscle cell goes through a phase during which it can't be excited again---a steady decrease in voltage until its matches the rest state voltage---this period is called \textit{refractory period}.
+The refractory period duration is modulated by the cell excitation rate; this relationship is called \textit{restitution}.
 		\end{itemize}
@@
 	\includegraphics[width=0.95\textwidth]{spiralbreak2tachy}
-	\caption{\label{fig:fib}Idealised diagram of: (a), normal heatbeat propagation; (b), tachycardia and (c), fibrillation. }
+	\caption{\label{fig:fib}Idealised diagram of: (a), normal heartbeat propagation; (b), tachycardia and (c), fibrillation. }
 	%\end{wrapfigure}
@@
 \begin{itemize}
-	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---which is the scarring of cells.
-This creates dysfunctional/inactive cells, decreasing the number of connections between the cells in cardiac tissue.
-This promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation breaks \cite{Alonso}.
+	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---the scarring of cells.
+This creates dysfunctional/inactive cells, decreasing the number of connections between cells in the cardiac tissue.
+This promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation chaotically breaks, initiating fibrillation \cite{Alonso}.
 Even though many treatments have been developed, the reentry mechanism is not completely understood and, as such, AF remains a major topic in medical research.
@@
 \begin{enumerate}
-\item Replicate the model of a 2D cardiac atrium from Christensen \emph{et al.}
+\item Replicate the model of the 2D cardiac atrium from Christensen \emph{et al.}
 \item  Translate the model onto a sphere in order to simulate the effects of a more realistic morphology. 
@@
 \begin{column}{.3\textwidth}
-\caption{The structure of the sphere model upon which electric impulse propagation was implemented using a network.
-The connections in red are differentiated from the connection in blue in order to introduce fibrosis to the model.
-An effect of fibrosis is modelled by removing a percentage of the red connections.
-}
+\caption{The structure of the sphere model.
+The connections in red (muscle fibres) are differentiated from the connection in blue (connections between fibres) in order to introduce fibrosis to the model.
+The effect of fibrosis is modelled by removing a percentage of the blue connections.
+Fibre growth from the SA node is seen.}
 
@@
 In research by Fedotov, a model of the atrium was constructed on the surface of a triangulated sphere to model AF on a closed, heterogeneous surface \cite{Fedotov}.
-Inspired by this research, we built a new model which translates Christensen \emph{et al.}'s model of a two dimensional atrium onto a sphere.
+Inspired by this, we built a new model which translates Christensen \emph{et al.}'s model of a two dimensional atrium onto a sphere.
 
@@
 Normal wavefront propagation is seen starting in the leftmost figures .In figures b,e the excitation wavefront propagates unevenly due to fibrosis, initiating reentry.
-On the right, AF has spontaneously arisen from both models, generating fibrillation.}
+On the right, AF has spontaneously arisen from both models.}
 \end{figure}
@@
 
-In Christensen \emph{et al.}'s model \cite{Christensen}, the refractory period of each muscle cell was fixed to be constant.
+In Christensen \emph{et al.}'s model, the refractory period of each muscle cell was fixed.
 In a real heart, the refractory period depends on the heart cell activation rate.
-The relationship between refractory period and heart beat rate is called restitution.
-We introduced a linear restitution function in  Christensen \emph{et al.}'s model in order to study its effects.
+The relationship between refractory period and cell excitation rate is called restitution.
+We introduced a linear restitution function into our replica of Christensen \emph{et al.}'s model in order to study its effects.
 \end{block}
@@
 \includegraphics[width=0.65\linewidth]{xriskcurvesphere}
-\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied (Christensen /emph{et al.}'s model, restitution curve model and sphere model) averaged over 50 simulations, against the percentage of transversal connections ($\nu$) in the models.
-A healthy heart is expected to have $\nu$ close to 1 while a fibrotic heart has low $\nu$.}
+\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied: Christensen \emph{et al.}'s model ("Kishan's Data"); our original model; a restitution model and the sphere model.
+Each point is an average risk over 50 simulations, using a specified percentage of transversal connections ($\nu$), mimicking fibrosis levels.
+A healthy heart is expected to have $\nu$ close to 1 while a fibrotic heart has low $\nu$.
+Incorporating restitution into original model increases the risk.
+The sphere model incorporating fibre growth increases the risk further, producing a more realistic model as less fibrosis is necessary to induce AF.
+}
 \end{figure}
@@
 
-A simplified mechanism of AF had been achieved by Christensen \emph{et al.'s} model through the modelling of the atrium muscle tissue while incorporating the effects of fibrosis.
-We implemented this model by adding more features: a more realistic heart morphology, using a sphere, and restitution.
+A simplified mechanism of AF had been achieved by replication of Christensen \emph{et al.'s} model: simulating a sheet of atrial muscle tissue and incorporating the effects of fibrosis.
+We incorporated more features into this model: a more realistic heart morphology,with muscle fibre growth, using a sphere, and restitution.
 Our results show a higher risk of AF for the new models created.
-
+This produced more realistic models as less fibrosis is necessary to induce AF.
 \end{block}
@@
 %\bibliography{sample1}}
+\small
 \bibitem{Comtois}

### 2017-02-23 Literature Review/Atrial Fibrillation Poster 22feb.tex
final changes

@@
 \begin{itemize}
-	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---which is the scarring of cells.
-This creates dysfunctional/inactive cells, decreasing the number of connections between the cells in cardiac tissue.
-This promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation breaks \cite{Alonso}.
+	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---the scarring of cells.
+This scarring creates dysfunctional/inactive cells, decreasing the number of connections between cells in the cardiac tissue.
+The absence of connections promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation chaotically breaks, initiating fibrillation \cite{Alonso}.
 Even though many treatments have been developed, the reentry mechanism is not completely understood and, as such, AF remains a major topic in medical research.

### 2017-02-23 Literature Review/Atrial Fibrillation Poster FINAL.tex
dsgafd

@@
  		 \item \textit{Rest/Excitable state}: the muscle cell is at rest with constant negative potential.								
-  		\item \textit{Excited state}: the muscle cell is excited by adjacent excited cells it is connected to, making the voltage is at a maximum. 
+  		\item \textit{Excited state}: the muscle cell is excited by adjacent excited cells it is connected to, making the voltage a maximum. 
   		\item \textit{Refractory state}: the muscle cell goes through a phase during which it can't be excited again---a steady decrease in voltage until its matches the rest state voltage---this period is called \textit{refractory period}.
@@
 \begin{itemize}
-	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{ fibrosis}---the scarring of cells.
-This creates dysfunctional/inactive cells, decreasing the number of connections between cells in the cardiac tissue.
-This promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation chaotically breaks, initiating fibrillation \cite{Alonso}.
+	\item \textbf{\textit{How AF Occurs}}: AF is correlated to the amount of \textbf{fibrosis}---the scarring of cells.
+This scarring creates dysfunctional/inactive cells, decreasing the number of connections between cells in the cardiac tissue.
+The absence of connections promotes a process called \textit{reentry}, during which, the electric signal wavefront propagation chaotically breaks, initiating fibrillation \cite{Alonso}.
 Even though many treatments have been developed, the reentry mechanism is not completely understood and, as such, AF remains a major topic in medical research.
@@
 \item Replicate the model of the 2D cardiac atrium from Christensen \emph{et al.}
-\item  Translate the model onto a sphere in order to simulate the effects of a more realistic morphology. 
+\item  Translate the model onto a sphere in order to simulate the effects of a more realistic morphology and muscle fibre growth. 
 \item Incorporate restitution of heart cells into the original model (such that the \textit{refractory period} of cells is modulated).
@@
 \caption{A comparison between Christensen \emph{et al.}'s model  (a, b and c)  and the sphere model (for d, e and f ) evolving in time in a fibrotic atrium.
-Normal wavefront propagation is seen starting in the leftmost figures .In figures b,e the excitation wavefront propagates unevenly due to fibrosis, initiating reentry.
+Normal wavefront propagation is seen starting in the leftmost figures.
+In figures b and e the excitation wavefront propagates unevenly due to fibrosis, initiating reentry.
 On the right, AF has spontaneously arisen from both models.}
@@
 \includegraphics[width=0.65\linewidth]{xriskcurvesphere}
-\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the three models studied: Christensen \emph{et al.}'s model ("Kishan's Data"); our original model; a restitution model and the sphere model.
-Each point is an average risk over 50 simulations, using a specified percentage of transversal connections ($\nu$), mimicking fibrosis levels.
+\caption{Risk of AF (defined as the fraction of time spent in fibrillation) for the models studied: Christensen \emph{et al.}'s model ("Kishan's Data"); our original model; a restitution model and the sphere model.
+Each point represents an average risk over 50 simulations, using a specified percentage of transversal connections ($\nu$), mimicking fibrosis levels.
 A healthy heart is expected to have $\nu$ close to 1 while a fibrotic heart has low $\nu$.
 Incorporating restitution into original model increases the risk.
-The sphere model incorporating fibre growth increases the risk further, producing a more realistic model as less fibrosis is necessary to induce AF.
+The sphere model increases the risk further, producing a more realistic model as less fibrosis is necessary to induce AF.
 }
@@
 
-A simplified mechanism of AF had been achieved by replication of Christensen \emph{et al.'s} model: simulating a sheet of atrial muscle tissue and incorporating the effects of fibrosis.
-We incorporated more features into this model: a more realistic heart morphology,with muscle fibre growth, using a sphere, and restitution.
-Our results show a higher risk of AF for the new models created.
-This produced more realistic models as less fibrosis is necessary to induce AF.
+A simplified mechanism of AF had been achieved by replication of Christensen \emph{et al.}'s model: simulating a sheet of atrial muscle tissue and incorporating the effects of fibrosis.
+We introduced more features into this model: a more realistic heart morphology, with muscle fibre growth, using a sphere, and restitution.
+Our results show a higher risk of AF for the new models created, resulting in more realistic models, as less fibrosis is necessary to induce AF.
 \end{block}
@@
 P. Comtois, J. Kneller, S. Nattel,
-\emph{Of circles and spirals: Bridging the gap between the leading circle and spiral wave concepts of cardiac reentry}
-Europace. 2005 Sep;7 Suppl 2:10-20.
+\emph{Of circles and spirals: Bridging the gap between the leading circle and spiral wave concepts of cardiac reentry}, 
+Europace. 2005 Sep;7 Suppl 2:10-20., 
 DOI: http://dx.doi.org/10.1016/j.eupc.2005.05.011
@@
 \bibitem{Alonso}
-S. Alonso, M. Bar and B. Echebarria
+S. Alonso, M. Bar and B. Echebarria, 
 \emph{Nonlinear physics of electrical wave
@@
 \bibitem{Christensen}
-Kim Christensen, Kishan A. Manani, and Nicholas S. Peters
-\emph{Simple Model for Identifying Critical Regions in Atrial Fibrillation}
+Kim Christensen, Kishan A. Manani, and Nicholas S. Peters, 
+\emph{Simple Model for Identifying Critical Regions in Atrial Fibrillation}, 
 Phys. Rev. Lett. 114, 028104 – Published 16 January 2015
@@
 N. M. Fedotov,  A. I. Oferkin, and S. V. Zhary,
-\emph{Modeling Sources of Atrial Fibrillation on a Triangulated Sphere}
+\emph{Modeling Sources of Atrial Fibrillation on a Triangulated Sphere}, 
 Biomedical Engineering, Vol. 49, No. 2, July, 2015, pp. 112115. 
